%Copyright © 2025 Brian Pomerantz. All Rights Reserved.

\documentclass[11pt,twoside]{memoir}

%------------------------------------------------
% Page layout for half-letter (5.5x8.5)
%------------------------------------------------
\setstocksize{8.5in}{5.5in}            % Final trimmed page size
\settrimmedsize{8.5in}{5.5in}{*}       % Trimmed size
\settypeblocksize{7in}{4.25in}{*}      % Text area (height x width)
%\setbinding{0.125in}
\setlrmargins{*}{*}{1.2}               % Inner/outer margin ratio
\setulmargins{*}{*}{1.2}               % Top/bottom margin ratio
\setheaderspaces{*}{*}{1.0}
\checkandfixthelayout
\setlength{\parindent}{1em}
\setlength{\parskip}{0.5em}

%------------------------------------------------
% Fonts
%------------------------------------------------
\usepackage{textcomp}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\linespread{1.1}

%------------------------------------------------
% Formatting
%------------------------------------------------
\pagestyle{myheadings}

\newcommand{\mychapter}[1]{%
  \chapter*{#1}
  \thispagestyle{empty}
  \markboth{Huck Finn's All Right}{#1}
  \addcontentsline{toc}{chapter}{#1}
}

\addtocontents{toc}{\protect\thispagestyle{empty}}

%------------------------------------------------
% Front matter
%------------------------------------------------
\title{Huck Finn's All Right}
\author{Abigail Haynes}
\date{}

\begin{document}
\frontmatter
\maketitle
\thispagestyle{empty}
\clearpage

% Copyright page
\thispagestyle{empty}
\vspace*{\fill}
\begin{center}
\tiny
Copyright \textcopyright\ 2024-2025 Abigail Haynes. All rights reserved.\\
No part of this book may be reproduced or transmitted\\
in any form without written permission of the author.\\
Printed in the United States of America
\end{center}
\vspace*{\fill}
\clearpage


% Table of contents
\tableofcontents*
\cleardoublepage

%\markboth{Introduction}{Introduction}
%\chapter{Introduction}
%\addcontentsline{toc}{chapter}{Introduction}
%\lipsum[1-10]

\book*{Huck Finn's All Right}
\thispagestyle{empty}


%------------------------------------------------
% Main matter
%------------------------------------------------
\mainmatter

\mychapter{Chapter 1}

Well, I was sound asleep in my old hogshead when a cat started howling
away out in the dark, really going at it, so ornery it waked me up. I
hove a shoe at it to get it to shut up, and heard a muffled curse.

I sat up so fast my head hit the top of the barrel (which was really the
side, on account of my barrel was overturned). There was a scrabbling
out in the dark somewheres. Clutching my scalp, I scanned the dark yard
behind the slaughter-house.

``Wh---who's there?'' My voice came out kind of a rasp.

``Who do you think, flathead?'' says a voice. ``And what did you have to
go and do a thing like that for? What kind of fellow chucks a shoe at
another fellow---''

I breathed out in a whoosh.

``Tom Sawyer, is that you?'' Then I nearly got my nose taken off as he
plunged into my hogshead, a beat-up tin lantern swinging wildly from his
hand.

``Didn't we \emph{say} we'd `meow' \emph{} if we ever needed to talk?''

``That's so, Tom. I reckon I forgot. What brings you out?''

``I got important news,'' he says, sitting cross-legged next to me.
``And it can't wait till morning. It's important.''

``Well, what is it?'' I says, but Tom shook his head.

``Not here,'' he says, all mysterious-like. ``Ain't safe here. Got to go
to church.''

``To \emph{church}?''

He nodded wisely, like a little owl. ``Absolutely vital.''

``Well can't you at least give me a hint before I go slinking around in
the rain like a fool?''

``Fine then, know this: it concerns your \emph{very} \emph{survival.''}

I shivered a little when he said that.

So we made a run for it, Tom with his roundabout (which was his little
jacket) above his head, me with my big overcoat over myself and Tom's
lantern. We made it to the church without any goons or pirates or
whatever was after us, catching us. Tom tugged on the heavy door and it
opened. I'm not sure, but I don't think churches is allowed to lock
their doors.

Inside it was still as a tomb. The thunderstorm echoed and rumbled, and
the rain was like a million cold fingers drumming on the roof. It
smelled like wood and polish and dust in there, like them places always
do.

I don't habituate churches as a general rule, not on Sundays but not on
regular days, neither, because as near as I can reckon, if the heavenly
father knowed I was in his personal space, he would shoot me in the
head, me being such an immoral, unrepentant varmint. I reckoned he'd see
it as a blazing bit of cheek.

I hoped he might look the other way tonight.

I collapsed into one of them long benches, but Tom motioned me further
in. We crouched behind the little pedestal thing at the front.

``What is it?'' I says, because by now I was pretty grouchy.

``You wouldn't take that tone, Mister Huckleberry Finn, if you knowed
why I brought you here,'' he says in a superior voice. ``Have you ever
heard of\ldots{}\emph{the blood-takers?''}

My mouth dropped open. ``The \emph{what?''}

``All right, Huck, don't feel too bad. I reckon it's understandable,
your being so uneducated and so kind of low-down and ignorant. If you
went to school and read books, I reckon you wouldn't have no excuse for
not knowing. But we'll overlook it, because you warn't brought up well,
even though it's \emph{critically hazardous} that you don't know.''

``Don't know what? What are they?'' I rubbed up and down my arms, which
was goosebumps all over.

``They're called \emph{vampyres} by intellectuals, I mean those who know
all about them and study them,'' says Tom. ``That's the scientific name.
They look like regular people, except they're pale as the belly of a
fish, and they have teeth like a mountain lion, and they \emph{drink
human blood.''}

``No fooling, now?'' I grabbed him. ``You're not pulling my leg, are
you, Tom?''

``Would I kid you about a thing like this?'' He pushed me off. ``If I
was fooling, don't you think I\textquotesingle d a waited and told you
in broad daylight---''

``I wish you had,'' I says---

``---In the middle of town? No, sir. Why do you think we came to the
church at this unholy hour?''

``Well\ldots why did we?''

``Because church's the only safe place to talk about blood-takers. Any
other place, it'd draw them like a swarm of locusts if you so much as
\emph{mentioned} `blood-taker.' But they can't come into a church,
see?''

By this point I was shaking like a leaf. ``I'm glad there ain't any
blood-takers in St. Petersburg.''

``Now Hucky, I know you're simple, but try to think. Why would I be
telling you this, if we warn't in \emph{imminent danger}?''

I gripped his wet collar. ``Tom Sawyer, you mean there\textquotesingle s
\emph{really} and \emph{truly} a blood-taker in this town?''

``He arrived today,'' Tom says dramatically. ``And his name is\ldots{}
\emph{Jackson Morley!''}

I stared at him blankly. ``Who?'' Because I knowed everybody in town and
I ain't heard of no such person.

Tom sighed. ``The new constable. He arrived today, on a Sunday---''

``That can't be a bad sign,'' I says---

``---And he \emph{didn't go to church}!'' says Tom. ``Moreover, he's
pale as a flounder and it was cloudy all day.''

``What's clouds got to do with it?''

``Blood-takers can't go out in daylight. It shrivels them like a slug in
a salt cellar.''

I was quiet while I digested this. ``So what are we going to do?''

``Take precautions. Don't leave home without your garlic and your silver
cross and your holy water.''

I gulped. ``I ain't got none of them things.''

``Don't interrupt. We also need some little bits--- beads or sand---to
scatter if we see him coming. Blood-takers are suckers for little bits.
They\textquotesingle re obliged to stop and count them, and while they
count them you run like the dickens for the next county. And the only
way to kill them is to ram a stake through their hearts.''

I looked around. ``Tom, I don't know about that other truck, but we're
in a church now, so I bet we could get some holy water.''

``Good idea. They probably keep it up here.'' Tom searched the pedestal,
but he didn't find none. We both set to looking and ransacked the place
pretty good, and eventually I found it in the back, by the door. It was
in a big barrel with a dipper, for drinking during the long, hot
services, where the preacher really makes you roast. It was the only
water we could find, so it had to be the holy water.

We didn't have nothing to put it in, so we just dumped a little on our
heads. I didn't mind, being already soaked from the rain anyhow.

``What about that other truck you mentioned?'' I says once we done it.

``Follow me.'' And out Tom went into the rain.

We ran through town, cut down an alley, and ended up behind a house with
a picket fence. Tom hopped the fence like a deer, and I followed in two
licks.

``Where's this?'' I says in a whisper. I warn't sure, but I reckoned the
owner wouldn't fancy seeing Huckleberry Finn outside his house at
midnight.

``This is the house of a great Smiter,'' says Tom, stealing up to a
window.

``What in the nation is a smiter?''

``A smiter is a person who smites blood-takers,'' says Tom. ``Honestly,
if I knowed as much as you I sure wouldn't let on.''

``I just didn't know there was any smiters in our town,'' I says. ``How
do we get in?''

``We got to break in, Hucky. Give me a hand.''

I saw he was trying to jimmy the window latch, and the way he was going
at it wouldn't never work. I showed him how it was done, and Tom blew
out the lantern, and in we slipped.

It was dark in the house. It smelled nice, like soot and bread and dried
flowers.

Tom went into the pantry and rummaged around. He came out with two long
strings of garlic, which he draped over our shoulders. Then he went to a
drawer and started going through it.

``Help me find the silver,'' he says, all soft. So I slid a drawer open
and started looking.

It took a couple tries, but we found the silver, and he put us each a
fork in our pockets, and said we could bend the tines out to make them
crosses. I liked this idea a lot, and said so.

Next he wanted to find some beads or sand, so he started into the rest
of the house. I was nervous because we'd been in there so long. I was
thinking about that stake through the heart business. I figured that was
a pretty efficient way to kill a human person, too.

``Tom---'' I says, when we was blinded by a sudden light.

``Thomas Sawyer!'' It was a woman's voice. ``What in the name of all
that's holy---''

I blinked like a mole. It was Tom's Aunt Polly, and beside her was Mary,
Tom's older cousin, and little Sid, his half-brother, who was wearing
nothing but a tow-linen shirt that reached his ankles.

``I tol' you, Auntie,'' says Sid, clinging to Polly\textquotesingle s
skirts, ``I tol' you I heard something.''

Tom couldn't say nothing, but the look he gave Sid was enough.

``Creeping around in the dark at Lord only knows what time, making
puddles and wearing vegetables!'' Aunt Polly says. ``Get back in bed
this instant, Tom Sawyer, because if I gave you the licking you deserved
right now there wouldn't be nothing left of you. And as for you---'' She
glanced at me and gave a start, like I'd slapped her.

``Huckleberry Finn. Get. Out,'' she says in this low, terrible voice to
curdle your blood. I was already halfway to the door before she
finished, and I\textquotesingle d a ``got out'' if I'd a been dead.

I scrabbled at the front door, leaped down the porch steps and flew away
into the night. Luckily she didn't make me turn out my pockets, so I got
away with the fork. And you can bet the first thing I done when I got
back to my old hogshead was bend the tines into a cross.

~

\mychapter{Chapter 2}\label{chapter-2}

Next morning the sun waked me hot on my face, which meant I slept in. I
snaked myself out of my barrel and stood in the cool morning air,
reaching my arms, having myself a good stretch. I was going to march
straight down to the river to see if I catched anything on my lines,
when I remembered what happened last night, and a little zing of
lightning shot through me. Forgetting the fish, I made like a deer and
high-tailed it into town.

I borrowed a couple of corn dodgers that was cooling on a windowsill and
made for the general store. On the way I stopped by the town pump, where
everybody and their sister has to get water, to hear what the locals was
jawing about.

I sort of wiggled my way in amongst the water-pump crowd and snatched a
drink while I listened.

It warn't nothing worth hearing. Just some pie-eater trampled Malcolm
Rogers's field and stole some watermelons, which was a fool thing to do,
because I know for fact them's the bitterest things this side of the
Mississippi and not worth the trouble of stealing. Willie Mufferson,
Master Frilly-frills himself, won some schoolhouse spelling bee, bully
for him. Some girl up in Goshen got snatched by roughs and held for
money, and the whole town chipped in to pay out. Ol' Mother Hopkins got
turned out of her house by her landlord, but no one cared because she
was a witch. That part was true, she was a witch, but I didn't think she
ought to be turned out of her house. Even witches got to live.

Just a lot of talky-talk. Nothing about no blood-takers.

I moseyed down to the general store and sat on the post fence. It was a
good position, because I could see the whole street, and it was in the
shade. I found an old corncob and whittled it into a pipe while I
watched people.

I saw ancient Major Ward make his way down the street to the Temperance
Inn, where he would drink nothing stronger than soda and play cards. He
was limping, and it was his left leg, not his right---which he'd injured
in the war---so I reckoned he fell down his porch and got his knee
twisted like he done last month.

I saw the mayor's wife go into the general store. She looked tired, and
I reckoned it was on account of her daughter was always getting stomach
sick and so was probably sick again last night. She was likely getting
some kind of syrup for her.

I saw two clerks walking along side by side, looking in shop windows and
preening. I felt sorry for them. I'd seen them at their work, sitting
inside all day, six days a week, with ink on their sleeves, getting
yelled at by their boss. That was no kind of life.

And I saw a lot of other people, too.

About an hour later, who should mosey up the street but Mister Thomas
Sawyer himself, hands in pockets, boater at a jaunty angle, smug as a
cat with cream. He'd slicked his curls down with water, because he hates
his curls because he thinks they're pretty, but they was already drying
and springing back up.

``\,'Lo, Huck,'' says Tom.

``\,'Lo, yourself,'' I says. ``Why ain't you in school?''

``I\textquotesingle m sick,'' Tom says. But I knowed him, and reckoned
he was so tuckered from running around last night that he decided to
play hooky.

``What was the big idea, anyways?'' I says. ``Why did we go jimmying the
window to your \emph{own blame house} at midnight to hook your own
dad-blamed things?''

``Because that's the way it's done, Hucky,'' he says, hopping onto the
fence beside me. ``When you need relics to fight blood-takers, you got to
break in and steal them. You can't just take them out of your own silly
pantry. It cheapens the stuff. And I can tell you, it was a bother to
lock that window before I went to meet you. That window is never locked,
and the latch was rusted open.''

I leaned my head back and gave him this look, with half-shut eyes. ``And
I suppose the blood-taker was a lot of rubbage as well?''

``Hush up, Huck, cause here he comes. Wearing a hat, of course, because
of the sun.''

I whipped around so fast I nearly fell off the fence. I spied a tall man
in a uniform, someone I never seen before, walking up the street from
the courthouse. He had short dark hair and was clean-shaven. He saw us
gaping and came over.

``Good morning, boys,'' he says. Tom touched his hat. I tried to see the
fellow's teeth, to see if they was like a mountain lion, but I
couldn\textquotesingle t tell. He was a mite pale, but my Pap is paler,
and he's \emph{} as human being as they come.

``Morning, Constable,'' says Tom.

``I remember meeting you yesterday,'' he says. ``Polly's lad, right? I
just found out the church was broken into last night. Apparently some
roughs had a good time. The whole place is overturned.''

``Shocking,'' says Tom.

``And what are two upstanding lads doing out of school on a Monday
morning?''

``Sore throat,'' says Tom. He rubbed his neck tragically.

``And you're out of bed? Does your Aunt Polly know?''

``---of my brother Sidney,'' Tom says hastily. ``I was sent to fetch
lozenges.''

``Nothing like a bit of sugar for a sore throat.'' The constable eyed me,
sitting there dumb as a bump on a log. ``Well, run along then, Thomas,
get your lozenges.''

Now Tom had to go in. Jingling his pocket anxiously, he did. And at once
I realized that left me alone with the new constable, who may or may not
have been a blood-taker. I scrambled off the fence and followed Tom into
the store.

It was full of the usual truck, you know; rusty cans full of nails,
crackers and apples in barrels, cans and jars on shelves, big rolls of
calicos and linens. Tom went straight to the counter to look at the jars
of candy. I loitered near the door so I could peek out the window and
keep an eye on the blood-taker.

He was talking with Mrs. Walters, who had just walked up. They lowered
their voices, but anyone could a heard them if he pressed his ear
against the door.

``Who was that young man?'' says the blood-taker.

``That was Tom Sawyer,'' the lady says. ``You met his family
yesterday---''

``Yes, fine lad,'' the constable interrupted. ``I meant the other one,
the one he was with.''

``I didn't see another,'' Mrs. Walters says.

``Shifty-looking lad, a head taller than Tom, skinny as a rail, with
unkempt hair to his collar. And his clothes were all rags and tatters,
about a hundred sizes too large---''

``That's enough, Constable,'' the lady says, laughing. ``No question who
you mean. You must have seen Huckleberry Finn, son of the town
drunkard.''

``Finn?'' says the constable thoughtfully.

``Yes, sir. The boy's a layabout just like his father. He lives in
barrels or on doorsteps, won't go to school, hasn't done a day's work in
his life. The town mothers won't let their sons play with him, he's so
vulgar and trashy.''

``Finn\ldots I\textquotesingle m familiar with the name,'' the new
constable says.

Well, \emph{that} warn't good. I could just imagine what he'd heard
about Pap. I wanted to get out of there, less time lost the better. But
they was standing right outside the door. I glanced at Tom, paying for a
sack of assorted medicines for Sid, all in bright colors and covered in
sugar.

I eased out the door, trying to kind of slink along and pretend I warn't
there, but of course they noticed me. The constable waved me over.

``Come here, Finn. I'd like to talk to your father. Where might I find
him?''

I didn't know what to say. It was my policy to never tell the truth if I
was in a tight spot, but Mrs. Walters was there, and she'd call me out
if I lied.

``I ain't seen Pap in a while,'' I admitted at last.

``You mean you're on your own?'' says the constable. ``A homeless
juvenile on the streets? That won't do. That won't do at all.''

I knowed I'd regret telling the truth. Sweat sprang to my skin. Luckily
Tom came out of the store right then, and when he did I seized the
distraction and took to my heels and lit out.

I knowed that conversation didn't go well, but I didn't know what to do.
So I went through the woods down to the crick and checked my lines, and
got a fish off one, and cooked it up and eat it.

As I eat, Tom came tearing into the clearing like he was chased by a
bull.

``Huck Finn, you done it now,'' he says. He collapsed next to my fire
and tried to catch his breath.

``What have I done?''

``You only went and told the \emph{constable} that you was
\emph{homeless} and \emph{without a father.} I tell you, if I was as
pinheaded as you, I'd jump into Big Muddy and drownd myself, save
everybody a load of grief.''

``Look, even the best fellow can tell an occasional truth, if the mood
suits him.''

``Well you sure picked a dumb time to start.''

``What can it hurt? All I said was I didn't know where Pap was.''

``You just showed how much of an ignoramus you are. Hain't you never
heard of a poorhouse?''

I allowed that I hadn't.

``It's where they send folks that ain't got a bean, nor roof nor living.
Children don't count usually, cause they got parents to take care of
them. But you ain't got none of those.''

``What is a poorhouse?''

``It's a big stone building without any windows. And you work all day,
crushing rock or grinding up bones. Just the same thing, over and over.
And you sleep on an iron cot with a hundred other people, and rats run
all over you, and chew off your toes while you sleep. And you can't
never leave.''

``Why can't I?''

``Because you owe rent. And they pay you for your work, but not enough
to cover your rent, and that means you gotta keep working and working
and you never catch up.''

``Maybe I'd skip out in the night.''

``You do and they'd hunt you with dogs and kill you.''

I felt pretty ill. I throwed my fish bones in the fire and asked Tom
what I should do.

``Make yourself scarce. Find a spot to hang your hat that's not in town,
because if that constable finds you behind the slaughter-house he'll
haul you off before you can say Jack Robinson.''

I stood up, ready to make tracks.

``Oh, don't go yet,'' says Tom, digging around in his sweets bag. ``They
ain't calling out the dogs \emph{yet}. Sit down and help me test Siddy's
medicine. We don't want to give him nothing he won't like.''

I sat. ``What you got?''

``Here's some of those ones that tastes like chalk. What'll you give me
for `em?''

I ransacked my pockets. ``A blue ticket?''

It was something they gave out at Sunday school for learning lines.

Tom was scornful. ``A blue ticket is fire kindling. What else you got?''

``A pebble I found in a fish gut.''

I knew I had him with that one. When I let him hold it, his eyes gleamed
with desire.

``All them chalky ones for the pebble,'' I says.

``Add the blue ticket, and you got a deal,'' says Tom. So we shook.

I crunched my sweets while Tom sucked on a horehound drop and rolled the
pebble between his fingers.

``From a fish's stomach,'' he says dreamily. ``Probably magic. Usually
are.'' After a while he asked me where I got the blue ticket. He knowed
I never earned it.

``Got it off a boy for a frog leg.''

``What'd he want with a frog leg?''

``Make a charm. Wear it round your neck, keep off the evil eye.''

``Never heard that. Why didn't he just toss salt over his shoulder?''

``Frog leg's better.''

Tom was quiet, absorbing this. A little later he says, ``So what did you
think of the blood-taker?''

Truth be told, I forgot the constable was a blood-taker. I thought he
was plenty scary enough as a constable.

``You left before you saw me spill the sugar,'' Tom says. ``Remember how
blood-takers can't help counting tiny things? I dumped the sugar from my
bag all over his shoes, then ran while he was distracted. It worked---he
never looked up. Say, why do you smell so ornery?''

I thought for a minute, then dragged out a dozen cloves of garlic, from
my hat band and pant cuffs and sleeves and pockets. Tom fell back in the
dirt, howling with laughter.

I wondered where I'd live now that I had to retire from my cushy life in
town. I knowed of a couple spots along the river that might serve, but
none as tidy as a doorstep nor as cozy as a hogshead.

``You want to help me pick a new place to live?'' I says, but Tom shook
his head.

``You can't tell \emph{nobody} where your secret hideout is, including
me. Then if they torture me, I can't tell them nothing.''

``Lordy, will they \emph{torture} you, Tom?''

``Probably. Maybe. Who knows? You got to think of everything.''

``Can't I never go back to town?''

``Well, sure, you can go back sometimes, just don't sleep there. That'd
be \emph{asking} them to scoop you.''

I breathed a sigh. I didn't think I could go cold-turkey from all the
modern conveniences all at once. I was too soft.

Back when I lived with Pap, I could a done it. We lived by ourselves,
just the two of us, for months at a time. Pap would run into town to get
another jug a whiskey, or a side of bacon or some cornmeal, but he'd
keep me locked up at home so I couldn't run away, and I'd go months
without seeing anyone but Pap. We had a high time, too. Sleeping all
hours, catching as many fish as we wanted.

I suddenly remembered this mill across the river from me and Pap's old
place. As I recollected, it had been in pretty poor repair, but now I
couldn't afford to be a snob.

``Say Huck, you ever play musketeer?'' Tom interrupted my thoughts.

``I don't reckon. What's that?''

``Well, you're a swordsman in a gang, and you're sworn to protect the
king from highwaymen and robbers and evil cardinals. Wanna play?''

I was agreed. So we played musketeer. And when school let out, Joe
Harper, who was Tom's best mate, came and played another musketeer, so
we was three, which Tom said was the correct number. And Tom played
Dartanyen, and I was Aramiss, and Joe was Poor Toss. And Tom said there
was actually a fourth, but he disremembered his name and anyways there
was really only supposed to be three.

Tom and Joe had to go home for supper after a while, so I sloped away
into the woods and found the river, and followed it along, past
Jackson's island, to where I remembered the mill.

It was even worse than when me and Pap lived nearby. I looked for me and
Pap's old cabin, but then I remembered it burnt down when Pap set his
beard on fire, the time he was in one of his raging drunks and threw a
lantern at me, and the grease spilled as he swung it over his head.

The mill was easy to get into because the door was hove off to one side,
ripped off its leather hinges. The inside was filthier than a sty. About
a hundred squatters must have stayed there over the years, and they left
a ton of bottles and rubbage behind and they burnt a hole through the
floor, which was derned careless of them.

So I set to fixing up my new home.

\mychapter{Chapter 3}\label{chapter-3}

It was a little rough living in that mill at first, but pretty soon I
got where I was used to it, and then I decided it was even better than a
doorstep. I set a trot line on the river, which is a long line across
the water with a lot of little fishing lines dangling from it.

I wallowed out a spot for my fire, where I could cook, and sit, and
think lonesome thoughts at night, and feel like I was the only fellow
alive on earth.

That suited me fine for a time, but pretty soon I felt I would die of
lonesomeness, so I decided to go into town. I buried my fire and rested
my door against its buckled frame. Just in case anyone was looking for
me, which I didn't reckon they was. Seemed like too much trouble to go
out and \emph{find} me.

Well, as it turns out, trouble did find me---at the bean feed.

A bean feed is this annual powwow the church holds to raise money. They
bring in gunnysacks of beans, all different types. They cook them in
these giant witches' cauldrons on fires. The pots was so big they had to
stir them with canoe paddles. There was about a million people in their
Sunday frills, looking like stuffed pheasants.

I moseyed around, trying all the beans and keeping an eye out for the
constable, when Tom Sawyer sidled up. He was dressed in a suit, real
starchy, buttoned to the chin and looking hot and uncomfortable. His
little boater hat was straight, his ribbon necktie a tight bow, his
curls combed around his ears. He was holding a china bowl with blue
flowers on it. It was bring-your-own-bowl. Mine was the lid off a
cracker tin.

``Hi-ya,'' Tom says.

``Hi-ya yourself,'' I says. I was eying a long table with a lace cloth
on it, which groaned under the weight of corn dodgers and pies and
biscuits and buttermilk and corn on the cob and watermelon wedges and
gingerbread. That's the truck you paid for. The beans was free.

And get this: there warn't nothing but a couple sweet little church
ladies guarding the table.

``Go over there and talk to Mrs. Riverson,'' I says.

Tom didn't ask no questions, bless his little heart. ``Morning, Mrs.
Riverson. What a turn-out, I reckon you must be real pleased.''

While they was jawing I idled around the table, innocent as a babe.
After they both agreed the turn out was excellent and the ladies had
done swell, Tom told her goodbye and I slinked away into the crowd.

Then while I ate biscuits and gingerbread and shook crumbs out of my
sleeves, Tom met up with me and led me behind a stand of trees where no
one would see us together. He ran his mouth all the while.

``I've been looking for you for days. Where you been? I knowed I'd find
you here. I knowed Huck Finn couldn't resist a bean feed.'' Then he
launched into a detailed account of some game of his and Joe Harper's. I
warn't overly attentive, being more interested in stuffing my pie hole.

Suddenly everybody at the feed stopped talking. Tom and I craned our
necks to see what was what. There was two strangers standing like
statues in the middle of the crowd, which had parted like the Dead Sea
around them. And never has the word \emph{stranger} been more useful, I
reckon.

The first fellow went on for miles---he was chest and shoulders taller
than anybody else. But he was skinny as a rake, and his long spine
curved like a shepherd's crook, like he was trying to ingratiate himself
shorter to see people right in the face. He had long straggly grey hair,
and a long hooked nose, and a long curved chin. He wore loose grey
robes, cinched around the waist with a wide fabric belt of a darker
grey, and he had pointed silk shoes on his feet. A bottle made of green
glass was tucked into his belt.

There was no way of knowing what his pal looked like, because his pal
was swaddled head to toe in dirty red fabric. Even his face was covered.
His eyes peeked out of a slit in the cloth, black and glittery.

Looking at the pair, a thrill ran up my spine into my scalp. I could
feel my hair stand on end. I traced a cross over my breast three times
and spat.

The Reverend approached them.

``Welcome,'' he says. ``Have some supper; we're raising money for new
pews in the church.''

``What a lovely, charitable, worthwhile cause,'' says the tall skinny
one, grinning and rubbing his hands. He made a funny kind of skip with
his feet. He and his red pal went off with the Reverend to talk.

I turned to Tom, and he was squirming with curiosity.

``We got to find out what they're saying,'' he says.

``How? They're right in front of the church. No way we could sneak close
enough without them seeing us.''

Tom looked around, and I saw him latch onto the twin oaks. They was
these great big trees, probably a thousand years old, rooted right in
front of the church. One was near our little stand of trees, and the
other stood overhead of the three men. Both had great spreading branches
that was thick as regular trees.

Tom looked at me, and I knowed exactly what he was thinking.

``I hoist you or you me?'' I says.

``You me,'' says Tom, and leaped for a low branch. I grabbed his feet
and pushed him up, and when he got settled I jumped and he took a
fistful of my shirt with one hand and my gallus in the other and hauled
me up.

We eased ourselves along the branch, to the furthest point that would
hold our weight, and then we was obliged to jump to the other oak. My
heart hammered in my throat the whole time, but I done it. The trees
shook hard, so we paused after we jumped, to make sure nobody noticed.

Then Tom and me picked neighboring branches and stretched ourselves
along them like cats. The bright green summer leaves was thick enough to
provide cover from below---we hoped.

We could hear what they was saying now. They was talking about the
church.

``See the bottle? That's how you tell,'' Tom says in a whisper.

``Tell what?''

``That the red one's a gin. That's a demon that lives in a bottle. The
tall, creepy one must be his master. Prob'bly a magi. They're the only
ones can properly control a gin.''

``What does a gin do?''

``All sorts of mischief. Horses with broken legs, fires, illnesses,
things that look like accidents.''

We listened to them gabbing. The tall, hunched one did the talking. The
gin didn't say nothing, but his head did this eerie bob now and then,
like a cobra scenting the breeze. \emph{That} was the varmint worried me
most, for some reason.

``We come from ancient Persia,'' says the magi man. His long arms jerked
skywards like they was attached to strings.

``What brought you to St. Petersburg?'' says the Reverend.

``It was a steamboat,'' the man says, and his feet performed a little
skippy jig. ``Steamboat steamboat steamboat! A lovely steamboat. We
sought family up north. But we hopped off,'' (he took a leap in the air,
his spindly knees rising like a grasshopper\textquotesingle s,) ``to see
your town, which is famed the world over for its beauty and its
charity.''

The Reverend turned pink with pleasure. Tom snorted.

I was sweating bullets trying to hear, and my dad-blamed hair kept
getting in my eyes and sticking to my face. I reached over and grabbed
one end of Tom's ribbon necktie and gave it a sharp yank, and pulled it
off his neck. He didn't seem to mind; in fact, he opened his collar and
rubbed his red throat.

I tied my hair behind my head, and for a moment enjoyed relief. Then I
saw a hornet, a big fat one, crawling up Tom's arm. I bit my knuckle and
pointed. Tom gently flicked it off. Then I saw one on my arm. And
another on the branch in front of me. I looked over at Tom, and to my
horror he was crawling with them. So was I. We was in a nest.

We stayed as calm as possible under the circumstances. We couldn't
scream bloody murder or we'd give up the jig, nor could we wriggle
backwards without crushing the nest and getting stung up.

Them hornets was pretty mellow. Tom catched my eye and mouthed,
`\emph{the smoke.'} I reckoned he was right. The smoke from the fires
was keeping the hornets drunk and lazy. Maybe if we stayed still they
would let us be.

``So you were seeking family, and happened to alight in St.
Petersburg?'' the Reverend says below.

``That's right,'' says the magi, placing the bony tips of his hand on
the Reverend's chest. ``Purely by happenstance. So you can imagine our
surprise when, lo and behold, we discovered the relative we sought lives
right here in town. It was the most blessed piece of luck.''

``Remarkable.''

``A lad by the name of Finn. Huckleberry Finn.''

I nearly fell out of the tree.

``We're distant relatives, and we came to America to tell him he's
inherited a fortune in jewels from his recently-deceased
great-great-uncle.''

Inherited jewels! I was rich! I turned to grin at Tom, and felt a
horrible prickling behind my head. I'd forgotten about the hornets. A
loud buzz came from behind my left ear.

``Blimey! Jewels, you say? Finn, you say? Well, I'm jiggered. Which
great uncle of his, did you say?'' says the Reverend.

``The Persian one,'' says the magi.

Things was getting hot for me in the tree. The prickling and tugging
behind my head got worse. I blew a drop a sweat off my nose and caught
Tom\textquotesingle s eye. `\emph{Hornet',} I mouthed, and made a tiny
frantic gesture. Tom waved me off; his attention was riveted below.

The only thing I could figure was that I had tied the blame thing
\emph{into my hair,} and now it was struggling to get out, and getting
madder and madder all the time. I bit my lip and tried to keep still.

Down below, the magi produced a folded piece of paper from his belt,
sealed with wax and all.

``This,'' he says, ``is the last will and testament of Finn's second
uncle, three times removed, who left the boy a fortune in emeralds,
diamonds, rubies, and sapphires. Anyone may see it's genuine.''

He held it up, and the Reverend examined the seal.

``May I read it?'' says the Reverend. The magi whipped it away and
tucked it back in his belt.

``Sorry! So sorry, dear, dear, reverend. Can\textquotesingle t allow
that. Must be opened in the proper legal manner, by the lawyer in charge
of the estate. If the seal is broken before that time, it would be void
and null. That\textquotesingle s to prevent any suspicion of forgery.''

The hornet in my hair stung me. It didn't feel like anything at first,
but then I nearly screamed. It felt like someone rammed a red hot needle
into my skull. Forget secrecy; I had to move. I yanked the ribbon out,
and my hair fell down around my face, and the hornet bumbled around, all
tangled up, then flew off. I screwed my eyes up and gritted my teeth,
and was so focused on not making a sound, I didn't hear nothing else
below.

Tom poked my shoulder. I shook my head. He grabbed my arm and squeezed
hard. I opened my eyes and looked down, and gasped. I had dropped the
ribbon, and it fluttered towards the men.

It fell at the feet of the gin. He stared at it for an eternity. And
then---and then---he looked up! I held my breath, eyes streaming,
trying not to move. You couldn't see his expression or nothing. Tom was
frozen, gripping my wrist like a vise. Neither of us twitched an
eyelash.

Finally, when I was sure the gin would say something and give up the
jig, he looked away. He picked up the ribbon and put it in his pocket.

I tried to pull myself together and listen again, and breathe, but the
talk was over. The Reverend assured the strangers he'd let them know if
he saw me, and they straggled off, and Tom and me dropped straight out
of the tree as fast as possible, covered in bark and leaves and trying
not to touch anything because we was all over hornets.

``Get them off quick---but gentle,'' says Tom, and for the next few
minutes we flicked bugs off each other as quick but as gentle as
possible. Even still, a number of them stung our fingertips.

``What a rum thing,'' Tom says, eyes blazing as he flicked. ``Persian
jewels. Did you know you have Persian relations, Huck?''

I shook my head. I went around behind him and flicked some off his back.

``Of course, stranger things have happened,'' says Tom. He flicked one
off his knee. ``It's not like we know a whole lot about your relations.
Wouldn't it be bully if you really was hair to a load of jewels?''

Tom ran on in this vein for a while, dreaming, until we finished our
business and was clean. Then he went to find his family.

I sucked on my red, stung-up fingertips and thought about the gin
looking up into the tree and picking up the ribbon, and I got a horrible
case of fan-tods. I decided to forget about the bean feed and the
Persian jewels, and ran for home as fast as I could. I reckon I went
three miles out of my way through the woods, down the river and back up,
to cover my tracks. Just in case.

\mychapter{Chapter 4}\label{chapter-4}

That folded sheet of paper, sealed with red wax, consumed my every
waking thought. I'd never heard of no Persians in the Finn line; far as
I knew, the farthest a relational of mine ever lived was St. Louis. And
I think he was incarcerated there.

Finally I decided there warn't no sense worrying myself to death over
it, there was only one thing for it. Barring actually \emph{speaking} to
the men, of course, I could at least get a looky at that will.

So I went to Tom Sawyer's house at midnight, and I stood in the garden
and yowled below his room like a cat, and when the window slid open, I
retreated and stood a ways back in the woods.

``Hi, Huck!'' He skipped up to me, swinging a tin lantern with its flame
turned low.

``Hi yourself, see how you like it. Listen, I got to know what's in that
will. Will you help?''

Tom clapped his hands with glee. ``I love you, Huckleberry Finn. I love
that you came to me for this. What's the plan?''

``Sneak in, steal it, read it, put it back.''

He stopped jumping. ``You call \emph{that} a plan?''

``What's wrong with it?''

``Oh, you poor varmint. You ain't got no head for plans. It's too
simple. It won't never work. We need a diversion, and a decoy will, and
a tincture of valerian to knock them out---''

I slapped my hand over his mouth. ``Tom Sawyer, shut yer trap. I ain't
got no patience for tingtures nor decoys. I just want to slip in all
private-like, take a peeky-peek, and get the blazes out. Now, this is my
expedition, and I reckon that makes me captain. That means you do it my
way, or you're out. Right? Good?''

``Good as wheat,'' Tom mumbled through my fingers. I removed my hand.

``Alrighty then.'' Off we set.

``Do you know where they're at?'' he says presently.

``Most every stranger comes through stays at the Temperance Inn,
right?'' I says.

Tom allowed it, so that's where we went. The Temperance Inn is this big
place that serves drinks that don't contain likker, and also has a bunch
a rooms upstairs for let. We looked up at the windows of the rooms, and
they was all dark.

``Which room?'' says Tom.

``I'm thinking.'' Truth was I didn't know how to cypher it out, there was
too many windows.

``Can I say something?'' says Tom after a moment of watching me sweat
and rub my noggin and get more and more worked up.

``What?''

``Suppose I know where they're at.''

I stared at him. ``How?''

``Suppose,'' says Tom, highly pleased, ``me and Joe Harper was playing
in the square after supper this evening, and suppose we happened to see
the two ruffians take a room, with our own eyes?''

``Did you, Tom? Did you really?''

Tom jumped and clicked his heels. ``Yes! Yes!''

``Well, which room is it?''

``It's the little crummy one around back, the one they let exclusively
to no-accounts who look like they ain't got a bean and will
prob\textquotesingle bly do a bunk. Come on.''

Sure enough, around back was a low, ugly little door. Tom covered the
lantern light.

``Hoist me up,'' says Tom, and I did, and he looked through the window.

``Asleep,'' he says, so we was a go.

The door warn't even latched. Tom gently pulled the loop of leather
strung through the door, and the door swung open. It was dark as pitch
in there.

``After you, Captain,'' Tom says, and saluted. So I slinked in.

I stepped delicately, on the edges of my feet. It was hot and stuffy,
and smelled like filth and sweat.

As my eyes grew accustomed to the dark, I saw the magi. He was curled on
a pallet like a long snake, snoring. But his snores was more like long,
nasal \emph{hmmmms}. The room was empty except for a tin basin and a
large wooden chest for storage. I didn't see the green bottle that the
magi carried in his belt. And though I raked the dark carefully with my
eyes, I couldn't see the gin, neither. That warn\textquotesingle t
comforting. If he had stepped outside to answer the call of the wild, he
would come back and catch us like cats in a coop.

I looked at Tom, and he motioned me to go on. So I crept over to the
magi. I remembered where he tucked the map, so I reached all slow into
his belt, and gingerly ran my fingers along it. I felt a crinkle, and
pulled something out. It was some paper, but it warn't the will, I could
tell that even in the dark. No red seal. So I put my fingers back and
oh-so-slowly eased along, feeling\ldots{}

Just as I felt the edge of something, I heard the most horrific roar,
like a grizzly or a mad dog. The magi jumped aright, and I jumped
aright, and for a second we stared flat into each other's faces.

``Run!'' says Tom from the door. I didn't need to be told twice---I
nearly runned him down. We heard another angry roar as we flew into the
night.

The Inn was right next to the woods that climb Cardiff Hill, so we made
for the trees. Behind us, I heard light, skippy-skip steps, like someone
doing a demented tap-dance. The sound gave me chills to the core. It was
the magi---he was following us.

``This way! Come on!'' says Tom, and I tried to follow the bouncing
light of his lantern, but I tripped and slid in a ditch. As I scrambled
up t'other side I heard something thud behind me. I risked a glance back
and the magi was crouched at the bottom, knees and elbows akimbo like a
giant toad, grinning at me.

I took off like a hare. I could hear Tom to my left, so I went that way.
I tripped on a root and rolled. Before I could get to my feet, the magi
loomed in front of me. His spine was bent double as he peered into my
face, grinning. I could see the shine of his teeth. His long arms rose
to grab me.

I crawfished backwards so fast I kicked up leaves. I scrambled to my
feet and lit out to the right.

I could hear the magi\textquotesingle s leaping feet behind me. I
couldn't see nor hear Tom. I ran for my life.

Eventually the magi's sounds grew fainter, and I saw a tiny light up
ahead. I thought it was Tom's lantern, so I made for it. Then it was
gone, but I saw another light to my right. I ran towards it, then it
winked out, and another lit up to the left, which was back the way the
magi was, and that was when I realized that these lights was
jack-o'-lanterns and not real at all, and I was completely lost.

``Got you!'' sings the magi, grabbing the back of my shirt. I screamed
like a banshee. Without thinking, I twisted and struck out blindly with
my fist. I heard a small ``\emph{ack''} as my hand connected with
something, possibly his ear. Then my shirt tore and I floundered away.

I saw a brighter light up ahead, and this time I was sure it was Tom,
come back for me. So I flew towards it, and suddenly my stomach was in
my mouth and my feet was in the air, and I was falling straight down,
down, into the earth.

I landed with a splash and a deafening echo of falling stones. I
couldn't see my hand in front of my face, but I could feel I was in a
couple feet of smelly mud and water, and I couldn't stretch my legs, nor
even budge an inch. The sound of the water splashing echoed loudly in my
head. I had found a well.

I was truly up a stump this time. The walls was slimy, and when I tried
to grab some stones and lift myself up they tumbled around my head and I
was afeared they would knock me out cold and then I'd drownd.

There warn't nothing to do but wait, and hope Tom Sawyer returned and
found me. I couldn't hear nothing from above, just my own breathing and
the water splashing. I dasn't call for help, in case the magi was up
there.

So I waited. And you've never been through a longer night than mine at
the bottom of that well. I felt like coffee grounds at the bottom of a
week-old pot. My legs hurt like the dickens, I was cold all over, and it
smelled. Even worse, I imagined slithering beasts and things with a
hundred legs all around me, in the water and on my clothes. I rubbed my
hands all over to try to get them off. I trembled like a leaf and I
don't know if there was really anything down there, but I swear I could
\emph{hear} them in the water. Maybe they was imps.

I tried to calm down before I completely lost my nut. I traced a cross
on my chest three times and spat. I counted my heartbeats and tried to
pretend I was back in my mill, and the water sounds was the river
outside.

Picturing the river actually did me some good. I imagined it so clear in
my head, like I was looking at it. I felt the old Mississippi ease down
my shoulders, through my arms to my fingers, like a warm blanket. My
muscles relaxed, and I stopped shaking. My heart slowed.

Time crept.

Eventually I saw the circle of grey sky above me, and I dared to call.

``Helloooo,'' I says, and it echoed so it nearly split my head. ``Hello,
hello, hello,'' I says, and \emph{hellohellohellohellohello} said my
echo, all tinny.

My hope was pretty much spent when I heard a far-off voice like a ghost
from the great beyond.

``Huck Finn? That you?''

``Help!'' I says. A rope hit me on the head, and I grabbed it, and they
pulled, and my legs unfolded with a pain like fire, and I was hauled up.

I flung my arms over the side of the well, which was just a hole in the
ground edged by the remains of a rotted wood cover, and my savior
grabbed the torn pieces of my shirt and hauled on me, and I was out.

I lay gasping on the ground, chest heaving, hair plastered to my face,
shaking all over. My legs was so numb I couldn't feel them, except in
sort of an achy way.

``You lay there a minute and calm down, and don't sit up till you're
good and ready,'' says a voice, and my eyes sought the speaker against
the pale morning sky. It was Uncle Jake, who worked on Ben Rogers's farm
(for no pay). Uncle Jake was older'n my pap. I did odd jobs for him
sometimes, split wood or hauled water, and every so often he fed me from
his own supper pail. It must a been the Rogers well I found, and their
farm must a been close by.

``Uncle Jake,'' I says, and then to my great shame I buried my face in
his lap and cried.

He patted my back, saying, ``There, there, that's all right, you safe
now,'' and other soothering things.

``I better tell them you was found,'' he says after a bit. ``I was sent
to find you.''

``How---why---?''

``That scamp Tom Sawyer come tearing up to the farmhouse last night,
hollering about being chased through the woods by a demon or some like,
and said you was still out here and maybe you was dead. So Rogers sent
me out, and I heard your teeny little itty-bitty voice just when I was
fixing to give it up.''

I tried to stand, but my legs was too weak from being in an unnatural
cramp all night, they couldn't bear the weight. So Uncle Jake turned and
I put my arms around his neck and hanged like a monkey, and he carried
me back to the house and set me in the kitchen, and Tom was there at the
table looking exhilarated.

``Tell me everything! Did he get you?''

I told him everything, and he was envious up to \emph{here} that I got
stuck in a well.

``The adventure of it,'' he says dreamily. ``Some fellows have all the
luck.''

Then he sighed. ``I reckon we failed, though. We didn't get the will.''

I reached into my breast pocket and pulled out the slightly damp will,
still sealed with red wax, and Tom snatched it and held it to the light
and felt of it and tested the wax with his nail.

``Well done, Huckleberry!''

``Now what? How do we read it?''

``What do you mean?''

``The magi said if we opened it, it was void and null.''

``Why, Huck, all you got to do is heat a knife and slide it under the
seal. Read the will, melt a little more wax, stick it back down again.
It ain't but old pie.''

So we done it. The seal broke a little, but Tom said that could be fixed
when we stuck it back.

We unfolded the paper, and Tom read it aloud.

``To whom it may concern: THIS MONTH only, for a limited engagement, the
FABULOUS Fabrizio, Witch Doctor of the Orient, is here with his
MARVELOUS Array of Mystic Potions and Remedies. FOUR DAYS ONLY. Bring
him your Sick and your Injured! He cures all ails---''

He trailed off and we stared at each other.

``What does it mean, Tom?'' I says.

``It means, them gents ain't here to give you no diamonds,'' he says.

\mychapter{Chapter 5}\label{chapter-5}

``The question is,'' says Tom, ``if they ain't got no jewels and the
`will' is bunk, how come they're searching for you?''

``It ain't a good sign,'' I says. ``If those men know me, it's because
they know Pap, and if they know Pap, then they got a bone to pick with
Pap, sure as you're born. And that could make things mighty hot for
\emph{me}.''

``Could be,'' Tom mused. ``Could be your pap crossed `em at cards, or in
some bad deal, and now they want to make you hurt. That takes.''

``So what do I do?''

``I reckon the only thing to do is avoid being alone with them at all
costs. If you run into them, make sure it's in a crowded place with the
constable and the mayor and everybody all around, so they can't do
nothing. And keep sleeping nights in your secret place, and make sure
you ain't \emph{never} followed back there.''

I nodded fervently.

``Reckon the magi got a good look at you last night?''

I thought back and shuddered. ``We looked right into each other's
faces.''

Tom looked sorry. ``Well, it was dark in that room, and even if he did
see you, he prob'bly wouldn't know you for Huck Finn, if you get my
meaning.''

I nodded glumly.

Then Tom left to get home before his aunty waked up, and I waited around
for Uncle Jake.

When he returned he gave me some brown bread and coffee, and asked me to
kindly explain what business I had in Rogers's well.

``Well, Tom Sawyer and I was out in the woods, and we was running about
playing headless horsemen and chasing jack-o'-lanterns, and I reckon we
got too close to the farm and didn't see the well in the dark, and I
stood on the rot cover and it broke, and Tom was off a ways so he didn't
see where I went.''

``That should teach you boys to go tearing round in woods you don't
know, in the dark, like headless fools,'' says Uncle Jake. ``How'd you
rip your shirt?''

``Catched it on the edge of the well cover when I bust through,'' I
says.

So Uncle Jake threaded a needle and fixed my shirt up real tidy, so you
couldn't even hardly see where it'd been torn. I reckon I was half in
love with him for that. When a fellow ain't got but one shirt, it takes
on a certain significance to him.

I watched him work. Folks liked when Uncle Jake looked at them and
talked to them, because he was real kind. He had curly white hair and
lots of wrinkles, and freckles from the sun. I was all over freckles,
too.

Finally the shirt was done and I put it on, and I wanted to sit in the
kitchen drinking coffee with him all day.

``Uncle Jake, how about you and I go down to the crick later and look
for frogs? And you could teach me how to tickle fish so they let you
catch `em and yank `em out of the water? And after that we could go
swimming, and---''

``I got to get back to work,'' he says.

``Oh. Well, I could help you. What do you got to do? I can chop wood for
the wood-rake, or---''

He held up a hand to stop me. His eyes twinkled. ``You had a long night.
Why don't you go on home and rest?''

``I feel all right now. I could fetch water, or if you're working
outside I could walk around with you and---''

He chuckled, a warm sound deep from his chest. ``Baby boy, you're sweet,
but with all due respect, get lost.'' He showed me the door. ``Run along
and let me work.''

I left. And I never saw Rogers nor any of the rest of the household. I
wondered if they even got out of bed, or just sent Uncle Jake to handle
it.

I went back to my mill and got a fishing line and sat on the bank and
dangled it in the water and sort of dreamed the day away. I watched the
big steamboats going up and down the river, smoking like chimbleys and
looking like wedding cakes, gleaming white and gold.

The well and the chase through the woods felt like a nightmare, and I
needed to loosen up after that. I stretched my toes and fingers and
closed my eyes and breathed deep of the smell of the river, and relaxed.
It was better than medicine. I reckon there ain't a lot a doctor knows
that the river doesn't know better.

Late afternoon I heard a snuffling sound in the woods behind me, so I
dropped my line quick as a flash and wallowed into the brush to hide.
Peeking out, I saw a boy a few years younger than me. He was crying and
kicking rocks along the path, all ornery.

``I know you're in there,'' he says, and I nearly jumped out of my skin.
``I see you.''

I poked my head out. ``Who're you?''

``Ain't none of your business. Go stick your nose in a hornet's nest.''

I pulled my head in and settled myself in the dirt. ``Fine. \emph{I}
don't care. And I don't care to come out, neither. I like it in here;
I'm perfectly comfortable. If you knowed about this old bush, I reckon
you wouldn't be so fast to tell me to get out. I reckon you'd want to
get in yourself.''

``Oh, it's just an old bush,'' says the boy. ``You're a catfish.''

``You're so right,'' I says. ``It is an old bush, absolutely nothing to
see, nothing that would interest you leastways, and anyways you ain't
allowed in even if you begged.''

``Am I begging? I don't see nothing so marvelous about it. What's in
there? I don't care.''

I put my hands behind my head and smiled up at the leaves and the
clouds.

``Is it really as nice as all that?'' he says after a second.

``You'll never know, because you ain't allowed in.''

``Oh, move over, we could both fit.''

``Ain't you got ears? I \emph{said} no.''

He shoved his hands into the bush. ``I reckon there ain't much you can
do to stop me.'' He pushed in and lay in the dirt beside me. I didn't
look at him.

``Well, I'm in,'' he says. He twisted and looked about with interest.

``What'd you say your name was?'' I says.

``Samuel Clemens. You?''

``Huckleberry Finn.''

We was quiet and just looked around at the inside of the bush.

``Is this a dream?'' Sam says. ``It feels like a dream. Everything looks
different; stranger, more colorful and lovely; even the river.''

``What'd'ya mean, a dream? You're nutty. How could it be a dream?''

But he was muttering to himself, happy as a loon, and didn't hear. ``Is
it my dream? Or is it your dream? Or whose dream is it?''

I rolled my eyes.

``If it's my dream, then you're not real,'' he says.

``I reckon I'd know if I warn't real,'' I says. ``So it must be
\emph{my} dream.''

``Then I'm not real,'' says Sam. ``And my life's not real, and nothing
that's happened to me is real. That would be nice, in a way.''

``What was you crying about, anyhow?'' I says.

``I could live in a bush someday,'' Sam says. ``It's peaceful.''

``Yeah, it's fine,'' I says impatiently. ``Listen, are you---''

``Look!'' Sam sat up like he was electrocuted. He gazed with adoration
at a big steamboat chugging up the river.

``Don't you wish you could live on a steamboat?'' Sam says. ``Look at
that beaut. A floating palace, a gilded cloud. A new riverbank every
day, new sights, not a care in the world\ldots''

``You ever been on a steamboat?'' I says.

``Yes. You?''

``No. Where you from? Ain't seen you before.''

``I ran away. Home is wretched.''

``Yeah? How so?''

``I'm always getting bossed about, no one cares for my opinion, I can't
have a horse of my own. They're shipping me out to apprentice to a man I
don\textquotesingle t even know, but I'm sure he's horrid. M-my---my
f-father---and mother don't love me. Anyways, they're always licking me
for things I didn't do, and we can't afford nice clothes, and all the
other boys make fun of me and call me names.''

``Tell me about your father,'' I says.

His voice catched in his throat like he was strangled.

``Nothing to tell. He's a judge.''

``Uh-huh. And?''

``And---and---he died. This morning.'' His voice was small.

A-ha. ``Sorry.''

``And my brothers and sister, too.''

Good grief. ``How'd it happen?''

``It warn't all in one go. My brother Pleasant died before I was born.
Then Margaret, then Benjamin, when I was real little. Then Daddy fell
ill two weeks ago, and this morning he\ldots he\ldots''

``What from?''

``Pneumonia.''

``So you're all alone?'' I wondered if I should invite him to live in
the mill with me. We could be brothers, and only look out for each
other, and when we sat by the fire at night we could be lonesome
together.

``Oh, no. I still have two brothers and a sister. And Mam.''

``Oh.''

We watched another steamboat glide past.

``Someday I'm going to be a pilot on a steamboat,'' he says.

``Your granny.''

``I am,'' he says. ``You could come with me; you could be a mud clerk.''

``Gee,'' I says, ``thanks.'' But it did sound kind of fun. ``Where would
we go?''

``Anywhere and everywhere,'' he says, rolling over to look at me. His
eyes was bright. ``America would be our oyster! We could go up to
Minneapolis in the north, or New Orleans down south. That's by the
ocean. Wouldn't you like to see the ocean?''

``I reckon,'' I says eagerly.

``And there's lots of other rivers that feed the Mississippi, too. We
could have grand adventures.''

We grinned at each other. He flopped back in the dirt and we dreamed
along for a few minutes.

I wouldn't have to worry about the poorhouse on a steamboat. I wouldn't
be plagued by gins and magis, nor by paps who drank too much. No
worries, no cares\ldots I wondered if there was any place that far away.

I sighed. It was all well and good to dream about, when you was by the
river, but that\textquotesingle s all it was---river talk. And dreams
was good enough for the likes of me, but it sounded like Sam had folks
that wanted him, and a future.

I didn't know much about doing the right thing, never having done it
before, but this seemed straightforward enough. I took a deep breath.

``Don't you reckon your mam and brothers and sissy would miss you?''

He was quiet. ``I would miss them, I lay,'' he says, all soft.
``Especially Henry. He's my little brother. He's my best friend.''

I plucked a leaf from above my head and rubbed it between my fingers. I
reckoned that would do. Don't ask me how I knowed, but we warn't going
on any steamboat trip, not today and not never. Me and Sam couldn't be
brothers. He already had family---everybody already had family---even
Tom Sawyer who was an orphan had his aunty and cousin and half-brother
Sid. Everybody was called for. No one in the world was free to drop
everything and dress like me and act like me and live in the mill with
me and look out for only me.

I felt so lonesome at that moment I thought I'd be sick.

``I reckon I should get back,'' Sam says. Took him long enough. He
squirmed out of the bush and stood, brushing himself off. ``Thanks for
the bush. You was right; it was swell.''

I didn't answer.

``Well, bye,'' he says, and started off. I didn't acknowledge him. I
laced my fingers behind my head and stuck my chin up and stared at the
sky.

Just when he was almost gone up the path, he turned and sang out, ``I'll
take you on that trip down the Mississippi some day, Huckleberry Finn! I
promise!''

``In a pig's eye,'' I says. But he was gone.

Sometimes being Huck Finn is the lonesomest job in the world.

I lay in my bush and watched the sky grow dark through the branches.

No, I didn\textquotesingle t cry. And that ain\textquotesingle t none of
your flaming business anyways, so shove off.

\mychapter{Chapter 6}\label{chapter-6}

It was all Tom's idea.

See, the school boys and girls was having what they call an Examination,
which was this event at night where they stood in front of their parents
and recited poems and stories and truck. The young folks hated it, the
grown folks ate it with a spoon.

Tom Sawyer was especially unenthusiastic about performing in this
exhibition, because he had to recite some humbug poem about a bird and a
man who wants to die. It was full of requiems and tender nights and
pining hearts. Tom said it out for me when we was by the crick, catching
crawfish, and I thought it was about the saddest thing I ever heard.

``It don't matter if you pay me a hundred yaller-boys,'' I says when he
got through, ``I don't care to hear \emph{that} one again.''

``Exactly,'' says Tom. ``It's too blamed down.''

``You'll have the whole crowd in tears,'' I says.

Tom paused with a crawfish in his fingers. ``I would?''

I had a little piece of ham on a string, and I was dangling it at the
entrance of a crawfish hole. The little bugger would creep out, try to
grab it, then pop back in. I dangled it lower.

``Oh, never mind,'' says Tom finally. ``I still don't want to. I hate
examinations. \emph{You'd} hate it. You have to stand in front of
everybody, and it's dead quiet, and the schoolmaster is over in his
chair scowling every time you mess up. I sweat bullets.''

The crawfish sneaked up on the ham and grabbed it with his big claw, and
I tugged gently, and he dug in with his little legs, and I tugged
harder, and he dug deeper, and---\emph{snap!}---I fell over, and the
crawfish was gone, and I was left holding a string with a single, large
claw still pinching it.

I pulled off the claw in disgust and tossed it in the bucket.

``What I wouldn\textquotesingle t give to have something happen that
would cancel the whole thing,'' says Tom.

``What \emph{would} you give?'' I says.

Tom looked at me, and I looked at him, and we started grinning like a
couple mullet-heads.

That's how I found myself climbing the oak outside the church that
afternoon, hunting hornets. And you wouldn't believe it, but there
warn't a single one to be found. They had packed up and moved sometime
between the bean feed and now.

Tom would be heartbroken. He\textquotesingle d been so drunk off our
plan he\textquotesingle d hugged me, and skipped around, and clicked his
heels. And in his excitement he knocked the bucket into the crick and
lost our crawfish, which I was pretty tore up about, since it was my
supper. But he said never mind, he'd feed me a thousand suppers if I got
him out of this. So I was agreed.

I couldn't let him down, so I wracked my brain and tried to think. What
else would break up a school meeting like a load of hornets?

I thought about me and Tom hunting crawfish, and I had an idea.

I'd gotten the meat for the bait behind the general store, where the
owner puts his scraps into covered barrels. He burns them every night,
or else the wild dogs and cats is all over them.

So I went and sidled up behind there holding my empty crawfish pail, and
I opened one of the barrels. It was full of bloody bones and sinews and
gristle, with little bits of raw flesh mixed in. The smell was terrific.
I rolled up my sleeves and just dug into that mess and filled my bucket
to the brim.

Then I moseyed around town and visited all the curs and cats I knew, and
let them get the scent of it. Some was suspicious, so I let them have a
little piece, just so they got the idea. Then I high-tailed it to the
schoolhouse, looking like the dad-blamed pied piper.

I looked in the window, and saw that the littlest children was only just
starting to recite. Good. I was in time. Tom told me they went by age,
youngest first.

I clenched the pail's rusty handle in my teeth and climbed the big oak
by the school. Then I hopped onto the roof and shimmied over to the hole
by the chimbley, the hole which was always getting plugged with rags as
an alternative to fixing it. I reached my hand in, and pulled the rags
out.

Down below, I had a captive audience.

The cats had followed me onto the roof, and was rubbing around on me
trying to get into the bucket. The dogs was down below, baying and
whining and feeling sore about getting left out. That was when I dumped
my truck down the hole.

The cats streamed down after it like a blessed flood, and I can only
imagine what that looked like to the ladies and gents inside. Like the
sky tore open and revealed the stuffing of the universe, and it was
cats.

The dogs on the ground heard the holy hullabaloo and burst through the
schoolhouse doors and windows to get in on the fun. It was the most
fantastic spectacle I'd ever seen in my life.

I slid off the roof and was in the woods when folks streamed out the
door. My work was done. I went and hid under the bridge and waited for
Tom.

Tom was nearly chewed apart in some places, and his face was scratched
within an inch of its life, but he was singing with joy. He grabbed my
hands and danced around.

``Hucky-boy, what a show! You did superb. That was even better than
hornets. What a commotion! The mayor got a cat to the face! And the dogs
went after the cats, and they was all fighting over the meat. Best of
all, I didn't have to exhibit!''

He jumped and grabbed a beam under the bridge and swung from it. I
grabbed him round the middle and tackled him to the ground and we
tussled, laughing and wrestling.

Suddenly the bridge creaked above our heads. We stopped in an instant
and held still, panting. Someone walked over the bridge. But it sounded
odd. Their feet went skippy-skip, skippy-tap, tap tap tap.

It was the magi. We could see him through the cracks in the boards. He
was grinning and humming, his spine arched in a deep hook, head bobbing,
elbows akimbo.

``That ain't a good sign,'' says Tom in a whisper. ``The gin ain't with
him.''

``So?''

``So, don't you remember what I told you? When a gin gets loose, he
makes trouble. They're evil spirits, Huck. All they like is trouble.
Mark my words, something bad is going to happen.''

This brought the mood down for me, but Tom was still pumped full of
happy. He promised me a full-spread supper anytime I stopped by, and
that made \emph{me} happy. We hung out under the bridge with our arms
around each other for hours, talking nonsense and congratulating
ourselves on how brilliant we was.

Finally as the sun set we split, him going down the road to home,
whistling, and me back into town to rustle up some grub.

There was a big commotion going on in front of the general store. I
eased into the crowd to see what the fuss was about.

The general store man was talking to the constable, and several lawmen
stood nearby swinging billy clubs. When they saw me, they sang out,
``There he is! Nab him!''

I tried to bolt, but seven hands reached out and grabbed me and held me
fast. I leaped and twisted and jabbed, but there warn't no shaking them.

``Huck Finn,'' says the constable. ``Come here, please.''

I had no idea what I was stepping into, but I went. I had no choice; all
those hands tugged me along.

``Have you been to the general store this evening?'' the constable says.

That's where I got the meat scraps from. Bugger. They knew.

``Ain't been nowhere near it.''

But the constable had noticed me hesitate just a second too long. ``And
did you happen to take anything away with you when you \emph{left} the
general store this evening?''

I hooded my eyes and set my jaw and determined to say nothing more. But
the constable nodded, as if I'd owned it all.

``Huckleberry Finn, I arrest you on suspicion of breaking and entering,
and burglary. Take him, boys.''

And the boys took me---straight to jail.

\mychapter{Chapter 7}\label{chapter-7}

I been in that jail before. I knowed it pretty well. It was a tiny thing
away at the edge of town. It warn't even guarded. I sat inside and
waited, and wondered. I never seen such fuss over a load of scraps.

Next morning they hauled me to court and chewed me over. The postmaster
testified he seen me hanging around behind the general store while the
owner was away. He said I looked ``mighty shifty.''

Then the owner returned and saw his shop had been broke into, and all
the money in the till was gone.

That's what it was about. Some crime I didn't know nothing about. I was
so mad I could bite, but I had to sit and take it and I warn't permitted
to say a word, which didn\textquotesingle t feel fair.

``How much money was in there, to the best of your knowledge?'' says Mr.
Riverson, the lawyer.

``Almost fifty dollars,'' says the general store man. ``It was the
week's income. I was taking it to the bank in the morning.''

The trial went on and on, and after a while I put my head down on the
bar and closed my eyes. Not that it warn't interesting. But I was tired
from spending the night in jail. All night long I imagined I heard
ghosts and devils outside the cell, and inside with me. Besides I felt
so low-down and rotten being locked away like that, with one thing and
t'other, I didn't get much sleep.

The judge asked me if I took the till. I lifted my head and told him no.
I told him this in no uncertain terms. And I spat.

``Where were you at the time the money was stolen?'' he says.

I thought, \emph{I was on the roof of the schoolhouse, flooding it with
rotten meat and cats}. Tom Sawyer could a vouched for me. But Tom Sawyer
warn't there.

If I told the truth, I\textquotesingle d get Tom in trouble.

If the truth was to be told, Tom would have to tell it.

I kept silent.

The judge asked if I would return the till, if I was let off light. I
said no.

The judge asked if I would pay a fine. I said no.

He says, ``Based on witness reports and prior offenses, and given that
the lad refuses to account for where he was at the time of the burglary,
I find him to be guilty. Penalty is public whipping, to be administered
at seven o'clock this evening at the whipping post in the square.''

The constable stood. ``Your honor, seeing as he's just a boy, I propose
we don't administer the full thirty-nine.''

The judge allowed it. ``We'll do ten.'' And the trial adjourned.

I was glad I warn't a judge. A courtroom's the most wretched place on
earth. And I was only in there a couple hours. I could only imagine how
much worse it must be for the man who's got to talk in front of
everybody, and argue all day, and say things he don't mean.

I got put back in jail to wait for seven o'clock. Tom Sawyer stopped by
a little after noon.

``Hi-ya, Huck.''

He peered in the barred window. I was sitting in the corner with my arms
around my knees.

``How're things?'' he says.

``Swell. Just swell.''

I waited for him to apologize for not coming to my trial, nor speaking
up nor vouching for me.

``You scared?'' he says instead.

I acted as if this idea was contemptible. ``I been whipped loads of
times. Ain't but old pie.'' This was strictly bilge. I got the stuffing
knocked out of me now and then, but I ain\textquotesingle t never been
whipped in public by lawmen. I was scared senseless.

I stood up. ``You fixing to tell `em where I really was yesterday
afternoon?''

He looked uncomfortable. ``What do you mean?''

I approached the window. ``You know I was with you, pulling that prank
at the schoolhouse. So I couldn't a lifted the till.''

Tom rubbed one of the rusty bars with his thumb. ``I warn't with you all
day\ldots''

I gasped. My face flooded with heat.

Why, that lousy, rotten, no-good, two-faced---

``You want me to bring you anything?''

I shot my arm through the window and grabbed his hair, yanking his face
close. ``Yeah. Bring me a bucket of meat scraps. I want to ram `em down
your throat.''

He prised my fingers loose and leaped back a jump. He looked scared.

I threw my hands up in disgust. ``Forget you, Tom Sawyer.''

I retreated to the rear of my cell and curled up on the ground, facing
the corner.

``Huck?'' Tom rapped on the bars. ``We're still friends, right? You know
it don\textquotesingle t matter to me if you took the money. Right?''

I ignored him, and after a while he went away.

I slept a few hours. When I woke there was small things hitting my back.
I looked around. Tom was at the window again, throwing pebbles.

``Hi,'' he says, all soft.

I growled. ``Can't a fellow at least find peace in jail?''

``Brung you something.'' He held up a bundle tied in a handkerchief. I
could smell it was food.

I went to the window, and Tom pushed the bundle through.

Inside the handkerchief was a big wedge of election cake, studded with
raisins and cherries and blueberries, plus an apple, a piece of cheese,
and a roll, which turned out to be a Sally Lunn roll packed with lemon
peel.

``Ain't a full-course meal, like I promised,'' Tom says, picking at the
bricks around the window. ``But I hope you like it.''

My stomach growled audibly. I shoved the bundle through the bars.
``Ain't hungry.''

Tom's eyes welled up.

``I--I got to go,'' he says. He scrubbed his face with the heel of his
palm. ``They rescheduled the schoolhouse examination for tonight. At
seven.''

So it had all been for nothing.

But Tom Sawyer would be roasting on his stage while I roasted on mine. I
sneered. ``Good.''

The constable came to get me a hair before seven. He led me to the post
in the square and stood next to me as his deputy tied my hands to it. He
scanned the crowd as if looking for someone.

``Master Finn,'' he says quietly. ``Not too late to return the money.''

``Mister Constable,'' I says, ``not too late to swap the danna in your
skull for some brains.''

He gave a tight smile. ``I'm surprised your father didn't show up to
make a scene. That's his style, isn't it?''

I scowled. ``I take my own licks.''

He glanced once more around the square. He seemed disappointed about
something. Clasping his hands behind his back, he turned away. ``Fine.
Take them, then.''

I took `em.

It warn't that the tanning itself was so ornery. It was that everyone
was watching. All the townsfolk. I guess you get to where you feel
comfortable with the people you see every day. Then something happens
like you get whipped up in front of them and they just watch, and you
feel like a dog.

The gin and the magi was there. The magi shook his head like he was real
sorry about the whole affair. The gin never took his eyes from me, and
his head moved side to side like a cobra scenting the air.

If the constable had just taken me somewhere private to do the deed, I
don't reckon I would a minded so much.

\mychapter{Chapter 8}\label{chapter-8}

I stayed out of town after that for the rest of the summer. It was awful
comfortable in that old mill of mine. I made a rough table and a
split-bottom chair out of wood. I reattached the door with some new
leather hinges. I also catched some logs floating by on the river now
and then, and sold them to the sawmill. The river was high that time of
year, and a lot of logs and truck floated along gentle as you please,
asking to be scooped. So I scooped them.

Life on that river was probably as close to the promised land as I would
get. The leaves was full green, and cicadas screamed during the day, and
crickets sang at night, and frogs would join in evenings and the three
would together make the most glorious din you ever heard, buzzing and
hollering and singing, so if you closed your eyes and listened you got
this thrill deep into the pit of your stomach and the music surrounded
you and filled you to the brim it was so soulful and beautiful, like
angels.

My favorite trees was cottonwoods, because they filled the air with
gauzy white fluff. It floated like snowflakes, getting in your hair and
collecting on your clothing and swirling like faerie dust. It built into
piles on the ground and turned that muddy riverbank into an enchanted
place.

One day I didn't catch any fish and my stomach was like to eat itself,
so I hunted up some berries and eat them. I eat so many I thought I
would bust, and that night I was in the most awful pain, I thought I
poisoned myself. I wracked my brains, trying to remember exactly what
the berries had looked like, and if they \emph{was} poison, but I
couldn't remember. One minute they was blueberries, the next
pokeberries, until I got so confused I didn't know my own name.

I reckoned I would die alone in that mill, and no one would find me for
weeks or even months. I would be an awful sight. It made me sick to
think how I would look when they finally found me.

Them's just the kinds of things you think about when you live alone.
It's best not to let your mind run free too much when it gets on one of
those tracks.

And then the next day I catched a coon and lived off it like a king for
a week. And I guess you've eat coon before, but maybe you ain't done it
the best way, so I'll go ahead and tell you how to cook it right.

Well, first thing is to dress it down. That means skinning it and
throwing out the musk glands and trimming the fat---and you must
absolutely get all that fat off you can, because it goes rancid in the
wink of an eye. Take my word; I learned this lesson the hard way.

Next just soak the coon in salted water overnight to draw out the blood.
Come morning, cut it into little pieces and toss it in a bit of meal or
flour, whatever you got. Brown the meat in the hot fat of your skillet,
cut up some onion and a tater or carrot if you got them, cover the whole
mess and let it simmer a couple hours. Add some more flour and water to
work up a gravy, whip up a couple corn dodgers for dipping, and lawsy.
You ain't never eat something so near heaven in your sainted life.

Finally along came the red leaves and the long afternoons, and I
reckoned it was about time to show my face in town again. So I skipped
in to check out the apple festival.

The apple festival is this thing the Ladies' Guild does when the harvest
rolls in. There's contests for best apple preserves and apple pies and
apple donuts and apple sausage and all that truck, and the young folks
go into the orchards and pick apples and eat as many as they can until
they bust. It's what you might call a good time.

It was real hot when I got there, and some fellows was wallowing out a
big patch of mud for wrestling, so I shucked my clothes and jumped in.
We brung sloshing buckets up from the crick and stomped around, pushing
each other and acting the fools. We was all over mud from our heads to
our toes, and it was the most delicious thing you ever felt, slimy and
squishy and slick, and so blessed cool.

It got pretty rough in there. The Sunday-school superintendent, Mr.
Walters, was a bully wrestler, and he pinned me twice, till I had to
holler `nuff. That man had no mercy. And Muff Potter was a champion mud
wrestler, since there was just plumb so \emph{much} of him.
He\textquotesingle d squish me into the hole till I thought I was
drownded. But then he showed me what he done afterwards, and learned me
it so I could do it myself next time. Muff Potter's all right.

We was having a ripping time, so it took me a while to notice Tom Sawyer
standing not far off. The young swells had returned from the orchards
and milled about, flirting amongst theirselves, crisp and starchy. Tom
stood under a tree by himself with his arms wrapped around his middle.
He watched me wrestle with a pinched, yearning sort of look.

I hadn\textquotesingle t seen him since I got walloped. And even though
I\textquotesingle d been livid at him then, I warn\textquotesingle t
livid now. I\textquotesingle ll allow I sort of missed him.

Plus the look of him made me sorry---trussed to the gills in itchy,
stiffy clothes and standing carefully in the grass to preserve his
shoes.

I clambered out of the mud and slouched over.

Tom watched me approach, looking like a kicked dog.

I got in his face and says, ``\textquotesingle What\textquotesingle s
the idea? No \textquotesingle Hi-ya\textquotesingle? No
\textquotesingle Hey, Hucky\textquotesingle? Ain\textquotesingle t we
speaking?''

Tom looked surprised. ``Hey.''

``Hey yourself. Look, when you\textquotesingle re pals with someone, you
don\textquotesingle t stand off and stare at him like a bedlamite. That
ain\textquotesingle t no way to behave.''

I poked the end of his freckled nose with my finger, leaving a big glob
of mud.

Tom gasped. 11Thanks! I mean...'' He looked down. He wiggled his nose like
it itched. ``I\textquotesingle m glad you showed up today.
It\textquotesingle s been a while. And I been thinking...a lot...about
last time...''

I rolled my eyes and peeled a clump of muddy hair out of my face. ``Just
forget it.''

Tom raised his head and glared. ``No! Listen. I got to say it. I been
thinking about it all summer. I know you didn\textquotesingle t take the
money! You might take food or clothes or what you need to survive, but
you\textquotesingle d never steal a till. And you needed me to speak up
for you, but I was too---too---''

He took off his hat and kneaded the brim, searching for the words. He
sighed in frustration. ``Look, if your pal\textquotesingle s in the soup,
you don\textquotesingle t leave him there. Not even if it risks your own
hide. Dartanyen wouldn\textquotesingle t have left Athos. Portos
wouldn\textquotesingle t have left Aramee. And I...left you.''

He hid his face in the hat. ``I\textquotesingle m sorry. I guess
I\textquotesingle m not much like a book hero. I\textquotesingle m not
even much like your friend.''

I heard a wet little sniffle. I pulled the hat away. His face was red
and blotchy, his nose running, his eyes squinched up and watery.

I slapped him. Not hard. Kind of gentle. His cheek splattered with mud.

Tom\textquotesingle s eyes got real big. He touched his face.

``I said, forget it.'' I grinned at him. ``I have.''

His eyes filled with gladness. He beamed.

He opened his mouth to say something, but his gaze focused behind me. I
turned. Constable Morley sauntered over, hands clasped behind his back.

``Huc-kle-ber-ry \emph{Finn,}'' he says.

``Con-sta-ble \emph{Morley}.'' I looked back to Tom, but he was gone,
disappeared like a jack-o'-lantern in the marsh.

``Haven't seen you around these parts in a while,'' the constable says.
``How's your father?''

Suddenly I remembered that the constable wanted to put me away in a
poorhouse because I didn\textquotesingle t have a pap nor a living nor a
home.

``Just fine, thanks. He'll be real honored when I tell him you asked.
Him and me are living in Goshen now. We're working at the lumbermill up
thataways. It's hard work, I mean sometimes it's dangerous and hot, but
it pays well. I mean it don't pay \emph{much,} because it's just rough
work, but we don't complain, because we're grateful. We live in a
one-room shanty down there---I mean up there---in Goshen. It's small but
it's comfortable. Pap is teaching me to play the fiddle.''

I have no idea where the fiddle came from. I knowed I was rambling, but
sometimes when I'm nervous my mouth runs. It was sort of enjoyable to
fantasize the grand times me and Pap was having, at least.

``Really?'' the constable says. ``That\textquotesingle s wonderful. Which
lumbermill? Rawley\textquotesingle s?''

``That\textquotesingle s the one.''

The constable nodded. ``Rawley closed his lumbermill two years ago on
account of his rheumatism. I know. He\textquotesingle s my uncle.
You\textquotesingle re a terrible liar.''

I bit my lip. That constable didn\textquotesingle t play fair.

Constable Morley smiled grimly. ``Enjoy the festival.'' He walked off,
and I let out my breath like I\textquotesingle d held it a year.

I didn't feel much like festivaling after that, so I crouched under a
tree and watched.

``Have an apple, Huck?'' It was Tom again. He\textquotesingle d cleaned
the mud off his face. ``It's a Milum.''

I took it. It was small and sweet. The outside was yellow and red, and
the inside was crisp and white as snow.

``I'm drying out,'' I says when I finished. ``I\textquotesingle m going
to wash off. Want to come?'' Tom eagerly agreed, and he even grabbed my
clothes from beside the mud pit as we passed. I aimed and threw my apple
core, and knocked a gent\textquotesingle s tophat clean off his head.
The man clapped his hand to his pate and whirled, trying to see who done
it, but we kept walking, making like nothing happened. Once we was out
of earshot we bust out laughing.

We tripped and skipped down to the crick, listening to boys and girls
yelling as they chased each other in and out of the trees.

The crick was small and rocky, with clear trickling water. I splashed
and wallowed, flinging water everywhere and shouting because it was like
ice.

``Come on in,'' I says to Tom, which was cruel of me.

He was burning to do it, I could tell, to get out of them clothes and
cool off. But I knowed, and he knowed, that if he did, his Aunt Polly
would lick him with a switch. I watched and grinned, waiting to see what
he would do.

The struggle was too much for his little heart to bear. First he waded;
he pulled off just his stockings and shoes. Then I splashed him, and he
splashed back, and that got his sleeves wet, so he took off his
roundabout and pretty soon all his things was off, and we was whooping
like a couple of savages, dunking each other and slipping about on the
slick green rocks.

His face was grave as we straggled back to the festival. I tried to keep
his mind off his licking by telling him jokes, and tripping him, and
putting bugs in his hair.

He was carrying what clothes he couldn't put back on, because some of
that awful garb he has to be sewn into, with a needle and thread.
That\textquotesingle s honest-to-goodness fact; I
couldn\textquotesingle t make it up.

When we got back to the festival, Tom ran off to see some of his pals. I
pinched some food. One lady catched me and scolded me, but the rest was
so wrapped up in jawing they wouldn't a noticed if I'd flipped a table.

``Afternoon,'' says a voice behind me. I whirled. It was the magi, with
the gin right behind. I was immediately fan-tods all over, because I
didn't know if they was going to jump me or what they wanted, all I knew
was that they finally catched me. I looked round for the constable, but
now I didn't see him nowhere. They ain\textquotesingle t around when you
need them.

``Howdy,'' I says. I turned to light out, but the magi whipped out quick
as a snake and curled his bony fingers around the back of my neck, and
then I couldn't a budged if I was hit by a bull.

The magi\textquotesingle s grey hair hung stragglier than ever, and his
teeth when he grinned was brown like a horse\textquotesingle s, and
crooked. The gin, glowering behind him, was still covered tip to toe in
red fabric, mighty dirty and rank. The gin stared out at me from the
folds of the cloth with black eyes, and his head made that swaying
motion.

``What's your name, my dear?'' the magi crooned in a baby-talk voice.
His long spine curled even deeper so he could peer into my face.

``Joe Harper,'' I says.

``Is that \emph{your} charming friend yonder?''

I followed his gaze and saw Tom, in his element: surrounded by girls and
boys. He was organizing some sort of game, and they hung to his every
word. He pointed and they ran. He nodded, and boys leaped and girls
clapped. The littlest boy, James, ran up to Tom and showed him
something, and Tom approved, and Little James fairly blew away with joy.
They swarmed all over him.

``You\textquotesingle re friends with that boy?'' the magi says again.

``Never seen him in my life.''

``He\textquotesingle s quite the darling of the town,'' says the magi,
showing his teeth. I got a creepy feeling.

``Wouldn't know. Not from around here,'' I says. ``Well, so long.''

I tried to walk away, but the magi didn't release my neck, so I jerked
back.

He nuzzled his long, hooked nose into my hair and spoke in my ear, all
intimate and private-like.``You acquainted with a jake by the name of
Huckleberry Finn?''

``Ain\textquotesingle t.'' I jerked my ear away from his snout and
frowned at him. ``Admit it; you made that name up. Ain\textquotesingle t
no fool alive named after a berry.''

``I have a surprise for him,'' the magi weedled. ``He's heir to a fortune
in Persian jewels.'' (I almost snorted, but catched myself.)
``Unfortunately, I lost the will. It was stolen, out of my room, under
my very nose.''

His face crept closer until the tip of his nose pressed against mine.
His eyes peered into mine like he could read the truth from my brain.

I focused my eyes over his shoulder and studied the mud wrestlers as if
they was the most fascinating things in the world. ``Did you---see who
stole it?'' I says, trying to keep my voice steady.

``Alas. I was within twelve inches of him, but it was too dark to make
him out.''

``Shame,'' I says, trembling with relief. ``Well, hope you find the
will.''

``So do I, lad,'' he says. ``And so should Huck Finn.''

And then to my joy the magi uncurled his fingers from my neck, and I
scampered.

Tom returned to me later, airy and expansive, just as the powwow was
winding down. He must a been having a grand time. He flung an arm across
my shoulders.

``This festival is for kiddies and old ladies, Hucky,'' he says. ``I
reckon we could show this town a good time.''

``How's that?''

``In Europe and New York and places, \emph{those} people know how to
throw parties. Look at us. A mud hole and apples. We need to organize
something big. Something that will knock their socks off. We need a
cotillion.''

``A \emph{what?''}

``It's like a dance. You dress up like a prince, and wear a mask so no
one knows you, and eat little fancy truck and drink punch and dance all
night, till dawn breaks the next day.''

``I don't reckon we ever had one of them in St. Petersburg,'' I says.

``Course we haven't,'' he says, gesturing at the general assemblage.
``Look at us.''

That got my hackles up. What was wrong with us? ``Well, you go and have
yourself a grand old time, but I don't reckon I can go to no
coe-till-yun, because I ain't no prince and ain't got a mask nor
nothing. Besides, where would we have such a blame thing? And what fancy
things would we make to eat, we who if we rubbed our skulls together
could prob'bly make a buttermilk biscuit?''

``Now looky there, that's your problem, you ain\textquotesingle t got no
vision,'' he says. He looked towards the tables where the Ladies was
gabbing. ``I know exactly where we can get as many fancy vittles as our
stomachs can hold.''

He rounded up a pack of boys, Joe Harper and Ben Rogers and Billy Fisher
and Johnny Miller and me, and told us his idea, and the other boys
thought it was swell, so over we straggled to the Guild Ladies, trying
to look nonthreatening.

We let Tom do all the talking.

``Miss Robinson, this is the best apple festival this town has ever
seen,'' he started. ``Why, all the cakes and pies and things is
beautiful. I know for sure we're going to bungle ours. In fact, Joe, I
don't think we ought to have ours at all. I hope you lot won't be too
disappointed.''

We set up a-wailing. The ladies looked on with interest.

``Why, Tom Sawyer, whatever are you talking about?'' says Miss Robinson.

``Well, the young folks in town, meaning me and my pals, we was planning
on having---no, it's too silly. You\textquotesingle ll laugh.''

``Go on, Tom,'' says another lady. ``Please tell us, we promise we won't
laugh, no matter if it \emph{is} silly.''

``Well, we dreamed of having our own little dance. Sort of a cotillion,
you know. Just a little party that we could organize wholly by
ourselves, and do all the work, and have the satisfaction of a little
culture and society in our small lives. But we see how much work goes
into planning a function, and besides we ain't clever at making treats
like these here. But it was only a funny little idea, and not worth
mentioning.''

``Well, you have mentioned it, and I don't see any reason why you can't
have it, either,'' says Mrs. Rogers, Ben's mother. ``Why, food is only a
little part of a dance. You can still get together in the schoolyard and
have your function.''

``You're not listening, Letitia,'' says Miss Robinson. ``They want a
\emph{fancy} party, with music and \emph{hors d'oeuvres} and punch.
Isn't that right, children?''

We indicated gravely that it was so.

``Ladies, a word?'' Miss Robinson says. The ladies formed a huddle and
conferred, and we boys formed a huddle and conferred, too. Our
conferring was mostly a lot of nudges and winks and Tom putting a finger
alongside his nose. Then we all emerged and faced off again.

``Tom, the ladies and I were wondering---if you don't object---wouldn't
you like to have your function after a Ladies' Guild meeting? We always
have refreshments at those, nothing excessive, but if you children come
we might have cake. Would that be all right?''

We all looked at Tom with our mouths watering, but Tom he shook his
head.

``You ladies is too, too kind. But we couldn't ask such a bother of you,
and besides, we did want to do it on our own.''

The ladies turned and conferred again, and we practically throttled Tom
conferring him in our huddle.

``What's the idea, anyways, you twit?'' says Billy. ``They was going to
make us cake.''

Tom allowed it was a pretty grim day when the likes of us couldn't trust
the likes of him, and had he ever steered us wrong before?, and wouldn't
we rather just wait and see how the glorious plan played out, instead of
panicking like a pack of pigeon-toed pin-heads?

When the huddles broke up this time, the ladies was beaming.

``We love that you children want to do everything yourselves, it shows
real gumption. But won't you let us help a little? What if you assign
each of your little friends to a Lady, and they make the treats
together? Finger sandwiches and cookies and cake and punch and all sorts
of lovely things? The Lady would let you do all the measuring and
mixing, she would even let you pick the recipe. She would just be there
to provide her kitchen and any little aid you may require. How does that
sound? Would that be suitable, Tom? You won't begrudge us a little fun,
if we promise to leave you alone at your party?''

Tom made a big show of cyphering it out, and finally he turned to us and
asked if we would agree to it, and wouldn't be too disappointed if he
said yes. As for us, we put forth it was agreeable almost before he
finished speaking.

So it was settled, and we was having a ball.

\mychapter{Chapter 9}\label{chapter-9}

I warn't too keen on the idea of a cotillion, and didn't figure on
going. It sounded like a situation where one was obliged to scrub and
brush, and stand around and be looked at. That sounded mighty
uncomfortable to me. If I knowed one thing, it was that Huck Finn warn't
brought up for Society, and Society warn't brought up for Huck Finn.

The nights was cooler by the river, and I got more and more lonesome
living out there by myself. I missed knowing what was what in town. I
used to know just about everything that went on, and at the apple
festival I realized I didn't know nothing. I hadn't gotten my water-pump
gossip in months, and I hadn't touched my old post fence outside the
general store since early summer.

So what with one thing and another, I started heading into town more and
more, till I was there nearly every day. And if the constable saw me, he
didn't say nothing, and I didn't say nothing, and we lived and let live.
And the gin and magi warn't to be seen, so they must a moved on up the
river, which suited me fine.

First thing I learned when I put my ear to the ground was everybody was
in fits about Tom's party. Every young person wanted to go, and some
older folks wanted to go, too. It was up to Tom Sawyer though, and he
said only young folks was allowed. The grown people talked of having
their own dance, but it was weak talk and soon died out. If it warn't
Tom Sawyer's party, it warn't worth a shaved rat.

Tom passed out slips of paper to his friends, with the name of the Guild
Lady they was assigned to cook with. And he found me sitting in front of
the general store smoking, and gave me a slip, and I tore it in half and
tossed it in the dirt.

``Go to blazes,'' says I.

``Now, Hucky, don't be like that,'' Tom says. ``You know if you want to
dress up and eat fancy truck and dance, you got to make something. Ain't
nothing to it,'' he adds, all soft and persuasive. I hocked up and spat
on his toe. And don't go feeling bad for him, neither. I was wise to his
Tom-Sawyerish ways. He warn't fooling me one second.

``Huck, this is the very best slip,'' he says next. ``I reckon all the
other boys would be sick if they saw what I saved for you. But if you're
sure, and no changing your mind, I won't tell you what it is. No point
arguing with a brick wall.''

I eyed him from under the brim of my hat, taking a long draw from my cob
pipe. ``Well...I lay it wouldn't hurt to know what it was.''

Tom grinned. ``Best of all. The cake, with the Widow Douglas.''

``I don't make cake.''

``The widow does. But she won\textquotesingle t make it if someone from
the party don't help. That's the deal.''

``Get Joe Harper.''

``Joe's making cookies with my Aunt Polly.''

``Get Billy Fisher or Ben Rogers.''

``They can't, saphead, don't you understand? They're already paired with
other Ladies. I saved yours for last, because no one would refuse a slip
like \emph{this}.''

He looked down and rubbed the toe of his shoe in the dirt. He glanced
up, eyes soulful. ``The boys and girls will be mighty disappointed if
there ain\textquotesingle t no cake.''

I blew a lungful of smoke in his face and walked away. His tricks don't
work on me.

But the next day, I got to thinking it wouldn't hurt just to drop round
the widow's place and sort of look around.

The Widow Douglas was real young. A couple years ago she married a judge
who was ancient, who promptly kicked the bucket. According to word
around town, she used her youth and her looks to ensnare him, but all
she wanted was his money. Folks sneered when they said it, and called
her a few names.

The widow now lived alone in a big house on top of Cardiff Hill, the
biggest house in St. Petersburg. And people still sneered behind her
back.

I looked round the house and found the kitchen, which was at the back. I
peeked in there. It was a nice kitchen. Big. I reckoned if a boy made a
cake in there he'd enjoy himself. It had a big black stove and shining
black skillets, and copper pots and pans hanging from the beams, and a
wood-rank neat as a pin. There was a pump for water, and a pantry
stocked to the gills with all sorts of sumptuous truck.

I guess I got in there a little too far, nosing around. Next thing I
knowed I was looking in a drawer, and I knowed I was in trouble because
it was full of silver, and just then someone grabbed my collar. I nearly
shot through the roof.

``I didn't!'' I says.

``Huckleberry Finn, I presume?'' says a lady's voice. I gulped and
twisted around.

``Widow Douglas, I presume?''

``I asked you first.'' Her expression was stern, but her mouth twisted in
one corner. Her yaller hair was in a huge roll all the way round her
head.

``Yes'm.''

``Well?''

``Yes'm.''

She let go of my collar and I took one huge jump towards the door, then
stopped and turned. She stood watching me with folded arms.

``You should've been here yesterday.''

I swallowed. ``Y---yes'm.''

``I suppose we should get cracking.''

I glanced towards the open door.

``I---I reckon so.''

She fetched a little wooden box from the pantry. ``Your friend Tom
Sawyer already explained the procedure. You pick the cake, and I lend
you the use of my kitchen, and help as needed.''

I took the box and nearly dropped it; my hands was shaking. I put it on
the counter quick. It was full of little cards, each covered in squiggly
black writing.

I stared at those squiggles hard, willing them to turn into something
with meaning. ``I reckon---whatever cake you like best, let's make that
one.''

She waved a hand. ``Certainly not, don\textquotesingle t let me
influence you. I have several in that box that are treats. I've a
chocolate cake that would knock you flat, and an election cake that
rises like a song. You really can't go wrong with any of them. Choose
whatever you like.''

I shuffled the cards nervously. There warn't any pictures. I couldn't
even tell which was cakes. What if I picked a card at random, and it was
soup?

The widow watched as I flicked through. I didn't know what to say, and
she didn't say nothing. I grew more and more frustrated. Pretty soon I
couldn't even see the writing on the cards. My vision swam.

I snapped the box shut and rubbed my eyes angrily.

``I don't like any of `em.'' I shoved the box to her and stuffed my
hands in my pockets.

The widow studied me, and I tried to return her gaze, but found I
somehow couldn\textquotesingle t. I pulled my pipe out and turned it in
my hands, and suddenly I wanted a smoke more'n anything in the world,
but I dasn't light up in her kitchen.

``You're right,'' she says. I looked up startlish. ``None of those are
the thing. Follow me.''

She went out, and I couldn't think of the right thing to do, so I
followed her.

We went through a covered porch into her house. She took me to the
parlor, which was real nice. The furniture was upholstered in dark
velvet. Its wood shined dark, with no scratches. It looked store-bought,
or nice enough to be, even if it was old.

She picked a magazine off a little round table and flipped through.

``Such a clever recipe, I was praying for an occasion to try it. I hope
you will consider it.'' She held it out, and showed me a picture of a
frilly purple cake, and it looked dreadful nice.

``I don't reckon I can make anything like that,'' I says. ``It is real
pretty.'' I didn't want to hurt her feelings. But it looked beyond my
reach.

``Perhaps the two of us together might have some luck with it,'' she
says. Then she smiled, and before I thought a minute I smiled back.

We returned to the kitchen.

``Huckleberry, I have a confession,'' she says. ``I hate searching for
things. Drives me mad. Would you mind if I read out the ingredients, and
you hunted them up?''

Badly relieved, I told her that suited me fine.

We needed a powerful lot of things. Flour, and butter, and eggs, and
milk, and sugar, and lavender, and salt, and I don't know what else. But
I found them all, and we rolled up our sleeves and set to.

The widow was blamed particular on how I did everything. She showed me
how to \emph{sift} the flour; you couldn't dump it in like it was
cornmeal. It was too much hassle for one derned cake, in my opinion.

It took most of the day. But finally it was done, and was it a thing of
beauty. It made my mouth well up, looking at it.

``There,'' says the widow. She wiped her hands on her apron, satisfied.

``Lordy,'' I agreed, and wiped my hands on the front of my shirt.

She nodded. We was wore out, but proud as mother hens.

``I reckon the lavender flowers on top was a good idea,'' I says.

``It was a magnificent idea,'' she says. I nodded.

``I hope it'll be enough,'' I says.

``Cut it small, it'll go around,'' she says.

Eventually there warn't nothing more to do, so I lit out and shambled
down the hill kind of slow, more tired than if I'd spent the day
splitting wood, but pretty well satisfied with our work.

And the next day was TOM SAWYER'S PARTY.

\mychapter{Chapter 10}\label{chapter-10}

Every way you looked, boys and girls scampered. Some was already in
their party dress, and the sun barely up in the east.

I sat by the water pump at daybreak, and all I got was a wet seat. The
talk was rubbage about clothes and food and music. It made me sick. I
got up and went wandering, looking for some kind of fun. I found Johnny
Miller and asked if he wanted to trade me his taw marble for a red
ticket or a fishhook. But he asked me, did I think it mattered if he had
both brass and copper buttons on the same jacket?, and I got disgusted
and left.

All the excitement got me kind a curious to see what Tom Sawyer was up
to on the dawn of his great triumph. So I went to his house and hopped
the fence and peered in the window to the breakfast room. He was just
sitting down to breakfast, hair a haystack and rubbing the sleep out of
his eyes.

Aunt Polly warn't home; I knowed because I already saw her and some
other ladies at the Temperance Inn, sweeping the floor and giving the
place a thorough cleaning, which I reckoned was a rough shock to the
poor Inn, as I doubted it had ever seen a bucket of water in its life.

I jimmied the breakfast-room window and slipped in while Tom was digging
in a jelly jar, and I grabbed a knife that wouldn't cut butter and held
it against his throat from behind.

``Move and yer dead,'' I says.

Tom's eyes got real big. ``Huck?''

``Show me the silver,'' I says. ``Make it quick.''

``Over there somewheres,'' he says, gesturing with his spoon. ``Want
some tea?''

``Coffee,'' I says, and put down the knife and poured up.

``If only you \emph{was} a robber,'' Tom says glumly, putting buttermilk
on his bread. ``How rum that would a been.''

``You sure got the town in a state,'' I says. Tom's eyes lit up.

``Have I?'' he says dreamily. ``How delicious.''

``Deranged, you mean. What did you have to go and get everybody stirred
up for?''

``Who would a thought our little town had it in it?'' he says. ``Oh, if
I could only go back. I wouldn't a made it a piddly cotillion. I'd a
made it\ldots a fête\ldots{} a gala\ldots a \emph{soirée\ldots''}

He was off his gourd. I throwed a slice of bread at him. ``Shut it.''

He shoveled food in his mouth. ``Must fly. So much to attend to\ldots''

Ordinarily I'd a helped myself to some of everything that was on that
table: porridge, buttermilk biscuits, toast, sausage, fruit sliced in
silver cups. His aunty never did things by halves. But that morning I
had a blazing stomachache, on account of something I ate the night
before.

Sidney came into the room, saw me, screamed, and disappeared. Tom didn't
stop shoveling his eggs.

``I\textquotesingle m going to find a place to hide till this blows
over,'' I says, standing up. Tom glanced me over and frowned.

``Ain't you getting ready?''

``Don't reckon I need to change clothes to take in a trot line.''

Tom sat back, alarmed. ``But you're \emph{going.''}

``Ain't you listening? I \emph{ain't.} The likes of Huck Finn don't
belong at the likes of no gosh-derned cotillion.''

``But\ldots but\ldots you have to.'' Tom jumped up and leaned right into
the jellies to grab my sleeve. ``I schemed this whole thing so we could
have a good time and do something new and eat cake. Didn't you make
cake?''

``I---made it.'' I wrenched my sleeve out of his grasp.

``Then you're going.'' He was relieved. He sat down.

I left in disgust. Sometimes there just warn't no getting through to
that boy.

As I was leaving the breakfast room into the corridor, Mary rushed into
the house from the front door and brushed past me. She was dressed real
nice and looked hopping mad.

``Tom, Aunt Polly wanted me to tell you that the Widow Douglas told
\emph{her} that the cake for the party was stolen.''

I lingered in the hall, just around the corner, to hear what Tom would
say, but I didn't hear a peep. So then I got this bad ominous feeling
and I flew through the house and flung myself out the door and down the
porch, but before I took two jumps down the road Tom came hurtling from
nowhere and grabbed my ankle, and I hit the road with a thud.

Tom was flat on his stomach, hands around my foot. He sat up and brushed
at his shirtfront.

``Huck,'' he says conversationally, ``you left in a hurry.''

I wriggled a little ways and sat up, dazed.

``Huckleberry,'' says Tom. ``I wonder if you heard what Mary said just
now, in the house. Concerning the cake. The cake you made with the Widow
Douglas. You know. The cake.''

I nodded once.

``Did you hear?''

Another nod.

``Oh, you did? I wondered if you missed it, moving out like you did. The
cake is missing. Ain't that a shame?''

I didn't give away nothing.

``Now, who could a knowed that cake was in the good widow's home last
night? Did you tell anyone? You didn't? I don't think \emph{I}
did\ldots{} And I don't suppose a gang of thieves broke in and took
it.\emph{''}

I watched him through narrowed eyes.

Tom cut to the point. ``Huck. Did you take the cake?''

I didn't move.

``You stole the cake?'' He clapped a hand to his forehead. ``The one you
and Widow Douglas made for my cotillion? Why'd you do it, Hucky?''

It was quiet on the road, except for the birds singing.

``What did you do with the cake?'' he says. ``Did you eat it?''

I stood up. ``I ate it,'' I says. ``I ate all of it. And it was the best
dad-blamed cake in the world.''

I stalked away down the road, and Tom didn't follow.

I had crept back up the hill around midnight, after the widow was in bed
and all the lights was out, and gone in by a window, and the cake was
right in the cellar where we left it.

Tom wouldn't understand. I hadn't had anything to eat yesterday, nor the
night before, and when I got back from the widow's, my trot lines was
empty. How could someone like Tom Sawyer understand that?

And I didn't feel bad about it, neither. \emph{We} made it, the widow
and me, and I didn't want a load of boys and girls gobbling it like it
was some ordinary cake.

I warn't even going to stick around that evening, but just by chance I
found myself near the Temperance Inn around seven, and light streamed
out of every window golden, and there was music. Somebody must a brung a
fiddle and a joos-harp and even a melodeon. A crowd of grown-up people
milled about outside, watching the parade of party-goers file through
the doors, and talking to girls and boys they knowed. There was a fair
mass of young folks going in. A couple big boys guarded the door.

I sloped up to the door with my hands in my pockets and my pipe in the
corner of my mouth and tried to pass them, and as I expected they
stopped me. One says, ``No way is some river trash coming in here,'' and
they called me wastrel and souse and idle slouch, and some other names,
too. Then one put out his hand and shoved me.

That was just what I was looking for. I struck out and hit him on the
nose, and the second one grabbed me while the first held his fingers
against the blood. I ducked out of the second's hold and butted him in
the gut, and then the first wrapped his arms around me and held me while
the second whaled on me, over and over, till I pulled myself upright and
snapped my head back, hard, against the first's face. That got him to
let go quick. Then I leaped on the second.

It was a bully fight; I enjoyed it immensely. I got my licks in, and
they got theirs too, but I reckon they got the worst of it. And just
when some other boys was jumping in to join the fun, the constable
showed up.

They all backed off at once, shamefaced. The constable took in the scene
coolly. He observed the well-dressed boys, tousled and scratched, one
with a bloody nose, and me like a chewed-up cat right in the center.

``Finn. I should have known,'' he says. I wiped some blood out of my eye
and tried to stand. My vision was blurry, and I stumbled against the Inn
like I was drunk.

``Like father, like son,'' the constable said. He put up his hands and
ushered everyone away.

My ears rang. I shook my head, trying to clear them. One of the big boys
grabbed my arm.

``Are you going to disappear, or am I going to have to---''

Right then Tom Sawyer came out of the Inn.

``Huck!'' he says, delighted. ``You came! Why are you hanging around out
here? Come on in!''

I jerked away from the boy, straightened my clothes, shoved my hands in
my pockets, and strolled through the door.

It was hot and noisy inside, filled to the gills with young folk.
Everyone was natty and prouder than a pack of pea-hens. They all had
masks on. Some was store-bought, but most was cloth wrapped around their
eyes with holes to see through. Some'd lined the eyeholes with sequins
or beads or lace, and looked real smart.

They\textquotesingle d take a quick looky at me and avert their gaze. No
one said a word to me. The music turned to dance music, and the boys and
girls started dancing. I slouched in a corner and watched.

It was loud; you could barely hear the music over the shouting and
feet-stamping. I saw tables of food, and went over to see what there was
to eat, and \emph{there was the cake.}

It looked the same as the cake me and the widow made, all covered with
purple sugar and lavender flowers, with red jam in the middle. Half of
it was sliced up, but the rest was whole, so there warn't no mistaking
it. I shook my head. I couldn't fathom it.

Then I imagined the widow baking a second cake this morning, on the day
of the party, harried and maybe hurt, wondering why I done what I done.
She made the same cake again, only this time she made it herself. And I
didn\textquotesingle t sift the flour for her.

I left.

I went to my mill and digged up a rusted tin can. Inside was the money
I'd earned selling logs that I catched on the river. I never saved
money, as a rule; money was for spending and enjoying. I learned that
lesson from Pap. When you lived with someone dedicated to their likker,
all cash walked out the door as soon as it walked in. So you better
spend it before it got spent for you.

But this was the money I reckoned on showing the constable if he tried
to send me to the poorhouse, so I could prove I warn't completely
penniless.

I reached in and scooped all of it. I didn't know how much flour and
sugar and eggs and milk cost, but I reckoned it was a heap.

I went up Cardiff Hill and stood outside her house. There was a small
light shining from an upstairs window, but the rest of the house was
dark.

I went in by the cellar window like I done the night before. I went up
to the kitchen. I was going to leave the money on the counter and be
off. But I guess I stood around a bit, looking at things and feeling
sorry for myself, because next thing I knowed the widow stood in the
doorway holding a lamp, in a red velvet dressing gown, with her golden
hair falling loose.

She looked at me. I looked at her.

``You didn't go to the party,'' she says. She sounded so disappointed.

``I went. Didn't care for it.''

She nodded. ``Huckleberry\ldots''

I jumped for the door, but it was locked and the key was in the lock,
and that slowed me down. She came up behind me and grabbed my hand,
which was still clutching the coins so hard they bit my palm. I tried to
wiggle free, but that widow had a grip of iron.

I released my grasp on the coins. They rained down on the stone floor
and made a horrible sound, like tiny bells. We stared at them down
there, rolling and bouncing and then lying still. It was quiet.

``Sorry,'' I whispered.

She still held my hand, but her grip was soft. ``If you're ever
hungry\ldots'' she says. ``If you ever find yourself so hungry you can't
stand it\ldots and you don't know what to do\ldots''

I let out a choked sound.

``Come here,'' she says.

She led me real gentle out of the kitchen and down the covered walk into
the house. She sat me down in the parlor, on the sofa, the beautiful one
of polished wood and clean blue velvet. She went out, and returned with
a cloth and bowl. She cleaned the bashes on my face, my split lip and
bloody nose and black eye and the cut on my cheek. I didn't know where
to look while she was doing it, so I looked at her. Her expression was
sad, but also sort of loving.

Didn't she know I was Huck Finn, river trash, wastrel, idle slouch, son
of the town drunk? What was the matter with her? She ought to know
better.

Then she brung me a mug of hot milk and some sliced rye bread with
butter. It was the best things I ever eat. And while I eat I started
talking, just running my mouth, telling her useless stuff she probably
didn't care to hear. About Tom Sawyer's party, and the fight.

About my Pap, and how he left one day last spring without a word and I
ain't seen him since, and maybe he was dead. How I hated him but also
longed for him, deep in my gut where it wrenched, and it was real
confusing for me.

About Tom Sawyer. How he was as crazy as a loon, and thoughtless and
careless and self-centered. How he was the darling of the town, like the
magi said, and boys and girls worshipped him. But even though he was
rolling in friends, though, for some reason he still wanted to make
friends with me. He was the only boy in town who did, and frankly, point
of fact was I loved him more than I ought.

I told her about lucky charms, how to cure a cold or banish a hex. About
the river in the summer and the river in the fall.

I disremember everything I talked about. But she listened, and laughed
and sighed in all the right places. She acted like she was interested.
Warn't no reason why she should be.

Deep in the night sometime I stopped talking because my throat was wore
out. The fire had burned to ashes. I was lying on the sofa, and she was
sitting next to me on the floor. At some point during my talky-talk, she
had reached up and grabbed my hand, and she was still holding it. She
squeezed my fingers.

``You sleep here tonight, Huckleberry,'' she says, ``and in the morning
I'll feed you breakfast.''

``Thanks,'' I says, all drowsy, and drifted to sleep. And for the first
time in a long while, I didn't feel lonesome. Not one jot.

The Widow Douglas is all right.

\mychapter{Chapter 11}\label{chapter-11}

I sat on the docks, dangling my feet in the water and lazying along, not
doing nothing---my favorite occupation---when Tom Sawyer skidded up
behind me, shoes clattering on the boards.

``Huck!''

I flopped on my back and looked up at him, all warm and chummy. ``Hi-ya,
Tom,'' I says. I waited for him to say something smart back, possibly
along the lines of ``Hi-ya yourself,'' since that was what I always said
to him. But he didn't. He breathed hard, his face red. I patted the
boards beside me.

He flopped down and brushed a curl of sweaty hair out of his eyes. Now
he was here, he didn't seem in a hurry to spit out his news. I didn't
care. It was a fine day on the river. The sun shined its little heart
out, the water sparkled, and the breeze smelled like brown leaves as it
tossed the branches of the trees. I didn't have no place to be nor
nothing to do. And Tom Sawyer had come to rustle me up. Life was fine.

I leaned over and got him in a headlock and rubbed his noggin with my
knuckles, wanting to wrestle. He pushed me off. I pounced on him and
pushed him down on the dock.

``Ow,'' he says, face mushed against the boards.

``Say `nuff,'' I says happily.

``Get off, you ninny.'' He struggled.

``Shan't. Not till you say--''

``I \emph{said}, quit it,'' he says, all snappish, and wrenched himself
out from under me. I stared, surprised. He was really upset.

``What's eating you?'' I peered into his face and smiled. ``Ain't so bad
as all that.''

``Shows how much you know,'' he muttered.

This was odd and not like Tom at all. I got a little shivery. ``What is
it? What's happened?''

He cut his eyes at me guiltily. He was fully reluctant to own it,
whatever it was. I didn't like the signs. ``Well, go on.''

He sighed. ``I don't want to tell you but you\textquotesingle re going
to find out eventually, and it might come better from me than from
someone else. Just\ldots swear you won't lose your nut when I do,
okay?''

I frowned. ``Swear.''

Tom took a deep breath. ``I'll say it then. Uncle Jake\ldots it's Uncle
Jake.''

I got a stab of ice to the gut. ``What about him?''

Tom looked miserable. ``He got done near to death last night.''

I blinked. My brain refused to absorb this. ``Say again.''

``He got---''

``What do you mean, `done near to death?'\,'' I says in a careful voice.
``Explain it real slow, so I can understand.''

``Someone broke into his little house last night and beat him up within
an inch of his life.''

Words failed me. It felt like the air was heavy, the sky black.

I forced myself to suck some breath. ``He all right?''

Tom understood what I was asking. ``He'll make it.''

I felt like I was choking. I stood. ``Who done it?''

``No one knows.'' His words tumbled out fast as he scrambled to his
feet. ``But\ldots I think I do.''

``Tell me.''

``The magi and the gin are back,'' he says. ``I saw `em outside the
Temperence Inn, watching the boys and girls go into the party.''

``I'll kill them.'' I turned to go and do it.

Tom grabbed me. ``No, Hucky. Don't do that.'' I twisted my arm loose and
he got in front of me and put his hands on my chest, holding me back. I
shoved him aside, and he put his arms around me, pinning me. I threw up
my arms, breaking his hold easy.

He slapped my face. I struck him a blow with my fist which landed
squarely on his eye and sent him reeling.

He thumped to the dock and sat staring blankly up at me, his mouth an
``O'', his face the picture of shock. His fingers crept to his eye,
feeling it.

My vision welled up. Before I could say nothing, he was on his feet and
slinging an arm around my shoulders.

``I know,'' he says, resting his head against mine. ``I know, Hucky. We
won't let them get away with it. But we don't want to kill `em, right?
That won't do us no good, right, getting throwed away in prison for the
rest of our lives? You wouldn't like that. No, we got to be smart. We
got brains. Let's use them.''

I nodded and gave a big shuddering sniff. I wanted to make those
creepy-crawlies pay.

``Let's sit and we'll cipher out what to do,'' Tom says, and led me back
to the edge of the dock.

We talked for two hours and came up with The Plan.

Next day I went out to Rogers's farm and waited in the woods. It was
horrible uncomfortable waiting, knowing what I was waiting for. My heart
thumped against my chest like a fist hammering at a door. I didn't like
it. But I done it.

Finally I heard light, skippy-skip footsteps approaching through the
woods, and hid myself in a bush. Into the clearing came the magi,
trotting along with Tom Sawyer beside him.

``Here we are,'' says Tom.

``You're a good lad,'' says the magi, talking to Tom like he was a baby.
He rubbed Tom's head and patted his cheek and tweaked his nose and
pinched his chin, all quickity-quick. Tom smiled and took a careful step
away.

``Now, where's your lovely little friend---Reynard, you said? It was he
who took my very, very important document at the Temperance Inn?''

``That's right.''

The magi smiled craftily and leaned towards Tom. ``There was two boys
that night. Perhaps the other was you?''

``Don't recall,'' says Tom, stuffing his hands in his pockets and
looking away.

The magi bent his curved spine even deeper to peer right into Tom's
eyes. ``Are you \emph{lying}?''

Tom shuddered. The magi's arm jerked like a marionette's, and Tom
flinched. But the magi just patted his face.

``Charming,'' the magi says. ``Now, my paper. Please.''

``Reynard has it. He'll be along.''

I popped out of the bush. ``Here I am.''

``Ah-ha,'' The magi says, pointing a long, bony finger. ``I knew it. So
you're `Reynard' now, my dear? Last time we met, you were Joe Harper.''

He grinned. ``So you took my paper.''

He lurched towards me with frightening speed, elbows jutting, hands
hooked into claws to grab me.

I skeddadled backwards fast and took cover behind a tree. ``Hey, hey,
hey,'' I says. ``It was an innocent misunderstanding. I didn't mean no
harm. I thought it was money. I'll fetch it for you, and we'll go our
separate ways.''

``Fetch it?'' he says. ``I thought you had it.''

``I did, but I fell down a well that night and dropped the paper at the
bottom, and it's there now. See, here's the well if you don't believe
me.''

I led the way to it.

The magi approached cautiously, as if expecting something to pop out at
him. It was still broke open; they hadn't fixed the rot-up cover. He
peered down. ``Well, go on, get it.''

I made a great shudder. ``Not me,'' I says. ``I couldn't go down again
in a million years. I spent the worst night of my life in that well.''

``Then your little friend will go down,'' says the magi. He looked
around the clearing for Tom, but Tom had vanished. ``Where is he?''

``He was late for something.''

``How will we get it out?'' The magi was practically gnashing his yellow
teeth.

I rubbed my skull, making a show of ciphering it out.

``I suppose I could lower you down,'' I says. ``I got rope.''

He peered down in the well again, for a long time, and I knowed we had
him. I don't know why, but he wanted that old piece of rubbage about the
FABULOUS Fabrizio bad.

``Is it dark down there?'' he says.

``You can see light at the top.''

``Is it damp?''

``Dry as a camel's backside.''

He rubbed his chin. ``I'll go down,'' he says, ``but you shan't lower
me. I'll tie the rope to a tree.''

``Good as wheat,'' I says.

He went down, muttering about narrow sides and damp walls and slime. I
heard a splash when he landed.

``My darling, clever lad,'' he says, all echoey, ``you were right about
it not being damp. It's flooded.''

Quick as a fox, I yanked the rope out of his reach.

``Listen, you,'' I says down. ``Me and Tom Sawyer---''

``Here,'' says Tom, skipping over---

``We say you got to control your gin from now on. No more trouble around
town. Got that?''

``My \emph{what?''} says the magi.

``The red fellow,'' says Tom, ``We know he's a gin, and that he's the
one keeps causing trouble for innocent folks.''

The magi was silent. I heard him splashing quietly.

``Swear you'll keep him under control,'' I says.

``If I swear, you'll let me out? And give me my paper with the red
seal?''

``Cross our hearts, hope to die, stick a needle in your eye,'' says Tom.
``But you also got to admit you was responsible for the beating of an
innocent man last night.''

He was quiet. ``It's just you boys up there?'' he says.

``Yes,'' I says.

``Then I guess I admit I know something about it,'' he says. ``Now let
down the rope.''

``Let down the rope, Huck,'' says Tom. I winced, wondering if the magi
heard him use my name. I uncoiled the rope and dropped it down the well.
We listened to him grunting as he pulled himself up.

He clambered out, filthy and red in the face, like a big ugly spider. We
stayed out of reach, ready to light out if he made a move for us. But he
was too tuckered out. He curled up on the ground and wheezed.

And then the constable stepped out from behind a tree and laid hands on
him and slapped him in irons.

``I\textquotesingle m arresting you for conspiracy in the beating of a
man last night,'' says the constable.

``Lies!'' the magi yelled.

``I heard you admit it,'' says the constable.

The magi's eyes bulged. ``I was set up!'' He glared at Tom. ``That
little---he tricked me.''

``Where's your friend in red?'' the constable says. ``We'd like a few
words with him as well.''

The magi bared his teeth. ``Ain't telling.''

He whipped his head to Tom. ``You got to give me the paper,'' he says,
urgent-like. ``We had an agreement.''

``Oh, that?'' says Tom carelessly. ``Lost it ages ago. May have used it
to make spitballs. I disremember.''

The magi howled as if in pain. As the constable led him away, he caught
my gaze and his eyes narrowed with hate. ``I asked if you was alone up
here, and you said yes. You lied to me.''

``You beat up Uncle Jake,'' I says. And I gave him the rudest gesture I
knowed.

\mychapter{Chapter 12}\label{chapter-12}

I spent a lot of time visiting Uncle Jake the next few days, helping him
out, bringing him his vittles, because the way the people in the house
acted, you'd think they expected him to get up and go to work like
regular, like nothing was the matter. It disgusted me.

I was real glad we got the magi behind bars before anyone else got hurt.

``I only wish we'd managed to capture the gin as well,'' I had said to
Tom as we walked back to town after The Glorious Plan.

``I didn\textquotesingle t see him anywhere when I fetched the magi,''
Tom said. ``Just as well; I doubt we could a fitted them both down the
well. But don't worry,'' he added. ``A gin ain't worth salt without its
master. If we got one behind bars, the other is powerless.''

I warn't too sure about that. But I wanted to believe it.

``Sorry about your eye,'' I said. It had already swelled to a
magnificent shining purple where I clocked him. He grinned.

``Never mind. It's a beaut.'' He stroked it lovingly. ``Might even get
out of school for it. Can't read my lesson-book with a shut-up eye, can
I?''

And that was the last we'd see of the gin and the magi for ever, we
hoped.

About a week later I cut my knee with my knife doing I disremember what.
Splitting a sapling to make a snare, like as not.

It didn't bleed very much, so I took my hands and squeezed it, to make
it bleed, because any saphead can tell you there ain't nothing better'n
your own blood for flushing out a wound. Except maybe likker, but I'm
not too sure about that.

I wrapped it with some burlap and thought no more about it. But the next
day, when I waked up, the whole knee had swelled up big as a melon, and
stiff. The cut was about two inches long and edged in white, fiery in
the middle. It looked pretty ornery.

I didn't have no likker, so I sat in the river a few hours and let the
water flow over it. My knee was hot, so the cold water felt nice on
\emph{it}, but otherwise it was pretty uncomfortable.

The day after, all my joints was painful achy. It was the derndest
thing; I never saw the like. All my elbows and wrists and fingers and
toes hurt, even my neck, till I couldn't hardly move. I couldn't make it
into town. I could barely hobble down to the water to check my lines. I
didn't want to close my fingers, didn't want to even lift my arm nor
turn my head. So I just slithered down to the river and laid out there
in the mud and weeds and leaves, and did absolutely nothing for two
whole days and nights.

I used my overcoat as a cover. It was cool, even at noon.

The reason I crawled down to the river like that was\ldots well, it's
uncomfortable to admit, now I know I lived. But the shameful truth is
that if I was going to kick daisies, I didn't want to do it inside the
mill, and be all smelly and humiliated when they found me. I druther go
on the riverbank. I'd just sort of fade into the grass, and plants and
vines and trees and flowers would grow up around me, and no one would
ever have occasion to look at my dead self. It would be real private.

Them's just the sorts of things you think about when you judge you're
fixing to croak. Like I said, pretty embarrassing, now I lived.

One day I waked up and found I could move a bit, so I ventured down the
riverbank a ways, feeling about a million years old, sort of hobbling
along and seeing what was what. I loosened as I moved. I found a nest of
mud-turtle eggs and filled my pockets, and went back to my mill and
fried them up. But I could've eat about a thousand before I was
satisfied.

I made the trek to town, which felt about fifty miles and uphill all the
way. I straggled into back alleys looking for scraps. It was too cold
for anyone to leave baking on their windowsills, so I couldn't get
anything that way.

I went by Tom Sawyer's house and entered by the gate, because jumping
the fence was out of my reach. I looked in at the windows, but the house
was empty. Where was Tom and his aunty and cousin and little brother?

They was on Jackson Island, having a pic-nic. I could see them over
there across the river, jolly on their blanket, and couldn't figure how
to get there other'n to swim. So I done it, and it was cold.

I crept up in the brush, back of where they had their spread set up, and
made a soft \emph{tr-r-r-rill}, like a bird. They was talking and
carrying on so they didn't hear.

I done it again, and Mary looked over, so I ducked down, and when she
looked away I hove a pebble at Tom Sawyer's head. \emph{That} got his
attention.

``There's a bird over there a-ways I want to take a look at,'' he says.

``Tell him to join us,'' says Mary.

``What?'' says Aunt Polly, as Tom scampered over.

``Huck,'' he says when we crept a ways into the forest. ``Ain't seen you
around in an age. How you been?''

I showed him my knee, and he was suitably impressed. We talked a bit,
and I asked him why he was having a pic-nic, and he said it was Sid's
birthday anniversary. Then he asked if I wanted to meet him and Joe
Harper in the public square later to play something, and I was agreed.
He went back to his folks.

I watched them for a while, secret-like, from the brush. ``Did you
invite him to eat?'' says Mary.

``Who?'' says Aunt Polly.

``The bird,'' says Mary, with a wink for Tom. Tom said never mind, real
hasty. I knowed he didn't want Aunt Polly to know it was me.

They eat the most sumptuous truck you ever imagined---chicken, and
pickled pumpkin, and spiced pears, and corn dodgers, and roasted squash
with carrots, and pie, and cider. I nearly passed away from the smell
alone.

I swum back to shore, and by that point I was too hungry to look for
food, so I found my favorite doorstep that was warm in the sun and
sheltered from the wind, and slept till afternoon.

``Huck.'' I felt a touch on my shoulder, and opened my eyes. It was Tom.

``Ready?'' he says. I nodded and gapped and stretched. The sun had
moved, so it was cold where I was sitting. I wrapped my coat round
myself and snuggled into it.

``What're we playing?'' I says. He motioned me to follow.

``We're playing musketeers,'' he says as we walked to the square. Joe
Harper was waiting for us, along with Billy Fisher and Little James.

``Again with the musketeers?''

``It's more fun with five,'' says Tom. So he reassigned roles, and I was
some bloke called Athoes. Tom was Dartanyen again, Billy was Port-horse,
and Joe was Aramiss. Little James was My Lady, which he was sore about
till Tom swore it was the most vicious character, and did all sorts of
poisonings and stabbings. So James was satisfied.

``How come we're musketeers if we ain't got no muskets?'' says Joe, and
I thought this was a swell point. Tom said that's just how it is,
musketeers use swords, and swords is more fun than muskets any day, and
any chucklehead who thinks otherwise can get hanged. So we played
swords.

We stuck pheasant feathers in our hats. We took off our galluses and
untucked our shirts and tied the waists with rope. The boys unfastened
their shoes and pulled the tongues out so they looked like fancy French
ones, but I warn't wearing shoes. Tom said I could just let on I was
wearing shoes and let on they was French. So I done it.

My character, Athoes, turned out to be the worst. I was an old man
turned to drink from my sorrows.

I says, ``Fine, I'm old, and fine, I got sorrows, but I ain't no
drunk.''

Joe says, ``Don't be ornery, Huck Finn. Just play the part.''

I crossed my arms and glowered.

``If you're not desperate for the soporific solace of your wine, we
ain't got a tragic story, we just got a bunch of nuts chopping at each
other,'' says Tom. ``It's really the most important part.''

``Then you be crummy Athoes.''

``I got to be Dartanyen,'' he says. ``I'm the only one knows how to do
it.''

Billy looked round, making certain we was alone, and went to his
roundabout and pulled out a little flask, and he winked at us and made
us come close.

``You can use this, Huckleberry,'' he says. And he opened it and held it
out, and Tom sniffed it and wiggled with glee, and took a tiny sip, and
started hacking up a lung and his face got red and his eyes watery. He
gave it to Joe, who tried it with similar results. Joe passed it to
Little James, who passed it to me. I sniffed it and nearly retched. I
would a done, too, if there had been anything in my stomach. It was
forty-rod.

Forty-rod whiskey is the cheapest, foulest truck you can find, and it's
called that because that's the distance from which it can make you
drunk. A rod is seventeen feet.

I put the stopper back and shoved it into Billy's hands. ``I'm done
playing,'' I says, and untied the rope from my waist and went over to my
coat and gallus.

The boys set to wailing. They said I would ruin the game.

``I don't want to play with no numskulls anyways,'' I says, and I called
them pinheads and mugginses and all sorts of names. Billy tried to
snatch my coat, but I snatched it back, and he slapped me, so I
demonstrated how to use a fist, and my lesson sent him reeling. I
jeered, and he jumped to his feet and lunged at me.

Tom and Little James grabbed me, and Joe grabbed Billy. Everyone shouted
about how I went too far, and I said I disagreed, I went exactly the
right amount of far. Then next thing I knowed all the blood rushed to my
head, and the ground came up to me.

When I opened my eyes, the boys was gathered round looking at me.

``And he didn't even try the forty-rod,'' Little James says in a
whisper.

I shoved them all back. ``Get out of my face.''

Tom grabbed my arm and steadied me, and pulled me to my feet. I felt
like there was gas in my head.

``Huck,'' says Tom, real quiet. ``When did you last eat?''

I shrugged. ``I had some truck this morning,'' thinking of the eggs and
the scraps I found behind the shops. But I didn't tell him about laying
out by the river for two days. Maybe I oughtn't to a come to town. In
hindsight, I probably should a stayed by the river and fished.

The boys made a fuss, trying to put damp handkerchiefs to my head and
fetch me water, until I wanted to wring their necks. I would a done,
too, except Tom read my mind and shooed them off sharpish.

``Come on,'' he says, helping me up.

``Where are we going?'' I felt cross and grouchy.

``Just come on.'' He steered me off.

``I\textquotesingle m fine,'' I says, trying to pull away, because he was
clinging close to my side.

``I know. I\textquotesingle m just cold,'' says Tom. ``Some fellows might
be fortunate enough to have a great big overcoat, but all
I\textquotesingle ve got\textquotesingle s this tiny little roundabout.
Have mercy, for pity\textquotesingle s sake.''

I still felt like fighting a bit, but on the other hand I
didn\textquotesingle t want anyone catching ague. After giving the
matter some thought I wrapped my coat around him, and he snuggled close
with his arms around me, and I buttoned us up.

We looked like a big fat man with two tiny heads.

We peeked at each other over the collar and bust out laughing. It was
the most hilarious thing you ever seen.

We moseyed along through town, trying to tangle each
other\textquotesingle s legs. We each claimed a sleeve. I doffed
Tom\textquotesingle s cap at a lady. He thumbed my nose at her. She
glared at me. That set us off again.

Tom led the way to the general store. He opened the door with his arm,
and I used my arm to gesture him politely in, which looked so dern
hysterical. We tried to go in at the same time and got stuck.

Finaly he led me to the counter where all the candies was at.

``Pick whatever you want,'' says Tom, all grand.

I stared at the sugary bits, displayed in jars and baskets lined with
paper.

They had stain glass, which was sharp colored shards flavored with
lavender or rose or lemon. They had lickrish, and crystalline ginger,
and sugarplums, and peppermint sticks. They had brown horehound drops,
and red candies filled with strawberry, and chewing gum. It was too much
to absorb.

Actually, the sight of all that sugary stuff made me kind of ill. I
would a killed for a slice of ham or a biscuit.

I looked at Tom, and his eyes shone as he watched me. I couldn't refuse.
``You pick for me.''

So Tom bought a penny bag of candy, and when we got outside he selected
the biggest peppermint stick and gave it to me. I tried it and said it
was swell, and Tom was happy. We parted ways, Tom skipping for joy.

The peppermint did me good; it settled my stomach and steadied my head.
I moseyed along through town, nibbling and thinking.

Fall was a nice time for St. Petersburg. Summer was my favorite. But if
anyone asked at that moment, as my feet swished through the crackling
leaves and the breeze sifted my hair, I would a sworn my favorite was
fall, always was, always would be.

I remembered how funny Tom and me looked as a two-headed man. I grinned
like a fool.

Tom Sawyer was something else. They really breaked the mold when they
made that one. He was perpetually filled to bursting with hare-brained
schemes and nutty games and all sorts of nonsense he read or dreamed up.
Sometimes he was the most infuriating little goose that ever lived. No
one got my hackles up like he could.

But he also made me feel warm and lit-up sometimes, like my chest was a
lantern with the flame turned all the way up. No one else in the world
made me feel like that.

My overcoat felt too big with just me in it.

I remembered how Tom clicked his heels as he scampered away, and I
smiled.

Tom Sawyer's all right.

\mychapter{Chapter 13}\label{chapter-13}

Fall ended with a storm. We had one last warm day, balmy and still as a
dead dog. Not a breath of wind anywhere. That warn't so pleasant.
Everyone laid around on their porches and front steps and sweated.

But \emph{then}---that afternoon---the wind blowed in, bringing the
thunderheads. They was these huge, solid, dark clouds, like toad holes
for angels, that towered over town, massive.

It was positively delicious, because the hot air would suddenly get cut
through with cold, fresh air that smelled like rain. Hot and cold, side
by side, like magic. Then the two airs mixed together and tore the sky
to flinders.

Trees roared; their leaves tossed and strained and ripped loose. And
leaves warn't all that was in the air; just about everything you could
think of was blowing about. I could a reached my hand and snatched a
bible, or a top hat, or a kid glove, or anything I wanted.

Everything was moving and alive and not just alive, but \emph{violent}
in its enthusiasm for life.

I flew into town ahead of the storm, whooping and hollering. My big
shirt and pants and overcoat---all mens' things I had found in the
rubbage ages ago---provided extra sail or extra drag, depending on if
the wind was for me. Sometimes the cloth \emph{hurt} as it whipped
against my skin.

I lost my hat; no consequence.

It was the most alive I ever felt. All of me thrilled from my fingers to
my toes. \emph{Oh---}THIS was to be alive!---how glorious life was!

Hot and cold pummeled me as I jumped onto the water pump, balanced on
one foot, tumbled off and rolled. I swung from the beams of porches and
climbed stairs just to leap down them six at a time, and in general
acted like an absolute tom o'bedlam.

Every which way folks ducked into houses and shops and took cover,
because the rain was coming---you could see it, like a solid wall,
advancing from the east. I hopped onto the little stage in the square,
the one where I'd got tanned in the summertime, and spun in circles with
my arms flinged wide, whooping like a lunatic; I ripped off my coat and
released it to the wind, and it flapped away like a great dark bat.

I kept my arms raised and my face turned up as the rain hit---and it
\emph{did} hit---like a bull. I laughed and screamed and skipped about,
kicking water with my bare feet. It was \emph{so cold} it took my
breath.

Everything had been flat and dull and hot and dead, and now it was cold
and alive.

A brewing storm is the most fabulous fun.

\mychapter{Chapter 14}\label{chapter-14}

With the arrival of winter meant, for Tom Sawyer, the arrival of
CHRISTMAS. He was so jumped up about it he couldn't hardly stand it.

We got an inch of snow one night, so next morning me and Tom went to the
town square and Tom wrote bad words walking heel-to-toe. I gave
suggestions and bossed him about. It was so funny, we was rolling. Then
the lawmen came and ordered us to rub it out, and we took off, so they
got brooms and did it theirselves. But it was still funny.

A peddler came with his cart, with all sorts of truck for sale. The
ladies visited first and got their pots and pans mended, then other
fellows went to buy nails and horseshoes, and then finally the young
folk had the peddler all to theirselves, and bought anything they could
afford with whatever they could persuade him to take for it.

Tom wanted presents for Mary and Sid and his aunty, and I tagged along
for fun.

``What shall I buy for Aunt Polly?'' says Tom, looking at his reflection
in a shiny pot. ``I don't want any of these ordinary things which she
would buy for herself, and which she would only use for cooking or some
other work.''

``Clever lad,'' says the man. ``How about sharp sewing needles, six a
penny?''

``Ain't you listening? She would use those for \emph{work}. Anyhow she's
always complaining about losing needles, and it really vexes her. I'd
hate to get something that would vex her.''

``What about this?'' I says, lifting a copper ladle. Tom gave me such a
withering look I put it right back.

``How about this?'' says the peddler. He held up a copper bird on a
chain.

``Does it do anything?'' says Tom suspiciously.

``Steeps tea,'' says the man, reluctant to own it.

``If you don't tell her, maybe she won't guess,'' I says. ``It looks
proper useless unless you \emph{know} it does something.''

Tom considered this and bought it. He also bought some jacks for Sid and
the needles for Mary. ``Because,'' as he said, ``Mary don't mind losing
them, because she never \emph{has} to my knowledge lost one in her life.
She has about twenty of them, lined up in a row in her notions box.''

So that was settled. Then we got to jawing with the man, because a
peddler is always interesting and full of stories. He told us this bully
tale about how the folks down in Texas raised up a few years ago against
Mexico and became independent. It was a real engaging piece of drama. I
doubted any of it was true.

I peeked around the inside of his wagon while I listened. It was real
cozy. Like a little house, with everything a fellow could want: a
miniature bed that folded into a table, and a minute kitchen with a pot
and skillet and a pronged iron stand for cooking over a fire. There was
even a little whittled chair, with a fiddle resting on the seat.

``I like your wagon,'' I says to the peddler. ``Must be nice, roaming
the roads, seeing new sights, free to move along anytime you please.''

``Always there, never here,'' says the peddler. ``It's a rough life, but
a free one. A man is alone with his thoughts.''

``Don't you get lonesome?''

``A peddler has friends in every town. Family with every family.''

``Must be hard work,'' I says, ``mending pots and selling things.''

``Life is work,'' he says. ``The righteous man toils till he dies. The
harder he works, the more he makes of himself.''

``What do you do once you're made up?'' I says.

``You can always make more,'' says the peddler.

Tom listened to our talk with a frown.

``I don't reckon I like the sound of all that work,'' I says. ``Sounds
an awful lot like a poorhouse.''

The peddling man looked at me. ``Now what does an upright gentleman such
as yourself know about the poorhouse?''

``I'm going to get sent to one,'' I says. ``If the constable proves I
ain't got no pap and no means.''

``That's a shame. I had a brother in a poorhouse. It like to killed him.
\emph{Like to},'' he says hastily, seeing my face. ``He got removed to a
poor farm.''

``Poor \emph{farm}? What's that?''

``It's a place where lots of folks work, folks that can't make a living
nowhere else. It's like a big family. You live together, work together,
eat the fruits of your labor seated around a big table. You spend every
day outdoors, working the land as the Good One intended.''

I absorbed this. It didn't sound quite so bad as a poor\emph{house}.
Leastways, I liked certain bits of it.

Tom watched me mull it over, his countenance darkening.

``Maybe that don't sound so bad,'' I says.

``Breaking your back in the sun, all day every day,'' says Tom. He
scowled and shoved his hands in his pockets. ``Covered in blisters and
sores. Bone tired every night.''

``Nay, nay---'' says the peddler.

``It don't sound like anything worth notice to me. Come on, Huck,'' says
Tom, hooking my arm in his. ``Let's go wrap my presents for tomorrow.''

``Take care now,'' says the peddler.

``What's eating you?'' I says as we walked back along the road.

``Nothing,'' Tom says. ``I just didn't like all that jibber-jabber about
poor farms. Too derned depressing. It's Christmas Eve, for pity's sake.
Let's talk about something else.''

``I thought it sounded kind of passable,'' I says. ``I wouldn't mind
going off and seeing new places. I like growing things. I'd be allowed
to spend all my time outdoors like I do now. And I'd meet a lot of other
fellows who would become my friends and family. Sounds right cheery to
me.''

``That's because you're a half-wit, Huck Finn,'' Tom says, rounding on
me. To my surprise, his face was beet-red. His eyes glittered brightly.
``You don\textquotesingle t know nothing about nothing.
You\textquotesingle re a simpleton.''

He shoved me in the chest a few times, like he was trying to start a
fight. ``A simpleton. An \emph{innocent}. That\textquotesingle s what you
are.''

I didn't lift a finger. I was too astonished.

He threw up his hands. ``You know what? Suit yerself. Run off to some
forsaken farm at the farthest reaches of goodness-knows-where. Break
your back. See if I care. Just---just---don't bring back any of your new
\emph{friends and family}, if you return for a visit. Actually, don't
visit. I don't reckon I'd want to see you.''

He scrubbed his eyes with the back of his hand, then took off towards
his house. He hopped the gate and went in at the front door, slamming it
shut.

I was bowled. I couldn't fathom what it meant. But I wished he hadn't
gotten sore. It made me feel lousy, and I didn't understand why.

So much for Christmas Eve. And the next day was CHRISTMAS.

\mychapter{Chapter 15}\label{chapter-15}

Christmas is some huge big deal in St. Petersburg. Maybe it is
everywhere; I don't know.

First thing was, everybody went to church all morning. The town was
silent as the grave while you waited for this bit to be over, and it
could get pretty tedious. But don\textquotesingle t let that fool you,
because you'd wish for the quiet once church let out.

The people flooded out late, it being an extra long preachy-preach. Then
the fun began.

First we built a great bonfire, and set up to roast a cow. It would cook
all day. The smell was like the heady winds of paradise.

Everyone sat around jawing and waiting. Someone started singing
``\emph{Heav'n and Nature Sing,''} and it was took up. I didn't know the
words, but there was a lot of repeats in it, and if I felt like it,
which I'm not saying I did, I reckon I could a picked up a few choruses
and sung along, which I'm not saying is what happened.

Then I hunted up the boys and found them in the schoolyard, trading the
loot from their socks.

``I'll give you this peppermint for your wood horse,'' says Tom to Joe
Harper. Joe was rightly indignant and demanded Tom's new penny
\emph{and} the peppermint. Tom refused, and traded his peppermint to
Johnny Miller for a couple walnuts.

Johnny Miller traded Billy Fisher a slingshot for a brand new taw
marble, so I asked Johnny for his \emph{old} taw marble, which I still
wanted from when I tried to trade him for it a while back. But he said
he wanted to have two, and anyways I didn't have nothing to trade. So I
said, yes I did, I had a pebble from a fish's stomach, and this set Tom
wild.

``Are you expecting me to believe you found \emph{two} pebbles in a
fish's stomach?'' he says. ``Did that fish eat a lot of pebbles, you
think? Or was it two different fish? I suppose this means that pebble
you gave me was a piece of rubbage.''

``No, \emph{yours} was real,'' I says, and that made \emph{Johnny} wild.

``Oh, blame it and forget it,'' I says, and stayed out of the trading
after that.

One thing the boys had in their socks was these new candies shaped like
oysters. When you bit them they broke apart, and there was a slip of
colored paper secreted inside, and printed on it was a message. The boys
was wild about these, though the messages was strictly bilge.

They didn't trade these new candies; they was too valuable. But they
agreed to eat them at the same time and read the sentiments aloud. We
sat in a circle.

Tom shoved someone over and flopped down cross-legged next to me. He
didn't look at me. I suspected he was still sore about whatever happened
between us yesterday. But he dug into his sock and pulled out a shell
candy and balanced it on my knee.

``Merry Christmas,'' Tom says. He met my eyes shyly and flashed a smile.
I grinned and bumped his shoulder with mine, and he bumped back. I got
warm and wiggly inside, and I suddenly understood why everybody made
such a flaming big deal about Christmas. It was terrific.

The candies was neatly shaped and delicate shades of pink or green or
blue. They was average tasting; no one cared what they tasted like. The
messages was the thing.

Ben Rogers read his first. ``\textquotesingle Married in satin, love
will not be lasting,'\,'' he says. The boys hooted.

Tom Sawyer read, ``A friend to all is a friend to none,'' and the boys
nodded sagely.

Billy Fisher read, ``Children are certain cares but uncertain
comforts,'' and we said how appropriate.

Johnny Miller's candy told him, ``Gluttony kills more than the sword.''
They whooped and teased and poked his belly till he got red in the face
and allowed he would lick them if they didn't leave off.

I bit mine and pulled out the paper. I handed it to Tom.

``Drink is the ruin of man,'' Tom read out.

I felt a giggle rise up in my throat. I tried to quash it down; it
wouldn't. I fell over and rolled. I laughed till tears poured down my
cheeks.

The boys studied their papers and nibbled their candies and didn't say
nothing. Guess they didn't get it.

Afterwards Tom invited everyone back to his house. We skipped and
horseplayed as we went. Before we entered Tom's gate he drew us up,
evidently studying something out. He flung out an arm, pointed across
the street, and sang out, ``Willie Mufferson!''

The boy in question stopped short, where he had been slinking along
inconspicuous as possible. He looked about, then gave a cautious nod.

``Come here,'' says Tom, pointing to the ground at his feet.

Willie came. His fists was clenched and trembling, like he expected a
fight.

``Won't you come inside with us?'' says Tom, all courtesy. He bowed,
holding open the gate.

The boy was wary. He knew that Tom, and all the boys for that matter,
despised him for being a sanctimonious runny-nosed whistle-blower (which
he was).

``Well Tom Sawyer, thank you very---''

``Don't mention it. Look, stand here, next to Huckleberry,'' says Tom.
``No, not like that, more in front of him. What are you, scared? He
won't bite. There, that's good. Now, no matter what happens, just stick
close to him, and if my aunt looks towards you, step in front of him and
look her in the eye and smile and say some polite rot, as you do. You
know how you do it.''

In tight formation we proceeded into the house.

The inside was decked with loads of green truck brung in, holly and
laurel and garland and mistletoe, which was the only things still green
that time of year. Mary and Sid sat round the tree, singing
``\emph{Nowell}'' and giggling. Mary was in a dark red frock with
embroidered white bits at the bottom, and Sid trussed to the chin in
blue velvet, with a wide collar. His flaxen hair was combed flat to his
head.

``Budge,'' says Tom to them.

Sid set to wailing. ``Now, Tom,'' says Mary, all gentle.

``Oh, all right then, stay,'' says Tom.

We gathered round the tree. It felt like we was in a Christmas card
picture, it was such a pretty scene. The tree was big, almost two and a
half feet, set up on a side table. It was done up with candles and they
dripped long tendrils of grease down the branches, and there was candied
fruits hanged on there, and bits of ribbon.

Everyone had shining faces and rosy cheeks and they was all smiling and
singing, and they looked like they just \emph{fit}, if you know what I
mean. And there was I, sort of slouched in my rags, my toes cold where
they stuck out the hole in my worn-out shoes.

Aunt Polly came in and saw me right off, and her eyes narrowed, and her
mouth twisted. Before she could say nothing, I grabbed Peter, her yaller
cat, and put him on my lap and stroked him. I kept my face downturned,
petting and petting. My hands shook a little.

``Good boy, Peter. Good boy,'' I says in a whisper. He looked into my
face and squinched his eyes and smiled---I swear he did. I bent and
kissed his furry head.

When I finally dared look up, Aunt Polly was gone. She'd left without
making a row nor kicking me out. I breathed. Peter kneaded me with his
paws, purring his little cat heart out. Oh Lordy. I was in love.

We told scary stories and eat the candied fruit off the tree. I eat so
much I was almost sick.

We fought to tell the scariest story, and I nearly won with my story
about the time Pap got the delirium tremens when I was small and mistook
me for a pup, and tied me in a gunny sack and tried to sell me to a
trader on the river, only a steamboat came and breaked up pap's boat, so
I didn't get sold, though I did nearly get drownded.

But Tom won with his story about a man who saw three ghosts on Christmas
Eve. And it \emph{was} scary. I was glad I warn't that man.

We had great fun at Tom Sawyer's, and when we broke up late in the day I
moseyed back to town to check on the situation concerning the cow. It
was nearly cooked. While I waited, I watched folks, which was something
I loved to do.

There was a tiny baby clinging to her pap's hand, toddling along on fat
legs. The pap brung her to see some gent, who exclaimed with delight.

``Hel-\emph{lo,} ain't you the doll? How d\textquotesingle you do,'' he
says, and took her tiny mitt between his fingers and shook it gently.

Over to the left a man showed his closed fist to his boy, who tried to
prise it open. Chuckling, the man opened his fingers, but his hand was
empty. Then he showed his son t'other fist, and the boy forced that one
open, and there warn't nothing in it, either. The boy shouted in
outrage.

The man roared with laughter. Then he pulled a trinket from behind the
boy's ear, and the boy snatched it. He wrapped his arms round his pap's
neck and kissed him. My head hurt. Don't ask me why.

And I saw a dog dressed up as a man. It wore a suit and spectacles and
looked real smart. I liked that dog. I wanted to borrow it and take it
back to the mill with me. But I reckoned it wouldn't be too happy living
there; its fine clothes would get spoiled.

When the meat was done I gulped down a whole pile, then got some more
and carried it out of the village. I went to Ben Rogers's farm and
hunted up Uncle Jake. He was hived up in bed still, on account of being
nearly done to death a few weeks ago. He had mostly recovered, all
except his left leg, especially the knee.

``That's my baby,'' he says when he saw me in the doorway, holding the
grub. ``You so good to Uncle Jake.''

He sat up in bed and patted the covers. I eased down beside him and
handed him the food real careful, so nothing spilled.

``Ain't nothing,'' I says, tucking a kerchief round his neck. ``You
given me bits from your own supper a hundred times I guess. How's the
leg?''

He showed me the knee, which was blotchy, and I showed him the raw pink
scar on mine from my knife, and told him how it had swelled up, along
with my other joints, too. He said he ain't heard the like of it.

``I got a hole in my foot from a nail, once,'' he says. ``My jaw clamped
up and I couldn't talk for a month. Was just about your age.''

``How did you live,'' I says with reverence.

``Drank water through my teeth. And soup, real thin. It was wretched; I
like to have died.''

While he eat he talked to me and petted my hair and in general made me
feel exceptional---like I might not be completely without value,
leastways to him.

I\textquotesingle d had fun in Tom's aunty's parlor. But
I\textquotesingle d felt ugly and out of place. Here I could slouch all
I liked, laugh loud as I liked, talk rough as I pleased. Maybe me and
Uncle Jake didn't look as pretty as Tom and Mary and the rest, but the
feeling was there, I wager.

Uncle Jake finished his vittles, and he looked wore out. He asked if I
wanted to sleep in the hayloft, which Ben's father didn't mind
sometimes, but I wanted to go back and see more of Christmas. So I
tucked his holey quilt around him and said so long.

Back in town it was dark, and folks lit torches and bonfires. Some
fellows got the idea to put gunpowder in an anvil and light it off, and
Lordy I wish you could a heard the bang it made as it flew straight into
the air, maybe fifty feet. It balanced in the sky, weightless, for a
long beautiful moment. Then slowly it dropped. When it hit the ground it
sent a shock up my legs.

The boys soaked balls of rags in likker and lit them, and tossed them
back and forth. As they flew through the air they blazed brighter, like
cannonballs lobbed by Ol' Scratch himself. If you catched them real
quick and tossed them back, you didn't get burnt.

I got singed a little. I warn't always quick enough.

It was terrific fun.

A parade marched through, singing and carrying torches and candles and
lanterns. They sang ``\emph{God and Sinners Reconciled}'' at the tops of
their lungs.

Finally everyone straggled back to their homes in twos and threes,
gapping and rubbing their eyes, and Christmas was over.

I slinked into the woods and made my way back to my mill, bleary-eyed,
tired as a sack of melons but in a good way.

My door was ajar. I pushed it open, and on the floor a lantern burned,
and the lantern warn't mine. I saw in the shadows a glint of green. I
went over to investigate and it was the gin's bottle, sitting upright on
the floor, uncorked. I picked it up. It was half-full of something. I
lifted it to my nose and took a sniff, and gagged. Forty-rod whiskey.

``Merry Christmas, son,'' says the gin, stepping out of the shadows.

\mychapter{Chapter 16}\label{chapter-16}

``Pap,'' I says. I set down his forty-rod with care, because if there's
one thing I knowed, it was that I was allowed to drink it, but I warn't
allowed to spill it.

He unwound all that filthy red cloth from himself. It appeared to be a
whole bolt of the cheapest fabric he could find. He dropped it on the
floor and stepped out. The regular clothes he wore underneath warn't
cleaner\emph{,} I'll tell you that for free.

His black hair and beard looked like spider's legs, long and spiky, and
his skin was pale as a corpse. He was as I remembered.

He straightened and his head weaved side to side like a cobra, that
motion I'd seen the gin make a thousand times. That was him trying to
keep his balance.

His eyes focused on me with effort. ``Give your pap a hug.'' He smiled
real big, his beetle-black eyes crinklish. He opened his arms to me.

I hesitated, gauging his mood. He read as sentimental and teary-eyed.

I dove into his arms.

He swept me close and squeezed me, burrowing his face in my neck. ``My
boy,'' he says, near sobbing. ``My little Huckleberry.''

I clung to him. ``I missed you so much,'' I says, throat tight. ``Where
you been?''

He set me on my feet, and when he let go he kind of staggered.
Quickity-quick I took his paw and led him to the split-bottom chair.

``I made this chair,'' I says, helping him onto it. ``I fixed up this
mill. It was a wreck when I found it.''

``Good boy,'' he says, sniffing and rubbing his nose with the back of
his hand. I wrapped my arms round his wide, sloped shoulders and kissed
his cheek. When Pap was safe I couldn't get enough of him.

``You was the gin this whole time?'' I pulled a piece of cloth I used
for a handkerchief out of my pocket and dabbed his eyes. ``How come you
didn't say nothing?''

He took the handkerchief and blew his nose. ``Because that blamed Morley
is in town, watching like a hawk.'' He fell to cussing the constable,
drawing together all his choicest comparisons and descriptors, really
spreading himself. I always thought if Pap hadn't been a drunk, he might
a been a poet. Except I doubt a poet's ever heard some of the words Pap
knows.

I says, ``The constable watched me like a hawk, too.''

``Cause he knows we're cut from the same cloth,'' says Pap, and slapped
my back, hard enough to send me stumbling. ``My own boy. I saw when you
was whipped. I was there.''

``Yeah?'' I remembered the gin being there, with the magi. The memory
kind of soured my stomach. ``Why didn't you say nothing?''

He tossed my handkerchief over his shoulder and motioned for his bottle,
which I fetched. He took a swill.

``Because I was the one that scooped the till,'' says Pap.

I sucked my breath in.

``I figured it best to let it play out the way it done. They tanned a
Finn, and a Finn done it, so what difference does it make?''

``I was getting tanned in public. I could a used a pap,'' I whispered.

``Eh?''

I swallowed. Pap looked around my mill and sneered. ``Ain't much. A Finn
deserves better.''

His snuggly mood was wearing off, getting fast replaced by an uglier
one.

I tried to reconcile in my head that Pap and the gin were the same. I
thought back to everything the gin had done. Suddenly my face went cold.
``Did you beat up Uncle Jake?''

I dreaded the answer.

``Who?''

``The man beat near to death on Rogers's farm.''

The stream of swearing which followed chilled me to the core. You don't
want to know all he said about Uncle Jake. ``Tailed you to his hut a few
times. Didn't like him getting so friendly with my boy, so I moseyed in
and told him to keep off if he knowed what was good for him.''

I gaped. ``You did?''

That meant Uncle Jake knowed the whole time who beat him up, he knowed
it was my own pap. I'd visited him a dozen times since then, and he
hadn't said nothing about it, nor treated me any different.

I was mortified. ``What did he say?''

Pap snorted. ``Got real smart. Said if you wanted to hang around he
wouldn't stop you. That was when I taught him his lesson. Me and
Smith.''

``Smith? Who's that?''

``My pal I picked up in Goshen. Tall, bony fellow. Hunched as a
buzzard.''

``Oh---the magi. And when he got catched and turned over to the
constable, you let him take all the heat.''

Pap waved his hand. ``Don't concern yerself about him. He's free
already. You know that cell ain't even guarded?''

The magi was on the loose. I got goose pimples.

I says, ``Constable Morley's been pestering me all summer and fall and
winter about you. Why does he want you so bad?''

Pap took another swig. ``Morley's brother was constable in Goshen. I ran
into trouble thataways a while back, and the long and short of it is, I
shot him. Reckon now Morley wants revenge.''

I grabbed my skull like it hurt. ``You shot a man, Pap?''

He sneered. ``He was breathing when I left him.''

I struggled to breathe myself. ``Constable Morley hates me. He wants to
send me to a poorhouse.''

That set Pap off swearing again. ``Ain't no son of mine getting carted
off to no poorhouse. I got a plan.''

``Yeah?''

Pap tried to stand but caught his toe and went sprawling. I rushed to
help. He shoved me back with his arm.

``A plan to get us some cash,'' he continued like nothing happened.
``Smith is getting it started right now.''

``What plan?''

Pap eyed me slyly. ``You know that boy Tom Sawyer?''

Ice stabbed my belly.

``Never heard of him,'' I says.

``Don't you lie to me! I know you knows him.'' He reached into his
pocket and pulled out a ratty blue ribbon. At first I didn't recognize
it, but then I remembered all them months ago, when Tom and me was
spying up the tree at the bean feed. I gazed at the ribbon in horror.

``I seen the two of you, thick as thieves. He's what we call the darling
of the town. Everyone in the village kisses the ground he treads on.''

``So?''

``So, you hear what happened in Goshen a few months back?''

``No.''

``Town darling got borrowed. Held in a cave for five days, whole town
pitched in and paid \emph{two thousand dollars} to get her back.''

At that moment I did remember hearing water-pump gossip about it, back
when it happened in the summer. I felt ill. I reached out a hand and
touched the wall.

``Was it you, Pap? Did you steal the girl?''

``Might a done,'' he says craftily. ``And might be me and Smith hid the
yaller-boys in a secret spot no one knowed but us, and put the clues to
its hiding place in a coded advertisement so we'd remember how to find
it. And might be that you and your fool friend \emph{stole the coded
message} and now we can't recover the blooming cash.''

That dopey advertisement about the \emph{\textquotesingle Fabulous
Fabrizio\textquotesingle?} That explained why he and Smith had been so
anxious to get it back. If I recalled correctly, Tom made spitballs out
of it.

I slowly shook my head. ``How can you not remember where you hid the
money? Seems you'd remember something like that.''

``We had a touch of celebratory libations,'' says Pap. ``Anyhow, since
you lost us our hard-earned yaller-boys, I figure it\textquotesingle s
your fault we have to steal another darling. So you have to help; you
owe me.''

``You going to steal Tom Sawyer?''

``Already doing it,'' says Pap proudly. ``Where do you think Smith is?''

I shook like I \emph{} had the delirium tremens. ``I ain't helping.''

``If I says you're helping, you ignorant little know-nothing, you're
helping.''

``Ain't.''

Pap raised his hand, and I ducked.

The magi, Smith, arrived.

``Easy as pie,'' says Smith, grinning and rubbing his hands. ``I put a
note in the boy's room that said, \emph{``Please meet me at the old mill
by the river, signed Huck Finn.''}

He reached out to rub my head. I smacked his hand.

Pap growled. ``Ignoramus. My boy can't \emph{write.} I told you, you was
supposed to meow under the window. That's their little signal. Why
didn\textquotesingle t you meow?''

``Never mind,'' says Smith, all sulky. ``It worked. I seen him with my
own eyes, exiting the house. On his way here right now.''

Pap sighed. ``Fine. Head down to the dock, get the boat ready. I'll
bring the darling. But first I got a little more persuading to do with
my boy.''

Smith winked at me and left. And Pap came after me.

\mychapter{Chapter 17}\label{chapter-17}

I waked and my head ached. Pap was gone and the mill was empty, but it
smelled like a bonfire. Smoky.

I rubbed my head and tried to recollect what had happened. Pap told me I
had to help steal Tom Sawyer. I told him to go to blazes. He slapped me
around until his foot kicked the lantern over. In the sudden darkness I
tripped over my split bottom chair and my head hit the boards. I
couldn't remember more.

I saw a piece of paper next to me on the floor and picked it up. It was
covered in small, square coarse-hand. I shoved it in my pocket and
looked around, and realized I was surrounded by flames. Pap had set the
dern mill on fire.

``Not again,'' I says.

I looked for the door, but couldn't find it in the smoke. I was addled.
The walls and even the roof was covered in blankets of flame. Black
smoke bit the back of my throat like an adder.

I had to get out; I was being roasted alive. I covered my face with my
arms and ran.

I hit a wall. It was the one with a tiny window with no glass. I jumped
and catched the sill under my arms and heaved and wiggled my way
through.

It was a tight fit. I reckon I was more than a little cooked before I
made it out.

I hit the ground outside and rolled down the slope, coming to rest
against a rock.

Everything hurt.

I wanted to crawl into the dark woods and curl in the soft, cold dirt,
and sleep. But I couldn\textquotesingle t. Because I had a note in my
pocket which presumably said, \emph{``Meet me at the mill, signed Huck
Finn,''} and that meant that while I was taking a steamboat ride down
the river of fancy, Tom Sawyer had been there and got snatched. By my
own pap, no less.

I started towards town. It hurt so much to move that twice I had to stop
running and wait for the weakness to pass before I could keep on.

Finally I arrived at the docks and stood, listening. I couldn't see
nothing. It was black as pitch, the sky overcast, not even a single star
to shed a light.

I didn't see any boat. Maybe they was long gone. I didn't know how long
I'd been asleep.

Forgetting my recent personal acquaintance with fire, I crouched to the
ground to look for tracks.

You ever cook a potato in the coals then roll it out, and the skin is
cooked so dry and crisp it splits at the slightest pressure?

What the potato says, that is what I said.

I won't repeat the words; it ain't for your delicate ears.

Blinking away stars, I studied the dirt. There was so many different
tracks, it was impossible to tell if any was Smith or Pap's.

I heard a muffled curse. It came from out over the water.

Hope leaped into my throat. Without taking a moment to think, I hastened
along the dock and eased into the river.

The water felt like heaven on my skin. I closed my eyes and opened my
ears and surrendered to the language of the river.

The current pulled at my legs. The reeds and grasses whispered.
Somewhere far off, a fish jumped, landing with a faint splash. Something
had scared that fish.

I swimmed along gently, listening and feeling my way. I didn't see the
boat until I nearly rammed my face into its side.

Without thinking, I flinged up my arms. ``Help me in, quick!''

I clambered into the boat as graceful as a drownd cat. Smith didn't
help, but he didn't push me back under again, neither, which I took as a
real good sign. At first glance, it appeared the boat was empty except
for Smith. Where was Pap and Tom?

``What are you doing here, little man?'' Smith says in his goo-goo
voice. ``I thought you was too lily-livered to help.''

``Changed my mind,'' I says. ``Pap persuaded me.'' I fingered my busted
lip, and Smith giggled.

``Aye,'' he says, ``I seen Finn's arts of persuasion, and the cleverest
man in Missouri couldn't argue with him.''

I spied Tom Sawyer. He was trussed in the bottom of the boat like a hog,
with a gunnysack over his head. He warn't moving.

I wondered where Pap was.

``Where's Finn?'' says Smith.

Didn't he know, neither? That was interesting.

I decided it warn't too early to start muddying the waters. ``Pap told
me he'd catch up. He said to go on without him, he'd meet us there.'' I
didn't know where \emph{there} was, but I figured there was a there.

``Why?'' says Smith, suspicious.

``Well,'' I says, thinking hard, ``did you leave a note asking for the
money?''

Smith slapped his forehead. ``By jigs, I forgot.''

``Pap stayed behind to do it.''

Smith accepted this and pulled at the oars.

I figured Pap was the brains of this operation, and Smith was the
jump-to-it man, so the last thing we needed was for Pap and Smith to be
in the same place at the same time.

Pretty soon Smith flagged. ``I didn't know he could write.''

I cursed under my breath, because of course he couldn't.

``I wrote it,'' I says, closing my eyes.

I heard a slow creak, and my skin crawled. I opened my eyes. Smith
leaned into my face, leering. ``You're lying. Your pap told me you can't
write. I ever tell you how much I \emph{hate liars?}''

``Maybe I learned.''

``Prove it.''

My mind raced. I pulled out that note to Tom, covered in Smith's
coarse-hand print. It was damp, but serviceable. ``You got a pencil?''

I knowed he did, because he wrote it.

He handed it me. It was a stubby red cedar pencil with soft black lead.

``You know writing, or just coarse-hand?'' I says.

``I know both,'' he says, looking insulted. I took a chance and betted
he didn't, or he would a \emph{used} writing. No one \emph{chooses}
coarse-hand if they can write.

So I scribbled some long wiggles and loops on the back of the paper,
real pretty, and showed him. He says, ``Well, by jigs, you can,'' and he
picked up the oars again and put to. I flopped back in the stern and
breathed.

I looked behind us as St. Petersburg receded down the river, and
wondered if the alarm had been raised yet, and where Pap was. Then I
looked down in the bottom of the boat where Tom lay, all curled up and
so limp it scared the tar out of me.

\mychapter{Chapter 18}\label{chapter-18}

We rowed down the river to Hookerville, seven miles and with the
current. Got there just before break of day. To my relief Pap warn't
there waiting for us. I don\textquotesingle t know, I just had this fear
he would be.

We docked at an abandoned lumberyard at the edge of town. It was a big
yard of bare dirt, dotted with jagged stumps. Ugly as sin. A few
falling-down buildings hulked around. Smith scooped Tom in his long arms
like a praying mantis and led the way to one of these. He kicked in the
door and entered.

It was a single large room, once used for storing and cutting lumber.
The dirt floor was covered in wood dust and slivers of bark. There was a
rough fireplace on the far wall, with a skillet and coffee pot, and a
beat-up chair and table. Next to the fireplace was a big stockpile of
provisions in crates and sacks. Mostly food and water, but also I saw a
gun, and about a hundred boxes of percussion-caps and powder. Pap and
Smith must a set this up ahead of time.

Most disquieting, there was also a small wood coffin propped in one
corner. Well. I didn't care for \emph{that.}

Smith tossed Tom down on a pile of boards by the wall, all careless,
like he was a sack of rocks.

``Your job is to keep him alive till the money comes,'' Smith says to
me.

This was a job I was only too willing to accept. I scurried to Tom.

His hands was bound before him with rough rope, his feet with iron
shackles. The rope didn't worry me overmuch, unless it cut off the
blood. The shackles I hated, because I didn\textquotesingle t know who
had the key. Maybe Smith, maybe Pap. Without the key we
wouldn\textquotesingle t be able to make a run for it. And I sure as
fire warn\textquotesingle t leaving Tom alone for one solitary second
with that degenerate Smith. Not trussed and blinded.

I reached to take off the gunnysack. Smith darted over and slapped my
hand away.

``What are you at, you fool?'' His voice was a hiss.

``How am I supposed to keep him alive without taking it off?''

He pulled his lips back and showed his teeth. ``That's your lookout.''

He left to hide the boat, then came back inside and barricaded the door
with wood rammed up underneath, and checked the whole building over for
holes or windows, of which there was none. We was tighter than an oyster
in the shell.

Smith told me my other job was to cook our meals. I sorted through the
provisions and discovered enough food to feed a battalion. Sacks of
potatoes, cornmeal, coffee, a side of ham. We wouldn't starve, anyways.

I fried some potatoes for breakfast and hoped Tom would smell them and
wake up, but he didn't. So me and Smith eat, then slept through the day.

I waked when it got dark, and the only way I could tell was there was a
few chinks in the walls, slimmer'n my pinky, between the unfinished
boards.

Smith was still asleep. I crept to Tom.

Didn't look like he had moved at all since getting tossed there. I felt
for his pulse, and it was there. I saw his chest moving, gentle as a
kitten's.

I leaned near his head, still covered by the sack, and talked soft.
``It's me. If you hear, give a sign.''

He didn't give one.

I decided the least I could do was make him more comfortable. Shoot,
anything would be better than a pile of boards.

I pulled a ratty blanket out of the provisions and spread it in a
corner. Then I picked up Tom. His head lolled on my shoulder.

He was heavier than I expected, even though he was small. When a body's
limp it's pure dead weight. Soon as I picked him up I was afeard I would
drop him. I shuffled to the corner quickity-quick and lowered him,
trying not to bang his head. I arranged his arms and legs so he looked
kind of natural.

``There you go,'' I says as I worked. ``That ought to feel better. Sorry
about the sack. It'll be off soon.'' I patted where I judged the top of
his head was.

Smith waked and demanded supper; I cooked it.

I heard a sneeze from the corner, and rushed over. Tom hadn't shifted,
but I knowed it was him. He was awake!

Without thinking I started pulling off the sack.

``Hey,'' says Smith, leaping to his feet. ``What did I tell you?''

My insides quaked, but I held my voice steady. ``Got to feed him,'' I
says. ``It's my job.''

Smith considered, then slowly settled back, like a dog with its hackles
raised. He slurped his coffee, watching me over the rim.

I yanked off the sack. Tom's eyes was wrapped around several times with
a strip of filthy, frayed calico, possibly tore off a
girl\textquotesingle s dress. His mouth was gagged with a long, fat
twist of red material. A man's shirtsleeve.

``If you done all that, why bother with the sack?'' I says to Smith,
cross.

``So no one will recognize him,'' Smith says. That was when I decided
Smith and Pap didn't have a brain between them.

I poked Tom's cheek. ``You awake?''

He didn't twitch. His face was serene. But I knowed I heard him sneeze.
He was playing asleep, I guessed.

``Want some food?'' I glanced at Smith and said louder, for Smith's
benefit, ``I'll untie your mouth so you can eat. But you got to swear
not to yell out.''

Secretly I hoped he would yell as soon as I got it off.

Smith eyed us but didn't say nothing. With shaky fingers I fumbled at
the knot of the shirtsleeve. It was tight. I picked it apart with my
nails and pulled the grubby, soaked fabric out of his mouth.

He didn't yell out.

I brushed at his cheeks and the corners of his mouth, wet from the gag.
``You thirsty? Feeling ill? Talk to me.''

He didn't say a word.

I didn't know how much longer my nerves could stand this.

The night passed slow. I knowed they must a been looking for us by now,
but they had no way of knowing we was in Hookerville. We could a gone up
river just as easily, to Goshen or farther.

I wracked my brains, trying to figure how to give them a sign we was
there. If I could dangle a white kerchief out somewheres as a flag, it
might catch somebody's attention. But there warn't any windows, and I
didn't have no white kerchief.

The note! I still had that note Smith wrote, luring Tom to his
ambuscade. I pulled it from my pocket and unfolded it. It had got soaked
in the river. If the writing had been ink, it would a bled and been
useless. But it was pencil, so the words was still intact. I waited for
my chance.

Smith passed time playing cards with himself, gun across his lap,
muttering. He cursed Pap for not coming quicker and the townsfolk for
not paying up. He was antsy.

He drank a lot to comfort himself, and from that direction might come
our salvation, because it turned out likker had a real soporific effect
on him. He drank, then he dozed, and finally his chin hit his chest and
he let out a long, nasally \emph{hmmmmm,} and he was out.

Quick as a flash, I runned to a crack in the wall and shoved the note
out. Now if someone happened along and found it, they'd know we was in
there. It was the best I could do.

I slept fitfully, and waked to the most awful fright. Long, mournful
cries echoed through the walls, from far off. I assumed it was ghosts.
But gradually I made out a single word, over and over.

\emph{``Toooooom,''} the voices said.

They searched for us. I glanced uneasily to the corner, but Tom didn't
stir. Neither did Smith. I tried to go back to sleep, but how could I? I
was twitchier than a squirrel in a coffeepot.

When Smith waked we was hungry as hornets, so I fried a panful of
cornbread and took some to Tom.

He looked drawn and pale and kind of sickly. Sweat matted his hair, dirt
caked his clothes. He was in rough shape.

Why wouldn't he speak or give me some sign? Was he sore at me? Didn't he
know I was on his side? Come to think of it, I warn't doing much to save
him. I was pretty much only doing exactly what Smith wanted.

Never mind. Tom could be sore, he could think whatever he pleased, so
long as he lived to see the other side of this.

I held some cornbread to his lips. ``Eat,'' I says.

His mouth stayed shut. He breath moved his chest gently, like a sleeping
babe.

What a catfish. Suddenly I wanted to hold his nose and force his mouth
open, shove the food in. Poke him and tickle him and slap him, anything
to get a reaction.

I took a deep breath. I sat cross-legged on the blanket and put his head
in my lap and stroked his hair and his forehead, running my mouth,
trying to sound all relaxed and friendly-like. ``Tom Sawyer, won't you
please eat a little of this cornbread? I made it. It's good. Well, it's
not good--- it's pretty ornery---but you got to keep up your strength.
Think of your aunty and Siddy and Mary. How would they take it if you
was to show up all dead? I know if I had a brother, I'd be pretty tore
up if he got killed by starvation whilst on a kidnapping.''

No reaction. It was like talking to a window that was shut and bolted. I
sighed.

``But I don't have a brother,'' I says. ``It's just me. There will only
ever be one Huck Finn. And that's okay, I mean I reckon I'm up to the
job, but it sure would be a lot easier to be Huck Finn if there was a
Tom Sawyer, too. I know you must feel pretty wretched and uncomfortable,
because you're kidnapped and trussed like a turkey. You don't owe me
nothing. But all the same I would be greatly obliged if you was to eat a
little something so you didn't die and come back a ghost to haunt me for
the rest of my days for killing you. That's all.''

My eyelids drooped. I grasped one of his hands and pressed it. It was
cold and damp.

And his hand squeezed back! All of a sudden he was latched onto my
fingers with a grip of iron. I looked into his face. It revealed
nothing.

Hastily I snatched up the cornbread and held it to his lips, and he took
a tiny bite, and chewed, and swallowed. I grinned.

``Tom Sawyer, I could kiss you,'' I says, and done it. Planted one right
on him. I was so grateful to him for obliging me by not starving to
death.

I got him to nibble away all of the cornbread. Then I propped him higher
on my chest and held a tin cup of water to his lips so he could drink,
and he did. Drank the whole cupful, and another after that. I was
thrilled to pieces. I kept my arms around him and my chin in his hair
and squeezed him and rocked and laughed.

I was cracking up. Oh well. Nothing for it.

\mychapter{Chapter 19}

Smith started getting violent on account of his nerves was shot. I'd
seen it before with Pap and had sort of expected it. Except you could
generally see Pap coming a mile off, because he was slow and bumbly like
a bear. Smith's limbs was controlled by invisible wires. He'd sit there
grinning his oily grin, watching me cook or sweep or nurse Tom, and then
I'd walk too close and \emph{yank,} his arm would jerk out to slap me,
or his foot to trip me. It was like his limbs had a mind of their own.

Once I dropped the skillet with a clang, and he swung round with the gun
and fired it off and blowed a hole in the roof. And if he hadn't been
soused I lay he might a got me.

He got to ranting about how Pap done him wrong.

``He cooked up this scheme, then left me to handle it on my own.'' He
paced about, elbows swinging, neck bobbing.

``You're right,'' I says, stirring a stew over the fire. ``Pap's
prob'bly off in some warm tavern, having himself a grand time, waiting
for the yaller-boys to roll in.''

Smith gnashed his teeth.

``Now if I was Pap,'' I says, ``I would definitely cut you a share of
the payout. After all, you're the one doing most of the work.''

Smith stared at me. ``What do you mean? I \emph{am} getting a share.
Fifty-fifty, that's what Finn said.''

``Ain't what he told me,'' I says, fabricating recklessly. ``Told me
soon as we got the cash, we\textquotesingle d ditch you and take off for
Orleans.''

Smith's face went purple. He grabbed fistfuls of his long hair and
started yanking till I feared it would pull out.

``He was going to do a bunk? That---that---'' He started cussing out Pap
in the most flowerful language.

I smiled to myself.

Smith stormed around in a fury. He picked up the chair and threw it. He
kicked over the coffin. Then his eyes locked onto the corner where Tom
lay. Before I could twitch, he darted over and kicked Tom with all his
might. The toe of his boot landed square in Tom's gut.

Don't remember how I got there. Took less than a second. He was only
starting to turn when I jumped him. His knees buckled and his back
slammed the wall. I took hold of his neck with both hands.

I got up in his face, teeth bared. ``You do not do that again.''

``Let me go, you lunatic.'' His voice came out strangled. He scratched my
eyes furiously with his long, cracked nails. I jerked my head aside.

He twisted away and leaped, high-kneed, to the fireplace. He snatched up
the gun and swung it to point dead at my chest.

I stood.

``You do not lay one grubby paw on Tom Sawyer ever,'' I says,
pronouncing each word clear and distinct.

Smith\textquotesingle s shoulders heaved. His eyes darted wildly, hair
in clumps. The barrel of the gun trembled.

``You hear me?'' I walked forward till the muzzle dug into my chest.
``You hear?''

Smith pulled his lips back in a snarl. He looked like a cornered dog
trying to act fierce.

My face twisted with contempt. He was the worst kind of yaller.

The moment passed. Smith averted his face. The gun drooped to aim at the
floor.

``When's that stew ready,'' he says in a mutter. He sloped to the fire.
He reached for the pot, but I got there first and blocked him with my
arm, hand on the wall.

``I want you to say it.''

He wouldn't meet my eyes. He studied butt of the gun, rubbing it with
his thumb, sullen. ``Ain't gonna touch him.''

``Dern straight.''

\mychapter{Chapter 20}

One day a sharp knock came at the door. I reckoned we all nearly went
through the roof. Quick as a flash, Smith snatched Tom and tossed him
into the coffin. He slammed the lid and drove in a couple nails, while
the beating on the door continued.

He shoved me towards the pallet by the fireplace where he was accustomed
to sleeping, and told me to play along or I was dog meat.

``Open up,'' says the person at the door.

``Who's that?'' says Smith.

``Someone's been kidnapped. We're on a manhunt, searching every
building.''

Smith opened the door, arm hooked loose around the gun. I reckoned we
was going to have a shootout right there.

``Go away,'' he says. ``We got sick folk. Ain't safe.''

``Who's sick? Who are you? ''

``Smith. My two sons is sick. Leastways, one is sick and one died last
night.''

``Sorry to hear it,'' says the voice. ``May I come in? We're looking for
a boy who was kidnapped from St. Petersburg.''

``We ain't from there,'' says Smith. I heard the man push his way in.
``He's by the fire. Don't get too close, now.''

I heard the footsteps approach. I tried to play my part of being ill and
close to death.

``Goodness, what's wrong with him?'' says the voice. I judged he saw all
the scabs from my burns and welts and general griminess.

``Dunno. He got it bad, though,'' says Smith. I was desperate to see the
searchman, so I played like I was fitful, and flicked my eyelids and I
rolled my eyes, and catched a glimpse of him.

\emph{It was the general-store man!} The very man who accused me of
burglary last summer.

I forgot to be sick; I looked at him, and he looked at me. There was a
long pause.

``Hope your son gets better,'' says the general-store man. He turned and
walked out.

Smith cussed with relief and blockaded the door.

I shivered.

Smith didn't know the general-store man. But I knowed him, and he knowed
me.

Poor Smith---he must a been from a big city. He didn't realize everybody
in a small town knows everybody else.

So we was found at last. Now came the worst part---waiting for the final
bell. What would it be? I hoped there wouldn\textquotesingle t be a
gratuitous amount of shooting. A fellow could get damaged that way.

A tense, expectant stillness descended in our little hovel. It made me
twitchy.

What was they waiting for?

The temperature dropped. Steely air whistled through the cracks in the
walls. I reckoned it meant snow.

Smith sat by the fire and drank a whole bottle of something strong.
``When we get our money,'' he says, voice slurred, ``we'll keep going
down the river to Orleans. You and me.''

``What about Pap?''

``He left us,'' Smith says. He threw the bottle into the fireplace,
where it burst. ``I say, cut him out.''

``He'll be sore.''

``We won't let him find us. We'll pay people to keep him off. We'll be
rich as kings.''

``We'd also be on the run from the law.''

Smith growled. ``We'd pay them off too. Don't you get it, genius? We'd
be rich.''

I left him and went to the corner. ``What-ho, Tom,'' I says, flopping
down.

He was flushed as a tomato, his face misted with sweat, even though it
was cold in that corner. Every so often a shiver quaked through him. I
was sickened. It was ague if ever I saw it. And I didn't know how to fix
that.

I lifted his blouse and checked his stomach where he been kicked. It was
ugly black and blue.

``That was my fault,'' I says real quiet. ``I got him riled up.'' I
pulled his shirt down and swaddled him in his blanket like a moth in a
cocoon. Then I curled against his back and closed my eyes.

The likker finally reached Smith's head and he passed out. I never
thought I'd sleep, being so entertained imagining bullets tearing
through the hut and doors busting down and dogs with slavering fangs and
glowing eyes. But I guess I did sleep, because I waked when Smith kicked
my head.

I sat up, disoriented, looking around. The blanket beside me was empty.
The leg shackles sat neatly on top, key in the lock.

Smith grabbed my shirtfront and hauled me up. ``You useless, brainless
oaf.'' Flecks of spittle hit my face. He pointed to the corner of the
room where the provisions lived. A crate was shoved aside, and behind it
was concealed a hole, cut from the inside. A kitchen knife stuck in the
wall beside it.

And Tom Sawyer was gone.

I nearly cried.

``You let that monkey escape.'' Smith was screaming, veins bulging.
``Now we're dead men.''

He was so jumped-up I was afeard he'd kill me.

A knock came at the door.

``Constable Morley. Open up.''

Smith panicked. He dropped me and jumped to the hole behind the crate
and squirmed his long body through like a gangly mink. It was a small
hole, but the fear of death was upon him. He made it. As for me, I was
through that hole before you could say \emph{futile}.

``Find that scumalout,'' says Smith when we touched ground outside. ``If
he talks we're hanged.''

I figured we was hanged regardless, but I nodded. I did want to find
Tom.

Fresh snow lay about an inch deep on the ground, and there was
footprints in it. We followed them. They led into the houses, situated
about a hundred yards down the road. We followed the prints round a
house, down an alley, and between two shops. Then suddenly the prints
split in two directions.

That clever varmint! He must a doubled back and walked in his own
prints. I was so proud I could bust.

``You go that way,'' Smith says, so I took off following one set of
prints. I tracked them down a street and round a corner and through an
alley---\emph{where you going, Tom Sawyer}?---and then round another
corner and down a path, and I ran smack into Smith.

We was back at the lumberyard. We'd been led on a snipe hunt.

Constable Morley emerged from the shack with a pack of lawmen, and they
each had a dog.

``Get `em!'' the constable sings out.

We ran towards the river. The lawmen was a few hundred yards behind and
closing fast. We reached a pile of wood dross, and Smith dragged out his
boat. He shoved it into the water and scrambled in. ``Quick! They'll
kill us!''

I put one foot in the boat, then looked back. The lawmen released the
dogs.

The dogs streamed down the hill with eyes red and a thirst for blood.
Get in the boat, or get eat by dogs?

I got in the boat.

We rowed down the river, pulling hard. The lawmen couldn't keep up, the
brush was too thick. But the dogs paced us on the bank, snarling and
slavering, hoping we fell in. The second we did, they'd be in that water
and all over us.

``We'll go down to Memphis,'' Smith says. ``I know a fellow can hide us
till this blows over.''

I wondered how things got so far out of hand.

Assuming I survived the dogs, I warn't particluarly keen to end up in
court again. They wouldn't listen to my side. They didn't last time.

They would say I was a conspirator and a confederate and a kidnapper.
Forget the poorhouse. They'd put me in prison, or make me do hard labor,
or maybe even hang me.

For a moment I let myself imagine what life would be like if I went with
Smith. Freedom. Hiding out, running scared, stealing money, hurting
folk.

If I went with Smith now, I reckoned I would one day turn into
Pap---doing what the world expected of me and hating them for it.

I eyed the dogs on the bank. They showed their teeth, snapping, baying,
salivating for blood.

``I don't want to get eat by no dogs---'' I says.

``Course you don't, you brilliant piece of trash,'' says Smith.

``I don't want to get eat by no dogs,'' I says again, ``but if it's a
choice between you and the dogs...''

I stood and put my foot on the edge of the boat. ``Let's have the
dogs.''

I flipped us.

They was on us in a trice. They bit my arms and legs and clothes and
anything they could sink their teeth into. The lawmen shouted from the
shore and waved their clubs. It was a real danna-drag.

I felt pretty comfortable, all things considered.

\mychapter{Chapter 21}

They carted us back to St. Petersburg. It was a rough ride in that paddy
wagon, jostling with each bump in the road. Smith kept up a constant
rail against the lawmen and me and Pap and Tom and probably everybody he
ever met or ever would meet. I ignored him and explored my dog bites.

They dropped us at the jail, and there was a fat crowd waiting. I was
worried we was going to have a mob situation on our hands and get torn
to flinders. But they seemed content to palaver.

Pretty soon I noticed there was a lot of palavering about \emph{me} in
particular. A lot of the ladies of St. Petersburg testified to anyone
who would listen that I was a bad influence on their children, and the
apple don't fall far from the tree, and Clean Up the Streets, and so
forth. They appeared to be in favor of carting me away, and good
riddance to bad rubbage.

I plucked at the constable's sleeve. ``Where's Tom Sawyer?''

He brushed me off and kept talking with his men. Seems they was
concerned about putting Smith and me together in the cell.

``You seen Tom Sawyer?'' I says to a man in the crowd by the name of
Foster.

``Ain't telling \emph{you},'' he says with a heavy scowl.

Where's Tom at?'' I says to the lady who runs the Temperance Inn.

She glared at me. ``They brung him home this morning, starved to the
bone and out of his head with delirium. He's safe now, and no chance in
tarnation I'm telling you where he\textquotesingle s at.''

``We'll take him to Goshen,'' says the constable finally. ``They got
three cells up there.''

I assumed he was talking about Smith, but he grabbed me and steered me
towards the wagon.

Smith bolted for the woods. The lawmen pounced before he got five jumps
and piled on him. The crowd cheered. The lawmen tossed him into the jail
and slammed the door.

The constable loaded me into the wagon. I slumped on the seat and hid my
face in my hands.

The constable and driver climbed up front and whipped the horses into
motion.

I warn't overly thrilled to be in that wagon again so soon. The seat was
hard, and my dog bites hurt. It was going to be a long ride to Goshen.

We hadn't gone ten paces before the wagon jerked to a halt. The horses
shied and knickered.

``What the---'' says the driver. I crawled to the front and pressed my
face to the bars.

Tom Sawyer stood in the road before us. He was dressed in a long
tow-linen shirt and wrapped in a quilt. His feet was bare. He gazed at
the constable like a deer, and I could see him shaking from his head to
his toes.

He tried to say something, but squeaked. He coughed and rubbed his
mouth.

``You---'' His voice was barely there, scratchy as sandpaper. The crowd
and the constable and the driver and me, we all leaned in.

``You let me see him,'' Tom says.

The constable says, ``Sorry, son, I can't.''

Aunt Polly pushed her way to the front of the crowd. ``Come away this
instant, Thomas Sawyer. You\textquotesingle re out of your head with
ague.''

Tom shook his head. ``Let me see him.''

``It\textquotesingle s out of the question,'' says the constable.

At that moment Uncle Jake limped out of the crowd. Everyone stared in
surprise, because it was almost like they hadn't seen him till that
moment, like he'd been invisible.

Uncle Jake joined Tom in the street and put an arm round his shoulders.
``I know that boy you got in there, constable,'' Uncle Jake says.
``Can't speak for everything he done, but I can say he always been good
to me.''

The constable says, ``I need to take him.''

The crowd whispered and shifted and parted. A slight figure with yaller
hair stepped through, bundled in glossy brown furs. The Widow Douglas.

Folks looked at her with interest, and one woman curled her lip, but the
widow didn't cast them a glance. She went to the other side of Tom and
put her arm round him, too.

``Let Tom talk to him, constable,'' she says. ``Everybody knows they're
bosom friends. We've seen them all over town.''

The crowd looked down and rubbed their heads and shuffled their feets.

The constable considered the three figures on the road. He shook his
head like he never seen the like. ``All right then,'' he says.

He came round and unlocked the doors. He jerked his head at me. As for
me, I popped out of there like a jack-in-the-box.

I trotted round the wagon to Tom. He was thinner than I ever seen him,
with circles under his eyes. His cheeks was too red, eyes too bright.
His aunty was right; he was still in the grip of the fever.

``Hi-ya, Huck.'' He sounded like a frog.

I smiled sadly. ``Hi-ya, Tom.''

It was a little uncomfortable with everyone watching. They was so quiet
you could hear the wind clattering the branches of the trees.

Tom rubbed his nose on the quilt and grinned. ``Rum time, eh?''

I shook my head. ``Rum ain't the \emph{word}.''

Tom glanced around, and his gaze lingered on Aunt Polly. I thought for a
moment he lost his nerve.

But then he took a shaky breath and raised his voice. His eyes screwed
up with the effort. ``Harken up, you lot. I only aim to say this once,
before my throat combusts and I go up like a torch.

``Huck Finn is innocent. He warn't in cohoots with that orangutan Smith.
He was in the soup, same as I was. Except his feet warn't shackled, so
he could a runned away and saved himself any time. But he didn't. He
stayed. He protected me.''

He cast me a lovesome glance. ``It got real unnerving in there at times.
Honest truth, I might not be here if it warn't for him.''

He broke down coughing. Everyone kept froze like statues, hanging on his
words.

Tom turned to the constable. His shaking got worse, and I realized in
that moment that Tom Sawyer was terrified of the constable. Always had
been. Probably why he made up that story about him being a blood-taker.

Tom says to the constable, ``You don't lay one finger on Huck Finn. You
hear?''

He put his arms round me and pressed his cheek against my shoulder. He
glared at the constable. ``You hear?''

My heart swelled till I thought it would bust. I grinned.

The constable took his hat off and rubbed his head. He stared at the two
of us like he couldn\textquotesingle t believe the evidence.

He chuckled.

``All right, Tom Sawyer. I hear.''

\mychapter{Chapter 22}

Next morning I waked in the doctor's house, on a pallet by the fire. All
my bites was bandaged real tight and clean, and I wore a long nightgown
of the doctor's that reached my toes. Tom was draped along the doctor's
velvet Chesterfield sofa, and as soon as he saw I was awake, he
scrambled out of his covers.

``Oh, Huck,'' he says, sitting cross-legged by me. ``Warn't it
\emph{too---too---}wonderful!''

I could only blink and try to think what he meant.

``To be real and truly kidnapped,'' Tom says dreamily. ``Like something
straight out of a novel.''

``Sorry,'' I says, voice hoarse. Tom didn't hear.

``Of course, it would a been frankly beautiful if it could a gone on for
weeks. I could a gone proper skin-and-bones by then, and my hair would a
been long and caked like the Iron Mask, and I'd be haunted and
despairing. But I heard that blamed general store man barge in, and
knowed it was only a matter of time. So I decided to make my grand
escape. It would a been no fun at all to let them \emph{rescue} me.''

``I saw that hole you made behind the crates. How in blazes did you do
it?''

``Hacked away while you two slept, just a little bit every time, then
moved the crate back so you couldn\textquotesingle t tell. Oh, it was
glorious! Me, toiling away, thinking desperate thoughts, and the two of
you slumbering like babes! I had to let on I drugged your wine, of
course, to get around how dead asleep you was and how easy you made
it.''

``How\textquotesingle d you do that when you was tied up?''

Tom laughed. ``Oh Huck, you just warn\textquotesingle t made to be a
kidnapper. My hands was tied before me! It was the work of an infant to
reach up anytime I pleased and take off the blindfold.''

I frowned. ``Oh. Where'd you get the knife?''

Tom wiggled with joy. ``Hid it in my stocking before I left home. I got
`your note'---which was too, \emph{too} funny, because of course you
can't write. So I took a knife and went to the mill, just to spy it out
and see who was trying to lure me to my ultimate doom. And it looked
proper horrifying, I got not a light case of fan-tods, if you wish the
truth.''

``Why didn't you run away?''

``Well, I saw your face looking out from a window. So I went in.''

``My face,'' I says blankly.

``Once inside, I could see I'd been tricked, because you was lying there
stone cold, half in the shadows. So that meant someone had lifted your
head and held it above the windowsill, to make me \emph{think} you was
looking out.''

I shuddered.

``Then I heard a noise behind me, and spinned around and saw nobody. I
tripped over you and fell to the floor. Next thing I knowed, I was
clubbed cold. When I woke in the boat I was trussed like a trisket and
well and truly kidnapped!''

``I wonder when folks discovered you was gone?''

``Aunt Polly said she heard the commotion of folks running to see the
mill burn, so she checked Mary's and Sid's and my bedrooms and found me
missing, and she got the search going almost at once. They sent word up
and down the river for miles, so everyone was on the lookout, and then
some jake found that kidnap note in Hookerville, blowing in the wind.
And what a rum idea of dropping it, did you do it on purpose? You
\emph{did}? You clever critter!''

``Did they ever find Pap?''

Tom looked awkward. ``No.''

I sank back on my pallet, feeling weak.

``You're sweaty,'' Tom says. ``I'll call the doc.''

``Don\textquotesingle t bother. I just want to go home.'' Then it hit me
that my mill was gone, wiped off the face of the earth. It crushed me a
little.

I struggled to stand. Tom helped me, and I leaned on him and he leaned
on me, being still weak himself. We grinned at each other.

``Neither of you should be up,'' says a voice, and we spinned guiltily
to the door. It was the peddler. He came in and sat in a chair.

``Did you tell him what I said?'' he says to Tom.

I looked at Tom, mystified. Tom was all of a sudden real interested in a
run in his shirt, which he was still wearing like a nightdress.

``Slipped my mind,'' he says.

I looked from one to the other. ``What?''

``I'd like to offer you a position as my apprentice,'' says the peddler.
I gasped.

``You would learn a worthy trade, earn an honest wage, and see the
world. Or as much of it as can be seen by cart.'' He chuckled.

I looked at Tom, astounded. He went to the couch and started folding the
sheets real meticulous.

Suddenly the Widow Douglas stormed through the door. She went straight
to me and pinched my ear between her finger and thumb.

In a terrible voice, she says, ``Didn't I tell you---if you needed
me---anything at all---didn't I---was I not clear---''

I twisted my head, trying to keep from forfeiting the ear.

``Never mind,'' she says. ``I wrote a letter, and there's an opening at
the boys' academy in Macon. A sturdy education is what you need. I'll be
happy to pay the tuition.'' And she was off talking about quarterly dues
and performance reports and all sorts of rubbage. And right in the
middle, who should walk in but the constable.

``Hello, boys,'' he says, ``how are we this fine morning?''

I says, ``Is it finally time for you to send me to the poorhouse?''
Because I was tired of worrying myself sick over it. I was ready for him
to send me and have done with it.

He says, ``Poorhouse? I would never send you to a poorhouse. They're
reprehensible. They punish the poor for their poverty. Where on earth
did you get that idea?''

I looked at Tom. He had gone noticeably pale. He caught my eye and
shrugged.

The constable says, ``No, what you need is a firm foothold in life.
Nothing keeps a boy out of mischief like the military. I'll help you
secure a position in the youth division of the local militia. We'll
follow that with an illustrious stint in the armed forces, and then,
perhaps, a career in law. What do you say?''

I looked from one to the other. All three faces gazed at me expectantly,
waiting. I noticed Tom attending real close. He'd forgot he was making
his bed. He clutched the wad of bedclothes to his chest.

``What do you say, lad?'' says the peddler.

``Well, Huckleberry?'' says the Widow.

I went to my pallet by the fire and snatched my clothes. I dressed quick
as a lick and fastened my gallus and threw my arms into my coat. I
turned to them.

``I say, I don't give a plugged nickel what you lot do, but as for me,
I'm going fishing.''

I vaulted over the Chesterfield sofa and landed by the door. I turned
and thumbed my nose at the pack of them. Then I was out of the parlor,
down the hall, and exiting the house. Soon as I touched street I lit out
like Old Scratch himself was after me. And I didn't stop running till I
got to the river.

\mychapter{Chapter 23}

They never catched Pap, and I couldn't cipher out what happened to him.
No one saw him that night except me. Tom himself didn't see him. As the
days passed, my meeting with Pap took on a dreamlike quality, like maybe
I imagined him all up.

I found the green whiskey bottle half a mile down the river, embedded in
the mud. I didn't know what to make of it. Maybe he catched himself on
fire in the mill and ran down to the water and fell in? He was a bit
unsteady on his pegs, as I recall.

I doubted he was gone for good, though. Pap had a habit of recurring.

I visited the place by the river where my mill had been. It had burnt
flat to the ground, and nothing remained but a large blackened spot. It
was ugly, and made me real sick.

But by and by things would take root, and trees and vines and flowering
things would grow up in it and over it and cover it up like it had never
been. Little furry critters would build their burrows, and birds their
nests. The mill would be gone. But I reckoned that place would always
hold a certain peace there by the river.

I visited the constable the day after I ran out of the doctor's house
and told him everything. To my relief, he didn't question the truth of
the tale.

Before I left his office he stopped me. He stood by the door rubbing his
jaw.

``My brother was constable in Goshen when that girl was kidnapped.''

``I know it.''

``He got her home safe. Then he tracked your father and Smith to their
hideout. That night he was going to make his move. Before he could,
someone sneaked up behind him and shot him in the back.''

I looked down, shamefaced. ``Sorry. Is he---?''

``He died two weeks later from the infection.''

My face flooded with heat. ``Sorry.''

The constable nodded. ``Before my brother died he told me about your
father. Said he would probably slink back to St. Petersburg, that he had
ties here. So I took this job as constable to wait him out.

``Then I met you. I assumed you would be just as bad an egg. So I
watched you. I saw you around town. Picking fights, sleeping on the
streets, stealing food. It made me furious how your father neglected
you. I decided when I finally caught him, I would add child abuse to my
report.''

I was embarrassed. ``I ain't abused,'' I says, scuffing my foot. ``I'm
okay.''

He put a hand on my shoulder. ``You're better than okay. You might have
died from neglect. Or rotted to the heart, like your father. But you
didn't. You survived. And I look at you now, and\ldots''

I sneaked a glance at him. He was smiling. ``Huck Finn, you're all
right.''

I grinned and ducked out the door. All the way down the street I kept
grinning, all over, the biggest fool grin you ever seen.

I'd say the same about him.

Now that I didn't have to worry about the constable no more, things got
back to the way I liked them. I slept on doorsteps in fine weather, and
hogsheads in foul, and that suited me. I was glad to be in town again. A
fellow can get lonesome on his own all the time.

I went to see Uncle Jake a few days later. I catched him a big old
catfish in the morning, but forgot it at the last moment.

``I catched you a big catfish, but I forgot it,'' I says as soon as I
saw him.

``Never mind,'' he says.

He was sitting in the kitchen of the Rogers's farmhouse, whittling a new
lid for the butter churn. The old one broke in half; the two pieces sat
beside him on the bench. He was using them as a model to shape the new
one.

``Sit and visit a while,'' he says.

I plunked myself on the floor by his feet.

His eyes twinkled. ``I hear you been in a mess of trouble.''

I shrugged. I noticed his feet was covered in wood shavings and dust.

``I was worried for you,'' he says. ``I would a been pretty tore up if
anything happened to you.''

``Really? No fooling?'' I was pleased. Then I remembered something. I
looked down, picking at my toenail. I didn't know how to say it.

I pulled out the scrap of fabric I was in the habit of using as a
handkerchief. I started slowly cleaning wood debris off his feet.

``Uncle Jake\ldots{} I know it was my pap who beat you up that night. He
told me. I'm real sorry. And\ldots I'll understand if you want me to not
come round here no more.''

He reached down and lifted my chin. ``Why would I want that?''

``Because\ldots''

``You and me are friends,'' he says. He tweaked my nose. ``I want you to
come anytime. You're always welcome. You hear?''

``But---''

``You ain't your father,'' he says.

A lump came to my throat.

I finished cleaning his feet. Then I hugged him round the legs, hiding
my face in his knees. He patted my head.

When I left the Rogers place, who should I find but Tom Sawyer waiting a
few lengths down the road, kicking stones and looking mighty hot.

``There you are,'' he says at once. ``I been waiting hours for you.''

``Well, so what?''

``I reckon you wouldn't take that tone if you \emph{knowed}.''

``Knowed what?''

He grinned. ``We got one more piece of business to take care of.''

I shoved my hands in my pockets and kept walking. ``I ain't going to
beg. Either tell me or don\textquotesingle t, I got things to do.''

He jumped in front of me and thrust a piece of paper in my face. ``Are
you too busy to meet the FABULOUS Fabrizio?''

\mychapter{Chapter 24}

We rendezvous'ed at midnight, me and Tom Sawyer and Billy Fisher and Joe
Harper and Little James, each toting a shovel or spade as the mood
suited him, plus armfuls of lanterns and candles.

``Well, this is my expedition,'' says Tom, addressing the party, ``so
I'm in charge. Huck, you hold the paper.'' He bestowed it upon me, and
the other boys scowled.

``This is the paper we hooked from Pap and Smith?'' I says, examining it
close. ``The same one? I thought you said you made it into spitballs.''

Tom shrugged. ``Thought I had. I know I made spitballs out of something.
But Aunt Polly turned out the pockets of my roundabout before she washed
it, and she found it. She almost fed it to the fire, and that would a
been the end of it.''

The boys gasped. They'd been briefed on the contents of the document and
they hankered for treasure.

``Read it out, Huck,'' says Little James, staring at me with big eyes.

``Yes, read it aloud to us, Huck,'' says Billy with a sneer.

``Sure thing,'' I says, holding it close to my eyes. ``It says, `Billy
Fisher was born under a dunnakin.'\,''

``What! It does not.'' He pressed close to read over my shoulder. I
brought my hand up underneath the paper and smacked his nose.

He howled and clasped his face with both hands. I jeered.

``If you two are through being donkeys,'' says Tom.

I handed him the paper.

He cleared his throat. ``To whom it may concern: THIS MONTH only, for a
limited engagement, the FABULOUS Fabrizio, Witch Doctor of the Orient,
is here with his MARVELOUS Array of Potions and Remedies. FOUR DAYS
ONLY. Bring him your Sick and your Injured! He cures all ails with his
TEN STEP patented process. Inquire at NUMBER TWO, TWIN OAK Street, EAST
SIDE. Don\textquotesingle t Wait! Inquire today.''

It took Tom about two shakes to hunt out the meaning in this.

``It's old pie,'' he says. ``\emph{Ten step process}, that's ten paces.
Twin oaks, east side, that's where we start.''

I says, ``I know of about a thousand oaks in St. Petersburg.''

``Yes, but there's only one pair of \emph{twin} oaks worth mentioning,''
he says, ``and they're in front of the church.''

We went there.

``Now what?'' says Joe as we stood outside the church. He rubbed his
arms and shivered. ``How d'we know which one?''

``It says \emph{number two}---that means the second one. So we go ten
paces from the east side of the second tree.''

``Hold hard,'' I says. ``What was that bit about \emph{four days
only}?''

This made him pause.

``Well, first let's pace it out and see,'' he says, so we counted. I
took longer paces than Tom, and no one's paces was shorter than Little
James's, but it didn't matter, because we all ended up at---

``The church doors!'' Tom says, excited. He tugged one open and led the
way in.

It was dark and quiet inside, and smelled like new wood.

``I don't reckon we should dig in a church,'' says Little James in a
trembly voice.

``Even for treasure?'' says Billy.

``I'll do it,'' I says, real quick. I was getting kind of hot for gold.
Tom handed me a shovel.

``Just tell me where to dig,'' I says, hefting it. ``Don't forget,
\emph{four days only}.''

``Four paces?'' Tom mused. ``Four feet under?''

``How could the gin and magi have dug here?'' says Joe. He tapped the
tip of his shovel against the stone flags of the floor. It made a
depressing little \emph{cling, cling.}

``I ain't digging in no church,'' says Little James stubbornly. He
plunked himself down in one of them long benches and crossed his arms. I
nearly jumped through the roof.

``Holy catfish,'' I says. ``Fourth row!''

We crept up the aisle and examined the benches on the fourth row.

I found a lot of little books and papers in the pockets behind the
seats. I took them out and flipped through. Nothing fell out.

``These new pews are pretty slick,'' says Joe, rubbing his hand along
the top of one.

I groaned. ``New pews! They replaced the pews last month. So if Pap and
Smith hid the money in the old pews, it's lost.''

We all slumped along the bench, defeated.

``What do you reckon they did with the old pews?'' Little James asked
after a while.

``Burned,'' says Billy glumly. ``Or gave them to another church, in
another town.''

``Just a minute,'' says Tom slowly. I could see that gleam returning to
his eye. I sat up.

``The treasure is two thousand yaller-boys,'' he says. ``I don't reckon
the gin and magi could fit that much gold into a pew, even if they
stuffed the wooden pockets full. Anyways, people would see it during the
services. I myself sat in this row one Sunday and didn't notice nothing
unusual.''

``So what, Tom?'' says Joe.

``I reckon we picked the wrong oak.''

We went back out to the trees.

``Second tree,'' says Little James.

``Who's to say which is first and which is second?'' says Tom. So this
time we paced off ten steps from the east side of the other oak. We
arrived at a stand of trees out to the left of the church. There was
four trees.

``Four days only,'' I says.

Little James started hopping up and down. ``Where do we dig, Tom? Where?
Where? Where?''

``Right in the middle,'' says Tom, pointing. ``Each of you pick an
instrument and set to.''

We did, and Tom watched, leaning on his shovel.

``You didn't bring that to lean on, Tom Sawyer,'' says Billy after a
while.

``You'll allow me to dig?'' says Tom, all shocked. ``Here I was,
restraining myself from hogging the best part. Any fool knows the most
enjoyable part of a treasure hunt is the digging---the feel of the
shovel in your hands, the \emph{clang} of hitting \emph{} something hard
with the tip of your blade. But if you really want me to have all the
fun---''

``Don't listen to him,'' I says. I heaved a shovelful of dirt over my
shoulder and wiped my perspiring face, smearing it with dirt. ``That's
Tom Sawyer up to his usual tricks. He could convince you to walk off a
cliff, and pay him for the privilege.''

Tom jumped me. Our scuffle was brief but enthusiastic, and ended when my
forehead hit the blade of my own shovel and began to bleed.

``I think Tom's right,'' says Little James, as I slinked off to lick my
wounds. ``I want to dig, because I want to be the one who finds the
treasure.'' The other boys went back to digging too.

I leaned against a tree and crossed my arms. ``If the lambs \emph{want}
to be slaughtered, let `em have it, I don't care.''

Tom winked at me.

We dug in that circle of trees all night, trying here and there, and
finally around dawn Little James's shovel hit something. We dug it out,
and it was a big rusted can. Inside the can was\ldots{}

``Gold,'' says Little James, eyes gleaming. Every one of us reached a
hand in and let the cold metal dribble through our fingers.

``Two thousand golden American dollars,'' I says, shaking my head.

We hauled the can back to town in the light of the speedily-rising sun,
talking of what we'd do with our shares. I said I'd buy a ride on a
steamboat and see the ocean, but Tom Sawyer said why not think
\emph{big}, so I said, fine, I'd buy the ocean, and he said \emph{good
grief,} and rolled his eyes. But I was just plaguing him.

We skipped and horsed all the way home. We put gold coins in our eye
sockets and pretended to be dead, and into our mouths and pretended we
had gold teeth. It was a big night for us. We was intoxicated with our
wealth. We was kings.

So you can imagine how tore up we was when Tom returned the gold to the
mayor of Goshen. I'd forgot it was theirs, that they'd pooled it to
ransom that girl of theirs.

The rest of us boys wailed and sulked and mourned for days, and none of
us talked to Tom. The name Tom Sawyer was mud to us.

But every other inhabitant of St. Petersburg extolled Tom as a hero, and
Goshen did too for that matter, for months. So I reckon he got his
reward.

\mychapter{Chapter 25}

Tom and I sat on the bank of the river, watching steamboats glide past
all shiny and grand, belching black smoke from their twin stacks. Old
glories strung at their sterns snapped crisply in the breeze.

We fished, not catching nothing, which was fine by me. Catching fish is
the last reason you fish.

It was early spring but still cold. I was swaddled up to the ears in my
great coat and warm as a bug.

``I ain't catching squat,'' says Tom, tugging his line impatiently.
``Let's go play something.''

``Like what?''

``Well, like headless horsemen, or musketeers, or Iron Mask, or
blood-takers, or---''

``Say, don't you care to admit now that Constable Morley ain't no
blood-taker?''

Tom smiled slowly. ``Well, he could a been.''

The wind stirred the tree limbs overhead, creaking and clattering. I dug
my toes into the cold mud and sighed with pleasure.

``Let's go find Joe Harper,'' says Tom, poking my arm. ``We can play a
terrific new game I thought up about river pirates. We sail up and down
the river, and rob people, and burn steamboats, and---''

``Tom Sawyer,'' I says, all lazy-like, ``Shut it.'' And he did.

I leaned against the tree and closed my eyes. I didn't want the
afternoon to ever end.

So I reckon I'll stop talking now and enjoy the peace and quiet, because
I know Tom will be up again soon and pushing to play river pirates, or
some other fool game, but for now we can just enjoy the trees and the
river and the clouds skudding like heavy white steamboats across the
blue expanse of sky.

And if you ever need me, you know where to find me. Look for me by the
river. I know I'll be here---I'll always be here---promise. And Tom
Sawyer may be here, or he may have run off to play something, but he'll
be back, that's the good thing about old Tom.

So just sit next to me on the bank and we'll wait for him, and while we
wait we'll fish, and talk, and daydream. We'll listen to the wind
whispering in the reeds, and the sound of the water as the river keeps
on.

All my love,

HUCK

\cleardoublepage
\pagestyle{empty}

\backmatter
\pagestyle{empty}
\phantom{A}
\cleardoublepage
\pagestyle{empty}
\chapter*{About the Author}
Abigail Haynes lives on Galveston Island with her cat Silky Seymour in a historical building that survived the Great Storm of 1900. She writes a variety of tall tales, some in the style of Samuel Clemens, some not. Silky also likes to write, though he primarily types words like ``zdfggk'' and ``lllllhghs''. Undoubtedly this is very profound in the language of cats, however his wisdom remains impenetrable to humans.
\thispagestyle{empty}

\end{document}
